### 2022-03-03 tex/manuscript.tex
some additions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jctcce,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
%\usepackage{graphicx}
\usepackage{graphicx}
\usepackage{caption}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Sree ganesh balasubramani}
\affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
\author{Steven D. Schwartz}
\affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
%\author{I. Ken Groupleader}
%\altaffiliation{A shared footnote}
\email{sschwartz@email.arizona.edu}
%\phone{+123 (0)123 4445556}
%\fax{+123 (0)123 4445557}
%\affiliation[Unknown University]
%{Department of Chemistry, Unknown University, Unknown Town}
%\alsoaffiliation[Second University]
%{Department of Chemistry, Second University, Nearby Town}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]
  {Transition path sampling and the calculation of free energies of enzymatic
  reactions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{TPS,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{tocentry}

%Some journals require a graphical entry for the Table of Contents.
%This should be laid out ``print ready'' so that the sizing of the
%text is correct.

%Inside the \texttt{tocentry} environment, the font used is Helvetica
%8\,pt, as required by \emph{Journal of the American Chemical
%Society}.

%The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
%permitted for  \emph{Journal of the American Chemical Society}
%graphical table of content entries. The box will not resize if the
%content is too big: instead it will overflow the edge of the box.

%This box and the associated title will always be printed on a
%separate page at the end of the document.

%\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  We discuss a scheme for the calculation of free energies within 
  the transition path sampling (TPS) method. TPS is often used for
  the calculations of reaction coordinates of complex reactions in 
  the condensed phase and in this work we use it to sample the 
  equilibrium distribution of the reaction coordinate. This 
  distribution can then be used for the calculations of free energies
  using standard Boltzmann inversion schemes. We calculate the free
  energy profiles of the human Mat2a enzyme and the plasmodium 
  adenosine deaminase to demonstrate the applicability of this method
  for enzyme catalysis reactions.

  %This is an example document for the \textsf{achemso} document
  %class, intended for submissions to the American Chemical Society
  %for publication. The class is based on the standard \LaTeXe\
  %\textsf{report} file, and does not seek to reproduce the appearance
  %of a published paper.

  %This is an abstract for the \textsf{achemso} document class
  %demonstration document.  An abstract is only allowed for certain
  %manuscript types.  The selection of \texttt{journal} and
  %\texttt{manuscript} will determine if an abstract is valid.  If
  %not, the class will issue an appropriate error.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
\section{Theory}
\subsection{Transition path sampling}
%Transition path sampling is an importance sampling in the trajectory space.
Consider the molecular dynamics simulation of a molecular system consisting 
of $N$ atoms. Starting from the initial positions and momenta of each atom at time $t_0$,
the time evolution is carried out by finding the position ($\textbf{q}$) and 
momenta ($\textbf{p}$) of each atom at regular intervals of time $t_i = t_0 + i\Delta t$  
($i = 0,1,2,\ldots$). At each of the time slices, the positions and momenta can be collectively represented as
the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is run for a total 
time of $\mathcal{T}$ the number of time slices is given by $L = \mathcal{T}/\Delta t +1$ and for this sequence of 
times the state of the system can be represented as 
\begin{equation}
z(\mathcal{T}) = \{z_0, z_{\Delta t}, z_{2\Delta t},\ldots,z_{\mathcal{T}}\}
\end{equation}
The probability to obtain a particular sequence of states is determined by the initial 
conditions and the type of dynamics used for the time evolution. If the time evolution is 
Markovian, the probability to go from $z_{i\Delta t}$ to $z_{(i+1)\Delta t}$ depends only 
on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probablility can 
be expressed as the product of individual probabilities $p(i\rightarrow j)$ as 
\begin{equation}
\mathcal{P}[z(\mathcal{T})] = \rho(z_0)\Pi_{i=0}^{\mathcal{T}/\Delta t-1} p(z_{i\Delta t}\rightarrow z_{(i+1)\Delta t}),
\end{equation}
where $\rho(z_0)$ is the equilibrium distribution of the initial conditions and for a canonical ensemble this can be 
expressed as 
\begin{equation}
\rho(z_0) = \exp(-\beta H(z_0))/Z
\end{equation}
where 
\begin{equation}
Z = \int dz \exp(-\beta H(z)) 
\end{equation}
is the canonical partition function.

Within the transition path sampling, the trajectories of interest start in the reactant region of the 
phase space ($\mathcal{A}$) and end in the product region ($\mathcal{B}$). Such reactive trajectories have 
a restricted probability distribution function given by
\begin{equation}
\mathcal{P}_{\mathcal{AB}}[z_{\mathcal{T}}] = h_{\mathcal{A}}(z_0)\mathcal{P}[z(\mathcal{T})]
h_{\mathcal{B}}(z_{\mathcal{T}})/Z_{\mathcal{AB}}(\mathcal{T})\label{eqn:tpsensem}
\end{equation}
where 
\[
    h_{\mathcal{A}/\mathcal{B}}(z)= 
\begin{cases}
    1, & \text{if } z\in \mathcal{A}/\mathcal{B}\\
    0,              & \text{otherwise}
\end{cases}
\]
The set of all reactive trajectories characterized by the distribution function
given by Eq. \ref{eqn:tpsensem} is the transition path ensemble.

\subsection{Equilibrium distribution of order parameters}
%Monte carlo methods in trajectory space the probability of a path is 
%\begin{equation}
%P(x_A)
%\end{equation}
The free energy as a function of the order parameter $\zeta$ is defined
as 
\begin{equation}
\text{A}(\zeta) = -k_{\text{B}}T\text{ln}(P(\zeta)) + C, \label{eqn:fenergy}
\end{equation}
where $P(\zeta)$ is the probability distribution of the reaction coordinate
$\zeta$ and $C$ is an arbitrary constant. At equilibrium the probability distribution of the 
reaction coordinate 
can be obtained from the distribution function for the phase space $\rho(\textbf{q})$ as
\begin{equation}
P(\zeta) = \int d\textbf{V} \rho(\textbf{q})\delta\left[\zeta-\tilde{\zeta}(\textbf{q})\right].
\end{equation}
The integration is over the entire phase space. In practice the calculation of this distribution function
often proceeds using histogram based methods where the reaction coordinate is 
divided into bins and the frequency of the occurence of the reaction coordinate within a particular window
during the course of molecular dynamics simulations is used to calculate the probability distribution. 
Since some values of the reaction coordinates particulary close to the transition state are 
rarely sampled in a conventional MD simulations, enhanced sampling techniques such as the umbrella sampling
are necessary to sample the low probability regions. TPS is designed to sample these regions in phase space 
without the application of any external bias, but the trajectories sampled with TPS are only reactive and hence
they are not distributed according to the equilibrium distribution function. 
The restriction that the pathways start from the reactant state and end in the product state can be relaxed
to obtain an ensemble of trajectories that will be distributed according to the equilibrium distribution.

\subsection{The algorithm}
We follow the steps proposed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
to sample TPS trajectories within windows of the order parameter. 
\begin{itemize}
  \item An appropriate order parameter is chosen for the reaction of interest 
  that can distinguish the reactant and product states sufficiently well and the TPS ensemble is harvested 
  which consists only of reactive trajectories.
\item Divide the order parameter into windows $\{\zeta_i\}$ within which the 
order parameters are given by 
$\zeta_{i}^{min} < \zeta < \zeta_{i}^{max}$.
\item Choose a reactive trajectory $\tilde{z}(\mathcal{T})$ from the TPS ensemble as 
a guiding trajectory for the window sampling. Pick a time slice from this trajectory 
$\tilde{z}_{j\Delta t}$ such that the order parameter calculated for this time slice
$\tilde{\zeta}$ satisfies $\zeta_{i}^{min} < \tilde{\zeta} < \zeta_{i}^{max}$.
\item Using the shooting and shifting algorithm familiar from TPS simulations, harvest 
dynamics trajectories starting from the time slice $\tilde{z}_{j\Delta t}$ of the trajectory 
  $\tilde{z}(\mathcal{T})$ and accept the new trajectory if at any time slice the calculated 
  order parameter $\hat{\zeta}$ satisfies $\zeta_{i}^{min} < \hat{\zeta} < \zeta_{i}^{max}$.
  \item Calculate the probability distribution ($P_i(\zeta)$) within the window $\zeta_{i}^{min} 
  < \zeta < \zeta_{i}^{max}$ by constructing histograms obtained from the ensemble of 
  accepted trajectories for the window $i$. 
  \item Combine the probability distributions in successive windows by adjusting the constants $C$
  such that the free energy $A$ from Eq. \ref{eqn:fenergy} is continuous. 
\end{itemize}

\section{Results and discussion}
%\begin{figure}
%\centering
%\includegraphics[scale=0.2 ]{figures/ts253.png}
%\caption{Enzymatic reaction in human Mat2a enzyme}
%  \label{fgr:mat2a_reaction}
%\end{figure}
\section{Adenosine deaminase reaction coordinates and free energy}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/reactant-state.png}
\caption{Reactant state}
\label{fig:minipage1}
\end{minipage}
\quad
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/int1.png}
\caption{Intermediate state}
\label{fig:minipage2}
\end{minipage}
\quad
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/product-state.png}
\caption{Product state}
\label{fig:minipage1}
\end{minipage}
\end{figure}

%\subsection{Outline}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[scale=0.5]{figures/reac-sc-oc-167.png}
\caption{Trajectory 167}
\label{fig:minipage1}
\end{minipage}
\quad
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[scale=0.9]{figures/den-fenergy-1.pdf}
\caption{Free energy profile for the enzymatic reaction in human Mat2a enzyme}
\label{fig:minipage2}
\end{minipage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}

\end{suppinfo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{manuscript}

\end{document}

### 2022-03-24 tex/manuscript.tex
added some new stuff

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jctcce,manuscript=article]{achemso}
+\documentclass[journal=jpcb,manuscript=article]{achemso}
 
@@
 \section{Introduction}
+Atomistic level study of the mechanism of enzyme catalysis is necessary
+to understand the efficiency of enzymes as well as to device strategies for inhibiting 
+their activity. 
 Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
+Human Mat2a enzyme and plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008} 
 \section{Theory}

### 2022-03-25 tex/manuscript.tex
added some more references

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jpcb,manuscript=article]{achemso}
+\documentclass[journal=jpcb,manuscript=article,reprint]{achemso}
 
@@
 \usepackage[T1]{fontenc} % Use modern font encodings
-%\usepackage{graphicx}
 \usepackage{graphicx}
@@
 Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
+is an unbiased monte carlo method in the trajectory space which is used for the calculation 
+of reaction coordinates and transition states of complex chemical reactions. 
+TPS uses reactive trajectories which connect a reactant state to a product state and these 
+trajectories are rare in molecular dynamics simulations because of the energy 
+barriers separating the reactants and the products. To overcome these energy barriers and 
+to obtain ergodic sampling several methods have been developed over the years such as 
+umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
+steered MD, \cite{Park04JChemPhys120p5946}
+milestoning, etc. These methods rely on applying bias
+potentials for the system to overcome energy barriers and hence require the knowledge of 
+the reaction coordinates.
+
+TPS based free energy simulations have been attempted previously by Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
+
 Human Mat2a enzyme and plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008} 
@@
 \section{Adenosine deaminase reaction coordinates and free energy}
+\begin{figure}[ht!]
+\centering
+\includegraphics[scale=0.25]{./figures/ada-protein.png}
+\end{figure}
+
 \begin{figure}[ht]

### 2022-03-29 tex/manuscript.tex
added some more corrections

@@
 \usepackage{caption}
+\usepackage{amsmath}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Human Mat2a enzyme and plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008} 
+One example of the TPS based free energies calculations on real protein systems \cite{Silveira21JPhysChemB125p2018}
+
+Human Mat2a enzyme and plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618} 
 \section{Theory}
@@
 \section{Results and discussion}
-%\begin{figure}
-%\centering
-%\includegraphics[scale=0.2 ]{figures/ts253.png}
-%\caption{Enzymatic reaction in human Mat2a enzyme}
-%  \label{fgr:mat2a_reaction}
-%\end{figure}
+\subsection{QMMM simulations}
+CHARMM force field \cite{Brooks09JComputChem30p1545}
+
 \section{Adenosine deaminase reaction coordinates and free energy}

### 2022-03-29 tex/manuscript.tex
added Make file

@@
 \section{Introduction}
-Atomistic level study of the mechanism of enzyme catalysis is necessary
-to understand the efficiency of enzymes as well as to device strategies for inhibiting 
-their activity. 
-Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
-is an unbiased monte carlo method in the trajectory space which is used for the calculation 
-of reaction coordinates and transition states of complex chemical reactions. 
-TPS uses reactive trajectories which connect a reactant state to a product state and these 
-trajectories are rare in molecular dynamics simulations because of the energy 
-barriers separating the reactants and the products. To overcome these energy barriers and 
+Atomistic level detailed study of the mechanism of enzyme catalysis 
+reactions is necessary to understand their efficiency as well as to 
+design molecules that can act as inhibitors for their activity. 
+Molecular dynamics (MD) simulations 2 have often been used to study 
+large biomolecular systems in atomistic detail since they provide an 
+excellent computational cost to detail ratio. MD simulations to study 
+chemical reactions that go through free energy barriers are often 
+impractical to simulate since it would take a long time for the system 
+to cross over barriers.
+To overcome these energy barriers and 
 to obtain ergodic sampling several methods have been developed over the years such as 
-umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
+umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
+metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
 steered MD, \cite{Park04JChemPhys120p5946}
@@
 the reaction coordinates.  
+Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
+(TPS) is an enhanced sampling method for 
+rare events that transition a system between two stable states often 
+referred to as reactants and products. 6
+
+TPS focuses on transition paths and not on stable states which introduces 
+biases in the sampling of equilibrium distribution of the reaction coordinate.
 

### 2022-03-31 tex/manuscript.tex
modified the introduction a little more

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jpcb,manuscript=article,reprint]{achemso}
+\documentclass[journal=jpcbfk,manuscript=article,layout=traditional]{achemso}
 
@@
 \usepackage{amsmath}
+\usepackage{mathptmx}
+\usepackage[scaled=0.92]{helvet}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}
+  Transition path sampling (TPS) has been widely applied for the
+  calculations of transition state and reaction coordinates of 
+  condensed phase systems. Here, we discuss a scheme for the calculation
+  of free energies with the TPS method 
   We discuss a scheme for the calculation of free energies within 
@@
 \section{Introduction}
-Atomistic level detailed study of the mechanism of enzyme catalysis 
-reactions is necessary to understand their efficiency as well as to 
-design molecules that can act as inhibitors for their activity. 
-Molecular dynamics (MD) simulations 2 have often been used to study 
-large biomolecular systems in atomistic detail since they provide an 
-excellent computational cost to detail ratio. MD simulations to study 
-chemical reactions that go through free energy barriers are often 
-impractical to simulate since it would take a long time for the system 
-to cross over barriers.
+Molecular dynamics (MD) simulations are being 
+used to calculate quantitative estimations of experimentally measurable 
+kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
+\cite{Karplus02NatStructMolBiol9p646,
+Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
+Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
+This has helped in proposing mechanisms and explanations for 
+enzyme activity, high throughput screening of millions of inhibitor 
+molecues in drug discovery,
+
+This is quite useful in explaining 
+the efficiency and rate of such reactions, developing 
+inhibitors for various enzymes in pharmacology, de novo design of 
+enzymatic activity.
+MD simulations model chemical reactions by taking the system from a reactant state 
+to a product state via a high energy transition state.
 To overcome these energy barriers and 
@@
 Human Mat2a enzyme and plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618} 
-\section{Theory}
+\section{Methods}
 \subsection{Transition path sampling}
@@
 \centering
-\includegraphics[scale=0.25]{./figures/ada-protein.png}
+\includegraphics[scale=0.22]{./figures/ada-protein.png}
 \end{figure}
 
-\begin{figure}[ht]
-\centering
-\begin{minipage}[b]{0.32\linewidth}
-\includegraphics[scale=0.3]{figures/reactant-state.png}
-\caption{Reactant state}
-\label{fig:minipage1}
-\end{minipage}
-\quad
-\begin{minipage}[b]{0.32\linewidth}
-\includegraphics[scale=0.3]{figures/int1.png}
-\caption{Intermediate state}
-\label{fig:minipage2}
-\end{minipage}
-\quad
-\begin{minipage}[b]{0.32\linewidth}
-\includegraphics[scale=0.3]{figures/product-state.png}
-\caption{Product state}
-\label{fig:minipage1}
-\end{minipage}
-\end{figure}
+%\begin{figure}[ht]
+%\centering
+%\begin{minipage}[b]{0.32\linewidth}
+%\includegraphics[scale=0.3]{figures/reactant-state.png}
+%\caption{Reactant state}
+%\label{fig:minipage1}
+%\end{minipage}
+%\quad
+%\begin{minipage}[b]{0.32\linewidth}
+%\includegraphics[scale=0.3]{figures/int1.png}
+%\caption{Intermediate state}
+%\label{fig:minipage2}
+%\end{minipage}
+%\quad
+%\begin{minipage}[b]{0.32\linewidth}
+%\includegraphics[scale=0.3]{figures/product-state.png}
+%\caption{Product state}
+%\label{fig:minipage1}
+%\end{minipage}
+%\end{figure}
 
 %\subsection{Outline}
-\begin{figure}[ht]
+\begin{figure}[ht!]
 \centering
 \begin{minipage}[b]{0.45\linewidth}
-\includegraphics[scale=0.5]{figures/reac-sc-oc-167.png}
+\includegraphics[scale=0.4]{figures/reac-sc-oc-167.png}
 \caption{Trajectory 167}
@@
 \begin{minipage}[b]{0.45\linewidth}
-\includegraphics[scale=0.9]{figures/den-fenergy-1.pdf}
+\includegraphics[scale=0.8]{figures/den-fenergy-1.pdf}
 \caption{Free energy profile for the enzymatic reaction in human Mat2a enzyme}

### 2022-03-31 tex/manuscript.tex
added some more stuff

@@
 \section{Introduction}
-Molecular dynamics (MD) simulations are being 
+Molecular dynamics (MD) simulations are increasingly being
 used to calculate quantitative estimations of experimentally measurable 
@@
 This has helped in proposing mechanisms and explanations for 
-enzyme activity, high throughput screening of millions of inhibitor 
-molecues in drug discovery,
+enzyme activity, fast high throughput screening of thousands of inhibitor 
+molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
+directed evolution for designing enzymes with enhanced catalytic 
+activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
+Straightforward MD simulations are not employed in these cases
 
-This is quite useful in explaining 
-the efficiency and rate of such reactions, developing 
-inhibitors for various enzymes in pharmacology, de novo design of 
-enzymatic activity.
-MD simulations model chemical reactions by taking the system from a reactant state 
-to a product state via a high energy transition state.
-To overcome these energy barriers and 
-to obtain ergodic sampling several methods have been developed over the years such as 
+Ergodic sampling of the phase space within straighforward MD simulations 
+is often difficult to achieve with the presence of energy barriers. 
+To overcome such barriers enhanced sampling techniques 
+have been developed over the years such as 
 umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
@@
 steered MD, \cite{Park04JChemPhys120p5946}
-milestoning, etc. These methods rely on applying bias
-potentials for the system to overcome energy barriers and hence require the knowledge of 
-the reaction coordinates.  
+milestoning, \cite{Faradjian04JChemPhys120p10880} etc. These methods rely on applying bias
+potentials for the system to overcome energy barriers and hence 
+require the knowledge of the reaction coordinates.  
 Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
@@
 rare events that transition a system between two stable states often 
-referred to as reactants and products. 6
+referred to as reactants and products.
 
@@
 
-TPS based free energy simulations have been attempted previously by Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
+TPS based free energy simulations have been attempted previously by 
+Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
 
-One example of the TPS based free energies calculations on real protein systems \cite{Silveira21JPhysChemB125p2018}
-
-Human Mat2a enzyme and plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618} 
+In this work we will focus on the implementation and application of the 
+free energy based calculations within the TPS method on two enzymatic reactions: 
+One example of the TPS based free energies calculations on real 
+protein systems \cite{Silveira21JPhysChemB125p2018} Human Mat2a enzyme and 
+plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
+%------------------------------------------------------------------------------ 
 \section{Methods}
+%------------------------------------------------------------------------------ 
 \subsection{Transition path sampling}
+%------------------------------------------------------------------------------ 
 %Transition path sampling is an importance sampling in the trajectory space.
@@
 to obtain an ensemble of trajectories that will be distributed according to the equilibrium distribution. 
-
+%------------------------------------------------------------------------------ 
 \subsection{The algorithm}
+%------------------------------------------------------------------------------ 
 We follow the steps proposed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
@@
 \end{itemize}   
-
+%------------------------------------------------------------------------------ 
 \section{Results and discussion}
+%------------------------------------------------------------------------------ 
 \subsection{QMMM simulations}
+%------------------------------------------------------------------------------ 
 CHARMM force field \cite{Brooks09JComputChem30p1545}
 
-\section{Adenosine deaminase reaction coordinates and free energy}
+%------------------------------------------------------------------------------ 
+\subsection{Adenosine deaminase reaction coordinates and free energy}
+%------------------------------------------------------------------------------ 
 \begin{figure}[ht!]

### 2022-04-05 tex/manuscript.tex
added a figure and some text

@@
 activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
-Straightforward MD simulations are not employed in these cases
+Straightforward MD simulations are not employed in these cases because 
+trajectories that take the system from reactant state to product state through
+a transition state are rare and would take long time
 
@@
 TPS focuses on transition paths and not on stable states which introduces 
-biases in the sampling of equilibrium distribution of the reaction coordinate.
+biases in the sampling of equilibrium distribution of the reaction coordinate. \cite{Dellago09AdvCompSimAppp167}
 
@@
 $\tilde{\zeta}$ satisfies $\zeta_{i}^{min} < \tilde{\zeta} < \zeta_{i}^{max}$.
-\item Using the shooting and shifting algorithm familiar from TPS simulations, harvest 
-dynamics trajectories starting from the time slice $\tilde{z}_{j\Delta t}$ of the trajectory 
+\item Using the shooting and shifting algorithm familiar from TPS simulations, \cite{dellago02AdvChemPhys123}
+harvest dynamics trajectories starting from the time slice $\tilde{z}_{j\Delta t}$ of the trajectory 
   $\tilde{z}(\mathcal{T})$ and accept the new trajectory if at any time slice the calculated 
@@
 \includegraphics[scale=0.22]{./figures/ada-protein.png}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[scale=0.9]{./figures/ada-fenergy.pdf}
 \end{figure}

### 2022-04-07 tex/manuscript.tex
added some more figures and text

@@
 \title[]
-  {Transition path sampling and the calculation of free energies of enzymatic
+  {Transition path sampling based calculation of free energies of enzymatic
   reactions}
@@
 activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
-Straightforward MD simulations are not employed in these cases because 
-trajectories that take the system from reactant state to product state through
-a transition state are rare and would take long time
-
-Ergodic sampling of the phase space within straighforward MD simulations 
-is often difficult to achieve with the presence of energy barriers. 
-To overcome such barriers enhanced sampling techniques 
-have been developed over the years such as 
+Straightforward MD simulations adequately sample 
+the long-lived stable reactant and product states 
+but the high energy barrier-crossing events that occur rarely and at long 
+time scales are not easily accessible using this method alone.
+To access such states at a higher frequency using MD simulations 
+several enhanced sampling techniques 
+have been developed in the past few decades some of which are 
 umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
@@
 steered MD, \cite{Park04JChemPhys120p5946}
-milestoning, \cite{Faradjian04JChemPhys120p10880} etc. These methods rely on applying bias
-potentials for the system to overcome energy barriers and hence 
-require the knowledge of the reaction coordinates.  
-Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
-(TPS) is an enhanced sampling method for 
-rare events that transition a system between two stable states often 
-referred to as reactants and products.
+milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
+(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
 
-TPS focuses on transition paths and not on stable states which introduces 
-biases in the sampling of equilibrium distribution of the reaction coordinate. \cite{Dellago09AdvCompSimAppp167}
+TPS in particular is an attractive method since it samples reactive trajectories
+in which the rare barrier-crossing events are guaranteed to occur and
+it does not require a priori knowledge of the reaction coordinate which can be 
+a complicated entity in large biological systems. TPS has been successfully 
+applied to study problems such as enzyme catalysis, protein folding, etc. 
+The TPS ensemble which is used to obtain a unique statistical definition for the reaction coordinate
+and transition state of a reaction consists of so called reactive trajectories which 
+link the stable reactant and product states. In this Monte-Carlo procedure in the trajecotry space
+rejected moves correspond to non-reactive trajectories. But these trajectories are 
+required to obtain the equilibrium distribution function.
+
+The TPS ensemble consists of reactive trajectories which connect a reactant state to
+a product state by going through the transition state and the probabilities of configurations thus sampled 
+do not replicate the equilibrium probabilities. To calculate equilibrium properties such as the free energy 
+of a reaction, equilibrium sampling of the configurations along a reaction coordinate is 
+required which is not obtained with straightforward TPS. \cite{Dellago09AdvCompSimAppp167}
 
 TPS based free energy simulations have been attempted previously by 
-Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
+Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111,Dellago04MolSimul30p795}
 
@@
 %------------------------------------------------------------------------------ 
+\subsection{Human MethionineAdenosyltransferase 2A}
+%------------------------------------------------------------------------------ 
+The SN2 reaction catalyzed by the human Mat2a enzyme leads to the production of 
+S-Adenosylmethionine compound which is part of the ATP cycle \cite{Niland21Biochem60p791}
+\begin{figure}[ht!]
+\includegraphics[scale=0.5]{figures/comm-80-mat2a.pdf}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
+\caption{The free energy as a function of the order parameter.}
+\end{figure}
+%------------------------------------------------------------------------------ 
 \subsection{Adenosine deaminase reaction coordinates and free energy}
@@
 %\subsection{Outline}
-\begin{figure}[ht!]
-\centering
-\begin{minipage}[b]{0.45\linewidth}
-\includegraphics[scale=0.4]{figures/reac-sc-oc-167.png}
-\caption{Trajectory 167}
-\label{fig:minipage1}
-\end{minipage}
-\quad
-\begin{minipage}[b]{0.45\linewidth}
-\includegraphics[scale=0.8]{figures/den-fenergy-1.pdf}
-\caption{Free energy profile for the enzymatic reaction in human Mat2a enzyme}
-\label{fig:minipage2}
-\end{minipage}
-\end{figure}
+%\begin{figure}[ht!]
+%\centering
+%\begin{minipage}[b]{0.45\linewidth}
+%\includegraphics[scale=0.4]{figures/reac-sc-oc-167.png}
+%\caption{Trajectory 167}
+%\label{fig:minipage1}
+%\end{minipage}
+%\quad
+%\begin{minipage}[b]{0.45\linewidth}
+%\includegraphics[scale=0.8]{figures/den-fenergy-1.pdf}
+%\caption{Free energy profile for the enzymatic reaction in human Mat2a enzyme}
+%\label{fig:minipage2}
+%\end{minipage}
+%\end{figure}
 

### 2022-04-07 tex/manuscript.tex
added some more figures

@@
 
+TPS was developed as a reaction-coordinate free, statistical method 
+which samples trajectories that connect the stable reactant and product states 
+guaranteeing the occurence of the rare barrier crossing events. These trajectories 
+are sampled using a Monte carlo procedure leading to an ensemble of transition
+pathways that are real dynamical trajectories which contain information regarding the 
+mechanism and kinetics of the reaction.
+
 TPS in particular is an attractive method since it samples reactive trajectories
@@
 applied to study problems such as enzyme catalysis, protein folding, etc. 
-The TPS ensemble which is used to obtain a unique statistical definition for the reaction coordinate
-and transition state of a reaction consists of so called reactive trajectories which 
-link the stable reactant and product states. In this Monte-Carlo procedure in the trajecotry space
+ In this Monte-Carlo procedure in the trajecotry space
 rejected moves correspond to non-reactive trajectories. But these trajectories are 
@@
 
-In this work we will focus on the implementation and application of the 
-free energy based calculations within the TPS method on two enzymatic reactions: 
+In this manuscript we will focus on the implementation and application of the 
+free energy based calculations within the TPS method on reactions catalyzed by two systems:
+the human MAT2A enzyme and the plasmodium vivax adenosine deaminase enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
+These enzymes were chosen not only because of their importance in biochemical pathways and
+but also because of the availability of experimental results with which comparisons can 
+be drawn.
+
 One example of the TPS based free energies calculations on real 
 protein systems \cite{Silveira21JPhysChemB125p2018} Human Mat2a enzyme and 
-plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
+plasmodium vivax adenosine deaminase 
 %------------------------------------------------------------------------------ 
@@
 %------------------------------------------------------------------------------ 
-%Transition path sampling is an importance sampling in the trajectory space.
+Here we briefly describe the equations describing the TPS method for the purpose of
+introducing notations as well as for the sake of being complete. 
 Consider the molecular dynamics simulation of a molecular system consisting 
@@
 %------------------------------------------------------------------------------ 
-\subsection{Human MethionineAdenosyltransferase 2A}
+\subsection{Human Methionine Adenosyltransferase 2A}
 %------------------------------------------------------------------------------ 
@@
 S-Adenosylmethionine compound which is part of the ATP cycle \cite{Niland21Biochem60p791}
+\begin{figure}[ht!]
+\includegraphics[scale=0.25]{figures/mat2a-transition-state.png}
+\end{figure}
+
 \begin{figure}[ht!]

### 2022-04-08 tex/manuscript.tex

@@
   using standard Boltzmann inversion schemes. We calculate the free
-  energy profiles of the human Mat2a enzyme and the plasmodium 
-  adenosine deaminase to demonstrate the applicability of this method
-  for enzyme catalysis reactions.
-
-  %This is an example document for the \textsf{achemso} document
-  %class, intended for submissions to the American Chemical Society
-  %for publication. The class is based on the standard \LaTeXe\
-  %\textsf{report} file, and does not seek to reproduce the appearance
-  %of a published paper.
-
-  %This is an abstract for the \textsf{achemso} document class
-  %demonstration document.  An abstract is only allowed for certain
-  %manuscript types.  The selection of \texttt{journal} and
-  %\texttt{manuscript} will determine if an abstract is valid.  If
-  %not, the class will issue an appropriate error.
+  energy profiles of the reactions catalyzed by the human methionine 
+  S-adenosyltransferase (MAT2A) enzyme and the plasmodium 
+  vivax adenosine deaminase enzyme to demonstrate the applicability of this 
+  method. 
 \end{abstract}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 free energy based calculations within the TPS method on reactions catalyzed by two systems:
-the human MAT2A enzyme and the plasmodium vivax adenosine deaminase enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
+the human MAT2A enzyme and the plasmodium vivax adenosine deaminase 
+enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
 These enzymes were chosen not only because of their importance in biochemical pathways and
@@
 \subsection{QMMM simulations}
-%------------------------------------------------------------------------------ 
-CHARMM force field \cite{Brooks09JComputChem30p1545}
+%------------------------------------------------------------------------------ We performed computational studies on two enzymes where we have shown
+that fast dynamics is part of the reaction mechanism and calculated the electric field near
+the bond-breaking event.
+QMMM simulations were used for the TPS simulations in this work and the 
+CHARMM force field \cite{Brooks09JComputChem30p1545} is used for the 
+MM portion and the PM3 semiempirical method \cite{Repasky02JComputChem23p1601} 
+is used for the QM region.
 
@@
 %------------------------------------------------------------------------------ 
-The SN2 reaction catalyzed by the human Mat2a enzyme leads to the production of 
-S-Adenosylmethionine compound which is part of the ATP cycle \cite{Niland21Biochem60p791}
+Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
+reaction between Adenosine triphosphaste and L-methionine which leads to the 
+formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
+The mechanism of this reaction is determined to follow a SN$_2$ transition state
+with the sulphur atom of the L-methionine as the nucleophile with the 
+triphosphate leaving group.
+
 \begin{figure}[ht!]
@@
 \begin{figure}[ht!]
-\includegraphics[scale=0.5]{figures/comm-80-mat2a.pdf}
+\includegraphics[scale=0.5]{figures/comm-60-mat2a.pdf}
 \end{figure}
@@
 %\end{figure}
-
-%\subsection{Outline}
-%\begin{figure}[ht!]
-%\centering
-%\begin{minipage}[b]{0.45\linewidth}
-%\includegraphics[scale=0.4]{figures/reac-sc-oc-167.png}
-%\caption{Trajectory 167}
-%\label{fig:minipage1}
-%\end{minipage}
-%\quad
-%\begin{minipage}[b]{0.45\linewidth}
-%\includegraphics[scale=0.8]{figures/den-fenergy-1.pdf}
-%\caption{Free energy profile for the enzymatic reaction in human Mat2a enzyme}
-%\label{fig:minipage2}
-%\end{minipage}
-%\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-04-12 tex/manuscript.tex
added some more figures and some text

@@
 
-TPS was developed as a reaction-coordinate free, statistical method 
-which samples trajectories that connect the stable reactant and product states 
-guaranteeing the occurence of the rare barrier crossing events. These trajectories 
-are sampled using a Monte carlo procedure leading to an ensemble of transition
-pathways that are real dynamical trajectories which contain information regarding the 
-mechanism and kinetics of the reaction.
-
 TPS in particular is an attractive method since it samples reactive trajectories
@@
 it does not require a priori knowledge of the reaction coordinate which can be 
-a complicated entity in large biological systems. TPS has been successfully 
+complex for large biological systems.
+TPS can be considered as a generalization of the Monte Carlo sampling of the 
+configuration space to the space of trajectories where the collected ensemble of 
+reactive pathways are real dynamical trajectories compared to articial trajectories
+generated using bias forces. The TPS ensemble can be used to obtain kinetic information
+such as the rate of the reaction as well as the transition state, reaction
+coordinate and the reaction mechanism. TPS has been successfully 
 applied to study problems such as enzyme catalysis, protein folding, etc.
- In this Monte-Carlo procedure in the trajecotry space
-rejected moves correspond to non-reactive trajectories. But these trajectories are 
-required to obtain the equilibrium distribution function.
 
-The TPS ensemble consists of reactive trajectories which connect a reactant state to
-a product state by going through the transition state and the probabilities of configurations thus sampled 
-do not replicate the equilibrium probabilities. To calculate equilibrium properties such as the free energy 
+Since TPS focuses on collecting only reactive trajectories the information contained 
+in the non-reactive trajectories which is essential for describing the 
+equilibrium distribution is missing. 
+Methods that can be used to simultaneously calculate
+the dynamical as well as equilibrium properties of a system within 
+a certain set of asssumptions have garnered interest in recent times. \cite{Suarez14JChemTheoryComput10p2658}
+Transition interface sampling (TIS) is a variant of TPS which can be used for 
+this purpose. \cite{vanErp03JChemPhys118p7762}
+
+A common theme among such methods is to divide the reaction coordinate into
+windows within which independent sampling can be performed and interpolated
+in a way that can used to calculate equilibrium properties
+
+To calculate equilibrium properties such as the free energy 
 of a reaction, equilibrium sampling of the configurations along a reaction coordinate is 
@@
 \subsection{QMMM simulations}
-%------------------------------------------------------------------------------ We performed computational studies on two enzymes where we have shown
-that fast dynamics is part of the reaction mechanism and calculated the electric field near
-the bond-breaking event.
+%------------------------------------------------------------------------------ 
 QMMM simulations were used for the TPS simulations in this work and the 
@@
 Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
-reaction between Adenosine triphosphaste and L-methionine which leads to the 
+reaction between adenosine triphosphaste and L-methionine which leads to the 
 formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
@@
 \end{figure}
+We used transition path sampling to generate and
+analyze the ensemble of reactive trajectories. As order
+parameters that define the reactant and product basins, we
+used the donor-hydride and acceptor-hydride distances. The
+reactant basin was defined as a donor-hydride distance smaller
+than 1.6  and the product basin as an acceptor-hydride
+distance smaller than 1.6 . To seed the generation of the path
+ensemble, we generated a biased trajectory by applying a
+harmonic force between the hydride and the acceptor and
+allowed the system to propagate for 250 fs. New trajectories
+were generated by the shooting algorithm: we perturbed the
+momenta of a randomly chosen slice of a reactive trajectory,
+and using the new momenta, we allowed the system to
+propagate forward and backward in time for 250 fs. A total of
+120 reactive trajectories were generated.
 %------------------------------------------------------------------------------ 
@@
 \centering
-\includegraphics[scale=0.9]{./figures/ada-fenergy.pdf}
+\includegraphics[scale=0.6]{./figures/ada-fenergy.pdf}
 \end{figure}
@@
 %\end{figure}
+%------------------------------------------------------------------------------ 
+\section{Conclusions}
+%------------------------------------------------------------------------------ 
+We have discussed the application of a scheme to calculate the 
+free energies of enzymatic reactions within the TPS method using the MAT2A
+and the ADA enzymes as examples.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-04-13 tex/manuscript.tex
added some more text

@@
 complex for large biological systems.
-TPS can be considered as a generalization of the Monte Carlo sampling of the 
-configuration space to the space of trajectories where the collected ensemble of 
-reactive pathways are real dynamical trajectories compared to articial trajectories
-generated using bias forces. The TPS ensemble can be used to obtain kinetic information
-such as the rate of the reaction as well as the transition state, reaction
-coordinate and the reaction mechanism. TPS has been successfully 
-applied to study problems such as enzyme catalysis, protein folding, etc.
+TPS is based on the Monte Carlo importance sampling in the 
+space of trajectories where the collected ensemble of 
+reactive pathways are real dynamical trajectories generated without the 
+application of any bias forces. Thus the TPS ensemble can be used to obtain 
+kinetic information such as the rate of the reaction as well as to elucidate 
+the transition state, reaction coordinate and the reaction mechanism. 
+TPS has been successfully applied to study problems such as enzyme catalysis, 
+protein folding, etc.
 
-Since TPS focuses on collecting only reactive trajectories the information contained 
-in the non-reactive trajectories which is essential for describing the 
-equilibrium distribution is missing. 
+Calculating equilibrium properties such as the free energy of an enzymatic 
+catalysis reaction within TPS needs more attention. 
+Since the TPS path ensemble only consists of reactive trajectories, the information from 
+the discarded non-reactive trajectories which is required for obtaining the 
+equilibrium probability distribution of the order parameter is missing. \cite{Dellago09AdvCompSimAppp167}
+
 Methods that can be used to simultaneously calculate
@@
 this purpose. \cite{vanErp03JChemPhys118p7762}
-
-A common theme among such methods is to divide the reaction coordinate into
-windows within which independent sampling can be performed and interpolated
-in a way that can used to calculate equilibrium properties
-
-To calculate equilibrium properties such as the free energy 
-of a reaction, equilibrium sampling of the configurations along a reaction coordinate is 
-required which is not obtained with straightforward TPS. \cite{Dellago09AdvCompSimAppp167}
 
@@
 %------------------------------------------------------------------------------ 
-\section{Results and discussion}
+\section{Computational details}
 %------------------------------------------------------------------------------ 
-\subsection{QMMM simulations}
+\subsection{System preparation}
 %------------------------------------------------------------------------------
-QMMM simulations were used for the TPS simulations in this work and the 
-CHARMM force field \cite{Brooks09JComputChem30p1545} is used for the 
-MM portion and the PM3 semiempirical method \cite{Repasky02JComputChem23p1601} 
-is used for the QM region.
+We used the crystal structure of human MAT2A 
+bound to methylthioadenosine and di-imido triphosphate (PNPNP)
+(7L1A from the protein data bank (PDB)) reported by Niland 
+et al. \cite{Niland21Biochem60p791} and the crystal structure of \textit{pv}ADA 
+in complex with MT-coformycin (3EWC from the PDB) reported by Ho et 
+al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
+simulations. Molecular dynamics simulations within classical and hybrid 
+quantum mechanics/molecular mechanics (QMMM) approximations were performed 
+using the CHARMM program package and using the CHARMM 36 force 
+field \cite{Brooks09JComputChem30p1545} for the MM part and the 
+approximate PM3 semiempirical method \cite{Repasky02JComputChem23p1601} for the QM part.
 
-%------------------------------------------------------------------------------ 
-\subsection{Human Methionine Adenosyltransferase 2A}
-%------------------------------------------------------------------------------ 
-Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
-reaction between adenosine triphosphaste and L-methionine which leads to the 
-formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
-The mechanism of this reaction is determined to follow a SN$_2$ transition state
-with the sulphur atom of the L-methionine as the nucleophile with the 
-triphosphate leaving group.
+Each protein was solvated within
+a 15 $\AA$ sphere using TIP3 waters and neutralized using
+potassium ions. We minimized the system with 50 steps of the
+Steepest Descent method, followed by 2000 steps of the
+adopted Basis Newton-Raphson method. The system was
+heated slowly to 300 K for 30 ps, starting with harmonic
+constraints except on H atoms and TIP3 waters that were
+gradually reduced to force constant 1. We started the
+equilibration with 15 ps with harmonic constraints of force
+constant 1, which were then removed and the equilibration ran
+for 60 ps.
 
-\begin{figure}[ht!]
-\includegraphics[scale=0.25]{figures/mat2a-transition-state.png}
-\end{figure}
-
-\begin{figure}[ht!]
-\includegraphics[scale=0.5]{figures/comm-60-mat2a.pdf}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
-\caption{The free energy as a function of the order parameter.}
-\end{figure}
 We used transition path sampling to generate and
@@
 120 reactive trajectories were generated.
+%------------------------------------------------------------------------------ 
+\subsection{Plasmodium vivax adenosine deaminase}
+%------------------------------------------------------------------------------
+
+%------------------------------------------------------------------------------ 
+\section{Results and discussion}
+%------------------------------------------------------------------------------ 
+\subsection{Human Methionine Adenosyltransferase 2A}
+%------------------------------------------------------------------------------ 
+Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
+reaction between adenosine triphosphaste and L-methionine which leads to the 
+formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
+The mechanism of this reaction is experimentally determined to follow a 
+SN$_2$ transition state with the sulphur atom of the L-methionine as the 
+nucleophile with the triphosphate leaving group.
+
+\begin{figure}[ht!]
+\includegraphics[scale=0.25]{figures/mat2a-transition-state.png}
+\end{figure}
+
+\begin{figure}[ht!]
+\includegraphics[scale=0.5]{figures/comm-60-mat2a.pdf}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
+\caption{The free energy as a function of the order parameter.}
+\end{figure}
+
 %------------------------------------------------------------------------------ 

### 2022-04-14 tex/manuscript.tex
added a lot more text now

@@
 TPS is based on the Monte Carlo importance sampling in the 
-space of trajectories where the collected ensemble of 
-reactive pathways are real dynamical trajectories generated without the 
-application of any bias forces. Thus the TPS ensemble can be used to obtain 
+space of trajectories. It uses the shooting and shifting algorithm 
+to collect an ensemble of reactive pathways that are real dynamical 
+trajectories generated without the application of any bias forces. \cite{dellago02AdvChemPhys123} 
+Therefore the TPS ensemble can be used to obtain 
 kinetic information such as the rate of the reaction as well as to elucidate 
@@
 TPS has been successfully applied to study problems such as enzyme catalysis, 
-protein folding, etc.
+\cite{Antoniou01JPhysChemB105p5553,Hay12NatChem4p161,Schwartz09NatChemBiol5p551} 
+protein folding, \cite{Bolhuis03ProcNatlAcadSci100p12129,Juraszek12ChemPhys396p30} etc.
 
 Calculating equilibrium properties such as the free energy of an enzymatic 
-catalysis reaction within TPS needs more attention. 
+catalysis reaction within TPS is not straighforward. 
+The Monte Carlo algorithm used in TPS has an acceptance criterion that 
+is only satisfied if the trajectory begins in the reactant basin and ends in 
+the product basin or vice versa, hence trajectories that are localized in the 
+reactant or product basins are rejected. The localized trajectories contribute
+to the probability density distribution function ($P(\lambda)$) which is the 
+probability to find the system at a particular value of the order parameter $\lambda$.
+
 Since the TPS path ensemble only consists of reactive trajectories, the information from 
 the discarded non-reactive trajectories which is required for obtaining the 
-equilibrium probability distribution of the order parameter is missing. \cite{Dellago09AdvCompSimAppp167}
-
-Methods that can be used to simultaneously calculate
-the dynamical as well as equilibrium properties of a system within 
-a certain set of asssumptions have garnered interest in recent times. \cite{Suarez14JChemTheoryComput10p2658}
-Transition interface sampling (TIS) is a variant of TPS which can be used for 
-this purpose. \cite{vanErp03JChemPhys118p7762}
-
-TPS based free energy simulations have been attempted previously by 
+equilibrium probability distribution of the order parameters is 
+missing. \cite{Dellago09AdvCompSimAppp167} Radhakrishnan et al. developed 
+`BOLAS' which combines the shooting and shifting algorithms of 
+TPS with a window based sampling technique 
+in the spirit of umbrella sampling to calculate equilibrium free energy 
+differences. \cite{Radhakrishnan04JChemPhys121p2436} 
+Baron peters developed a variant of this scheme called the 
+equilibrium path sampling (EPS) where they use aimless shooting 
+to create the ensemble within each window. \cite{Peters08JAmChemSoc130p17342,Silveira21JPhysChemB125p2018} 
+Apart from these schemes a recent work by 
 Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111,Dellago04MolSimul30p795}
+has attempted to look at the TPS based free energy calculations for 
+reactions that are diffusive in nature where 
+Transition interface sampling (TIS) \cite{vanErp03JChemPhys118p7762}
+could work well.
 
@@
 be drawn. 
-
-One example of the TPS based free energies calculations on real 
-protein systems \cite{Silveira21JPhysChemB125p2018} Human Mat2a enzyme and 
-plasmodium vivax adenosine deaminase 
 %------------------------------------------------------------------------------ 
@@
 momenta ($\textbf{p}$) of each atom at regular intervals of time $t_i = t_0 + i\Delta t$  
-($i = 0,1,2,\ldots$). At each of the time slices, the positions and momenta can be collectively represented as
-the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is run for a total 
-time of $\mathcal{T}$ the number of time slices is given by $L = \mathcal{T}/\Delta t +1$ and for this sequence of 
+($i = 0,1,2,\ldots$). At each of the time slices, the positions and 
+momenta can be collectively represented as
+the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is 
+run for a total 
+time of $\mathcal{T}$ the number of time slices is given by 
+$L = \mathcal{T}/\Delta t +1$ and for this sequence of 
 times the state of the system can be represented as 
@@
 %------------------------------------------------------------------------------
-We used the crystal structure of human MAT2A 
+We used the crystal structure of the human MAT2A enzyme 
 bound to methylthioadenosine and di-imido triphosphate (PNPNP)
-(7L1A from the protein data bank (PDB)) reported by Niland 
+(7L1A from the protein data bank) reported by Niland 
 et al. \cite{Niland21Biochem60p791} and the crystal structure of \textit{pv}ADA 
-in complex with MT-coformycin (3EWC from the PDB) reported by Ho et 
+in complex with MT-coformycin (3EWC from the protein data bank) reported by Ho et 
 al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
-simulations. Molecular dynamics simulations within classical and hybrid 
-quantum mechanics/molecular mechanics (QMMM) approximations were performed 
-using the CHARMM program package and using the CHARMM 36 force 
-field \cite{Brooks09JComputChem30p1545} for the MM part and the 
-approximate PM3 semiempirical method \cite{Repasky02JComputChem23p1601} for the QM part.
+simulations. For the human MAT2A enzyme the substrates were modified to 
+S-Adenosyl methionine (SAM) and adenosyl triphosphate (ATP) whereas for the \textit{pv}ADA
+enzyme the substrates were modified to $2^{'}-$deoxyadenosine and hydroxide ion.
+
+MD simulations within classical and hybrid 
+quantum mechanics/molecular mechanics (QMMM) approximations were carried out 
+using the CHARMM program package. \cite{Brooks83JComputChem4p187,Brooks09JComputChem30p1545} 
+For the MM calculations the CHARMM36 force field \cite{Brooks09JComputChem30p1545} 
+was used and for the QM calculations the approximate PM3 semiempirical 
+method \cite{Repasky02JComputChem23p1601} was used. Both the
+systems were solvated with water molecules using the TIP3P model
+in a nanodroplet, whose volume was set to be $15\;\AA$ away from
+the surface of the protein, and then neutralized with potassium
+ions.
 
@@
 120 reactive trajectories were generated.
-%------------------------------------------------------------------------------ 
-\subsection{Plasmodium vivax adenosine deaminase}
-%------------------------------------------------------------------------------
-
 %------------------------------------------------------------------------------ 

### 2022-04-14 tex/manuscript.tex
added some more text

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jpcbfk,manuscript=article,layout=traditional]{achemso}
+\documentclass[journal=jpcbfk,manuscript=article,layout=twocolumn]{achemso}
 
@@
 \title[]
-  {Transition path sampling based calculation of free energies of enzymatic
-  reactions}
+  {Transition path sampling based calculations of free energies of enzymatic
+  reactions: the case of human MAT2A and plasmodium vivax adenosine deaminase}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 to the probability density distribution function ($P(\lambda)$) which is the 
-probability to find the system at a particular value of the order parameter $\lambda$.
+probability of a system at equilibrium to have a particular value of the 
+order parameter $\lambda$. \cite{Dellago09AdvCompSimAppp167}
 
-Since the TPS path ensemble only consists of reactive trajectories, the information from 
-the discarded non-reactive trajectories which is required for obtaining the 
-equilibrium probability distribution of the order parameters is 
-missing. \cite{Dellago09AdvCompSimAppp167} Radhakrishnan et al. developed 
-`BOLAS' which combines the shooting and shifting algorithms of 
-TPS with a window based sampling technique 
+Frenkel has shown that inclusion of rejected states in Markov state Monte 
+Carlo algorith can help speed up the convergence of the calculations of 
+order-parameter distribution of a two-dimensional
+Ising model. \cite{Frenkel04ProcNatAcadSci101p17571}
+Inclusion of rejected trajectories within TPS was first discussed by Radhakrishnan 
+et al. \cite{Radhakrishnan04JChemPhys121p2436} who developed `BOLAS' which combines 
+the shooting and shifting algorithms of TPS with a window based sampling technique 
 in the spirit of umbrella sampling to calculate equilibrium free energy 
-differences. \cite{Radhakrishnan04JChemPhys121p2436} 
-Baron peters developed a variant of this scheme called the 
+differences. Peters et al. used a variant of the BOLAS scheme called the 
 equilibrium path sampling (EPS) where they use aimless shooting 
-to create the ensemble within each window. \cite{Peters08JAmChemSoc130p17342,Silveira21JPhysChemB125p2018} 
-Apart from these schemes a recent work by 
-Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111,Dellago04MolSimul30p795}
-has attempted to look at the TPS based free energy calculations for 
-reactions that are diffusive in nature where 
-Transition interface sampling (TIS) \cite{vanErp03JChemPhys118p7762}
-could work well.
+combined with window based sampling for free energy calculations. \cite{Peters08JAmChemSoc130p17342,Beckham10epsbook} 
+More recently Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
+have developed a reweighted path ensemble scheme within the standard TPS method 
+called virtual interface exchange TPS (VIE-TPS) to calculate the free energy 
+landscape. VIE-TPS is shown to produce reasonable free energies for 
+diffusive reactions whereas it is less accurate otherwise.
 
-In this manuscript we will focus on the implementation and application of the 
-free energy based calculations within the TPS method on reactions catalyzed by two systems:
+Free energy calculations based on the TPS shooting algorithm is 
+appealing but not a lot of examples exist in the literature 
+which compare the computational calculations with experimental
+results for enzymatic reactions.  
+In this manuscript we implement and apply the 
+algorithm developed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
+to calculate the free energies of reactions catalyzed by two systems:
 the human MAT2A enzyme and the plasmodium vivax adenosine deaminase 
 enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
-These enzymes were chosen not only because of their importance in biochemical pathways and
-but also because of the availability of experimental results with which comparisons can 
-be drawn. 
+These enzymes were chosen not only because of their importance in various 
+biochemical applications but also because of the availability of experimental 
+results with which comparisons can be drawn. The manuscript is organized as 
+follows: first we discuss the TPS method and the algorithm that we use 
+to calculate the free energies briefly, then we show calculations which apply this 
+method on the two enzymatic reactions. Finally we compare our results to expeiments
+and provide conclusions. 
 %------------------------------------------------------------------------------ 
@@
 bound to methylthioadenosine and di-imido triphosphate (PNPNP)
-(7L1A from the protein data bank) reported by Niland 
+(protein data bank ID: 7L1A) reported by Niland 
 et al. \cite{Niland21Biochem60p791} and the crystal structure of \textit{pv}ADA 
-in complex with MT-coformycin (3EWC from the protein data bank) reported by Ho et 
+in complex with MT-coformycin (protein data bank ID: 3EWC) reported by Ho et 
 al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
-simulations. For the human MAT2A enzyme the substrates were modified to 
-S-Adenosyl methionine (SAM) and adenosyl triphosphate (ATP) whereas for the \textit{pv}ADA
+simulations. For the MAT2A enzyme the substrates were modified to 
+L-methionine and adenosyl triphosphate (ATP) whereas for the \textit{pv}ADA
 enzyme the substrates were modified to $2^{'}-$deoxyadenosine and hydroxide ion. 
-
-MD simulations within classical and hybrid 
-quantum mechanics/molecular mechanics (QMMM) approximations were carried out 
-using the CHARMM program package. \cite{Brooks83JComputChem4p187,Brooks09JComputChem30p1545} 
+MD simulations within classical and hybrid quantum mechanics/molecular mechanics 
+(QMMM) approximations were carried out using the CHARMM program 
+package. \cite{Brooks83JComputChem4p187,Brooks09JComputChem30p1545} 
 For the MM calculations the CHARMM36 force field \cite{Brooks09JComputChem30p1545} 
 was used and for the QM calculations the approximate PM3 semiempirical 
-method \cite{Repasky02JComputChem23p1601} was used. Both the
-systems were solvated with water molecules using the TIP3P model
-in a nanodroplet, whose volume was set to be $15\;\AA$ away from
-the surface of the protein, and then neutralized with potassium
-ions.
-
-Each protein was solvated within
-a 15 $\AA$ sphere using TIP3 waters and neutralized using
-potassium ions. We minimized the system with 50 steps of the
-Steepest Descent method, followed by 2000 steps of the
-adopted Basis Newton-Raphson method. The system was
-heated slowly to 300 K for 30 ps, starting with harmonic
-constraints except on H atoms and TIP3 waters that were
-gradually reduced to force constant 1. We started the
-equilibration with 15 ps with harmonic constraints of force
-constant 1, which were then removed and the equilibration ran
-for 60 ps.
-
-We used transition path sampling to generate and
-analyze the ensemble of reactive trajectories. As order
-parameters that define the reactant and product basins, we
+method \cite{Repasky02JComputChem23p1601} was used. 
+We solvated the two proteins using TIP3P water molecules 
+in a nanodroplet sphere whose volume was fixed at $15$ {\AA}
+away from the protein's surface and the total charge was neutralized 
+using potassium ions. Both the systems were minimized with 50 steps of the
+steepest descent method, followed by 2000 steps of the
+adopted basis Newton-Raphson method. The minimized systems were
+heated slowly to 300 K for 35 ps beginning with harmonic
+constraints on all atoms except on the H atoms and the TIP3 
+waters with gradual reduction of the restraint forces to 1. 
+15 ps of equilibration was carried out starting with harmonic 
+constraints of force constant 1 followed by 20 ps of contraint free 
+equilibration to prepare the systems for production runs.  
+%------------------------------------------------------------------------------ 
+\subsection{Transition path sampling}
+%------------------------------------------------------------------------------
+We used TPS to generate an ensemble of reactive trajectories. 
+The time step that we used for all our calculations was $1\;fs$ with a total
+of 250 steps forward and 250 steps backward which amounts to a total of $0.5\;ps$
+trajectory.  
+\subsubsection{human MAT2A}
+As order parameters that define the reactant and product basins, we
 used the donor-hydride and acceptor-hydride distances. The
@@
 120 reactive trajectories were generated.
+
+\subsubsection{plasmodium vivax ADA}
+%------------------------------------------------------------------------------ 
+\subsection{Free energy calculations}
+%------------------------------------------------------------------------------ 
+For the free energy calculations we divided the reaction coordinate into 13 windows for 
+MAT2A and 20 windows for pvADA with a $15\%$ overlap on each side of the window.
+
 %------------------------------------------------------------------------------ 
@@
 \begin{figure}[ht!]
-\centering
-\includegraphics[scale=0.22]{./figures/ada-protein.png}
+\includegraphics[scale=0.15]{./figures/ada-protein.png}
 \end{figure}
@@
 \begin{figure}[ht!]
-\centering
-\includegraphics[scale=0.6]{./figures/ada-fenergy.pdf}
+\includegraphics[scale=0.5]{./figures/ada-fenergy.pdf}
 \end{figure}

### 2022-04-14 tex/manuscript.tex
added some more text

@@
   Transition path sampling (TPS) has been widely applied for the
-  calculations of transition state and reaction coordinates of 
-  condensed phase systems. Here, we discuss a scheme for the calculation
-  of free energies with the TPS method 
-  We discuss a scheme for the calculation of free energies within 
-  the transition path sampling (TPS) method. TPS is often used for
-  the calculations of reaction coordinates of complex reactions in 
-  the condensed phase and in this work we use it to sample the 
-  equilibrium distribution of the reaction coordinate. This 
+  calculations of reaction rate, transition state and reaction 
+  coordinates of condensed phase systems. Here, we discuss a scheme 
+  for the calculation of free energies with the TPS method in combination 
+  with a window based sampling technique. This 
   distribution can then be used for the calculations of free energies
@@
 \subsubsection{human MAT2A}
-As order parameters that define the reactant and product basins, we
-used the donor-hydride and acceptor-hydride distances. The
-reactant basin was defined as a donor-hydride distance smaller
-than 1.6  and the product basin as an acceptor-hydride
-distance smaller than 1.6 . To seed the generation of the path
-ensemble, we generated a biased trajectory by applying a
-harmonic force between the hydride and the acceptor and
-allowed the system to propagate for 250 fs. New trajectories
-were generated by the shooting algorithm: we perturbed the
-momenta of a randomly chosen slice of a reactive trajectory,
-and using the new momenta, we allowed the system to
-propagate forward and backward in time for 250 fs. A total of
-120 reactive trajectories were generated.
+The key bond parameters for the reaction in this system consists of the 
+oxygen atom on the ATP, the carbon atom on the adenosyl group and the 
+sulphur atom on the L-methionine. The reactant basin is defined by the 
+oxygen-carbon bond distance being less than 1.7 {\AA} and the product basin 
+is characterized by the suphur-carbon bond distance being less than 
+1.7 {\AA}. The initial reactive trajectory was generated by the application 
+of a bias harmonic force between the oxygen and the carbon atoms 
+and running dynamics for $250\;fs$. The shooting algorithm was used to generate
+new trajectories within which first a random time slice is chosen from the reactive 
+trajectory and perturbing the momenta of the system. Then with the new momenta
+the system is propagated forward and backward in time for $250\;fs$.
+A total of 120 reactive trajectories were generated.
 
 \subsubsection{plasmodium vivax ADA}
+The $pv$ADA reaction is 
 %------------------------------------------------------------------------------ 

### 2022-04-20 tex/manuscript.tex
added some more text and figures

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jpcbfk,manuscript=article,layout=twocolumn]{achemso}
+\documentclass[journal=jpcbfk,manuscript=article,layout=traditional]{achemso}
 
@@
 equilibrium path sampling (EPS) where they use aimless shooting 
-combined with window based sampling for free energy calculations. \cite{Peters08JAmChemSoc130p17342,Beckham10epsbook} 
+combined with window based sampling for free energy calculations. 
+\cite{Peters08JAmChemSoc130p17342,Beckham10epsbook} 
 More recently Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
@@
 al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
-simulations. For the MAT2A enzyme the substrates were modified to 
-L-methionine and adenosyl triphosphate (ATP) whereas for the \textit{pv}ADA
+simulations. For calculations involving the MAT2A enzyme the substrates were modified to 
+L-methionine and adenosyl triphosphate (ATP) whereas for calculations with the \textit{pv}ADA
 enzyme the substrates were modified to $2^{'}-$deoxyadenosine and hydroxide ion. 
@@
 away from the protein's surface and the total charge was neutralized 
-using potassium ions. Both the systems were minimized with 50 steps of the
+using potassium ions. 
+Both the systems were minimized with 50 steps of the
 steepest descent method, followed by 2000 steps of the
@@
 constraints of force constant 1 followed by 20 ps of contraint free 
-equilibration to prepare the systems for production runs.  
+equilibration to prepare the systems for TPS simulations.  
 %------------------------------------------------------------------------------ 
@@
 %------------------------------------------------------------------------------
-We used TPS to generate an ensemble of reactive trajectories. 
-The time step that we used for all our calculations was $1\;fs$ with a total
-of 250 steps forward and 250 steps backward which amounts to a total of $0.5\;ps$
-trajectory.   
-\subsubsection{human MAT2A}
-The key bond parameters for the reaction in this system consists of the 
-oxygen atom on the ATP, the carbon atom on the adenosyl group and the 
-sulphur atom on the L-methionine. The reactant basin is defined by the 
-oxygen-carbon bond distance being less than 1.7 {\AA} and the product basin 
-is characterized by the suphur-carbon bond distance being less than 
-1.7 {\AA}. The initial reactive trajectory was generated by the application 
-of a bias harmonic force between the oxygen and the carbon atoms 
-and running dynamics for $250\;fs$. The shooting algorithm was used to generate
+TPS calculations collect reactive trajectories based on the Monte Carlo algorithm 
+and requires a proper definition of the reactant and product states.
+
+The reaction catalyzed by the human MAT2A enzyme is shown in Fig. \ref{fig:mat2a-reaction}
+between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules resulting in the formation of
+S-adenosyl-L-methionine (SAM). SAM is an essential metabolite which is distributed to almost all 
+body tissues and fluids, is an universal methyl donor and is of
+fundamental importance to metabolism of compounds such as hormones, neureotransmitters, proteins, 
+and nucleic acids. \cite{Friedel89Drugs38p389} The reaction was experimentally characterized by 
+Firestone et al. \cite{Firestone17JAmChemSoc139p13754} to follow a S$_{\text{N}}2$ transition 
+state. The key bond parameters for this reaction consists of the distance between the
+nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP ($S-C$), and the distance between 
+oxygen atom of the phosphate group and the $5'-C$ atom of ATP ($O-C$).
+For the TPS calculations we define the reactant state to have $O-C$ distance to be 
+less than 1.7 {\AA} and the product state to have the $S-C$ distance being less than 1.7 {\AA}.
+
+\begin{figure}
+\includegraphics[scale=0.6]{figures/mat2a-reaction.png}
+\caption{The reaction catalyzed by human MAT2A enzyme}
+\label{fig:mat2a-reaction}
+\end{figure}
+
+The reaction catalyzed by $pv$ADA is shown in  
+Fig. \ref{fig:ada-reaction} which is between the adenosine molecule and a hydroxide ion
+which results in the formation of inosine with ammonia as a by product. 
+$pv$ADA is an essential enzyme in the purine salvage pathway for plasmodium vivax and 
+inhibiting this enzyme is a promising approach to treat malaria. \cite{Madrid08JBiolChem283p35899}  
+This reaction proceeds with the nucleophilic O atom (O$^{\text{hydroxyl}}$) of the hydroxide anion
+attacking the electrophilic C6 atom of the adenosyl ring leading to the formation of a Meisenheimer 
+complex while the N1 atom on the adenosyl ring starts to get 
+protonated by the Glu229 residue of $pv$ADA. At the transition state, experimental findings 
+suggest that the dissociation of the C6$-$N6 bond as well as the protonation of the N1 
+atom are almost complete. \cite{Luo07JAmChemSoc129p8008} The key bond parameters are 
+the distance between the approaching hydroxyl oxygen and the C6 atom ($O-C$) and the 
+dissociating C6$-$N6 ($N-C$) bond distance. For the TPS simulaions we defined the reactant state to 
+have the $O-C$ distance being less than $1.7$ {\AA} and the product state to have the $O-C$ 
+distance being less than $1.7$ {\AA}.
+
+\begin{figure}
+\includegraphics[scale=0.6]{figures/ada-new.png}
+\caption{The reaction catalyzed by adenosine deaminase enzyme}
+\label{fig:ada-reaction}
+\end{figure}
+
+A biased initial reactive trajectory is obtained by using harmonic restraint forces 
+of 1500 to break the $O-C$ bond and form the $S-C$ bond for the MAT2A enzyme and 
+to break the $N-C$ bond and form the $O-C$ bond for the case of the $pv$ADA enzyme. 
+Using the biased trajectory as the starting point and after removing the restraint 
+forces, the shooting algorithm was used to generate
 new trajectories within which first a random time slice is chosen from the reactive 
-trajectory and perturbing the momenta of the system. Then with the new momenta
-the system is propagated forward and backward in time for $250\;fs$.
-A total of 120 reactive trajectories were generated.
+trajectory and perturbing the momenta of the system. 
+The time step that we used for all our calculations was $1\;fs$ and we propagate the 
+dynamics for a total of 250 steps forward and 250 steps backward which amounts to a 
+total of $0.5\;ps$. 
+A total of 150 and 100 reactive trajectories were generated for the MAT2A and $pv$ADA
+enzymes, respectively.
 
-\subsubsection{plasmodium vivax ADA}
-The $pv$ADA reaction is 
 %------------------------------------------------------------------------------ 
@@
 %------------------------------------------------------------------------------ 
+Window based free energy calculations such as WHAM divide the reaction coordinate
+into windows and calculate the individual populations within the windows and 
+decouple the applied bias with some formulas. 
 For the free energy calculations we divided the reaction coordinate into 13 windows for 

### 2022-04-21 tex/manuscript.tex
added some text
added some system preparation text

@@
 \usepackage{mathptmx}
-\usepackage[scaled=0.92]{helvet}
+%\usepackage[scaled=0.92]{helvet}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   {Transition path sampling based calculations of free energies of enzymatic
-  reactions: the case of human MAT2A and plasmodium vivax adenosine deaminase}
+  reactions: the case of human methionine adenosyl transferase and plasmodium 
+  vivax adenosine deaminase}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %------------------------------------------------------------------------------ 
-\subsection{The algorithm}
+\subsection{Algorithm for free energy calculations}\label{ssec:algorithm}
 %------------------------------------------------------------------------------ 
@@
 enzyme the substrates were modified to $2^{'}-$deoxyadenosine and hydroxide ion. 
+We solvated both the systems using TIP3P water molecules 
+in a nanodroplet sphere whose volume was fixed at $15$ {\AA}
+away from the protein's surface and the total charge was neutralized 
+using potassium ions.
+
 MD simulations within classical and hybrid quantum mechanics/molecular mechanics 
@@
 method \cite{Repasky02JComputChem23p1601} was used. 
-We solvated the two proteins using TIP3P water molecules 
-in a nanodroplet sphere whose volume was fixed at $15$ {\AA}
-away from the protein's surface and the total charge was neutralized 
-using potassium ions. 
-Both the systems were minimized with 50 steps of the
+For the MAT2A system the partitioned QM region consisted of 
+the ATP and MET molecules along with three Mg$^{2+}$ ions 
+whereas for the $pv$ADA system it consisted of 
+the adenosine molecule, an hydroxide anion, Zn$^{2+}$ ion and 
+a portion of the Glu229 residue which was partitioned within the generalized 
+hybrid orbital (GHO) scheme. \cite{Gao98JPhysChemA102p4714}
+
+Both the systems were minimized within classical molecular mechanics 
+with 50 steps of the
 steepest descent method, followed by 2000 steps of the
-adopted basis Newton-Raphson method. The minimized systems were
+adopted basis Newton-Raphson method. 
+The minimized systems were
 heated slowly to 300 K for 35 ps beginning with harmonic
@@
 equilibration to prepare the systems for TPS simulations. 
+During the heating and equilibration steps, bonds that contain 
+hydrogen atoms in the MM region were restrained to their equilibrium values 
+using the SHAKE procedure. \cite{Ryckaert77JComputPhys23p327}
+
 %------------------------------------------------------------------------------ 
@@
 %------------------------------------------------------------------------------
-TPS calculations collect reactive trajectories based on the Monte Carlo algorithm 
-and requires a proper definition of the reactant and product states.
+TPS simulations require the definition of unique reactant and product states
+to collect reactive trajectories based on the Monte Carlo algorithm. Here we 
+discuss the details of the reactions involving the MAT2A and the $pv$ADA enzymes 
+to identify and define the parameters characterizing the reactant and product basins.
 
@@
 state. The key bond parameters for this reaction consists of the distance between the
-nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP ($S-C$), and the distance between 
-oxygen atom of the phosphate group and the $5'-C$ atom of ATP ($O-C$).
-For the TPS calculations we define the reactant state to have $O-C$ distance to be 
-less than 1.7 {\AA} and the product state to have the $S-C$ distance being less than 1.7 {\AA}.
+nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as $dis(SC)$), and the distance between 
+oxygen atom of the phosphate group and the $5'-C$ atom of ATP ($dis(OC)$).
+For the TPS calculations we define the reactant state to have $dis(OC)$ to be 
+less than 1.7 {\AA} and the product state to have the $dis(SC)$ less than 1.7 {\AA}.
 
@@
 atom are almost complete. \cite{Luo07JAmChemSoc129p8008} The key bond parameters are 
-the distance between the approaching hydroxyl oxygen and the C6 atom ($O-C$) and the 
-dissociating C6$-$N6 ($N-C$) bond distance. For the TPS simulaions we defined the reactant state to 
-have the $O-C$ distance being less than $1.7$ {\AA} and the product state to have the $O-C$ 
-distance being less than $1.7$ {\AA}.
+the distance between the approaching hydroxyl oxygen and the C6 atom (denoted as $dis(OC)$) and the 
+dissociating C6$-$N6 ($dis(NC)$) bond distance. For the TPS simulaions we define the reactant state to 
+have $dis(NC)$ being less than $1.7$ {\AA} and the product state to have the 
+$dis(OC)$ less than $1.7$ {\AA}.
 
@@
 %------------------------------------------------------------------------------ 
-Window based free energy calculations such as WHAM divide the reaction coordinate
-into windows and calculate the individual populations within the windows and 
-decouple the applied bias with some formulas. 
-For the free energy calculations we divided the reaction coordinate into 13 windows for 
-MAT2A and 20 windows for pvADA with a $15\%$ overlap on each side of the window.
-
+To obtain the free energy profile along the order parameter, we use the method 
+discussed in sec. \ref{ssec:algorithm}. The order paratemer for the MAT2A system 
+is defined to be $dis(OC)-dis(SC)$ and that for the $pv$ADA is defined as 
+$dis(NC)-dis(OC)$. For the MAT2A system, configurations are sampled within 
+the range $[-2.0,2.0]$ and this range is divided into 14 overlapping windows 
+($0.09$ {\AA} overlap on either side). For the $pv$ADA system configurations are
+sampled along the order parameter within the range $[-2.0, 2.0]$ where the range is
+divided into 20 overlapping windows ($0.09$ {\AA} overlap on either side of each window). 
+A reactive trajectory from the TPS ensemble is chosen to be the guide trajectory 
+for sampling within each of the windows. To sample in a particular window $w$, a 
+frame from this reactive trajectory is chosen such that the value of the 
+order parameter calculated for this frame falls within $w$. Using the shooting algorithm
+the momenta are perturbed and dynamics is run for 10 steps forwards and backwards with a 
+$1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.    
+Within each window 3000 TPS trajectories are collected using the acceptance criteria 
+discussed in sec. \ref{ssec:algorithm}. The equilibrium probability distribution function 
+for the order parameter within window $w$ is then calculated as normalized histogram
+similar to the WHAM procedure. \cite{Kumar92JComputChem13p1011}
 %------------------------------------------------------------------------------ 

### 2022-04-21 tex/manuscript.tex
added some more text

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\abbreviations{TPS,NMR,UV}
+\abbreviations{TPS,WHAM}
 \keywords{American Chemical Society, \LaTeX}
@@
 enzyme the substrates were modified to $2^{'}-$deoxyadenosine and hydroxide ion. 
-We solvated both the systems using TIP3P water molecules 
+We solvated both the systems using TIP3P water molecules \cite{Jorgensen83JChemPhys79p926}
 in a nanodroplet sphere whose volume was fixed at $15$ {\AA}
@@
 (QMMM) approximations were carried out using the CHARMM program 
-package. \cite{Brooks83JComputChem4p187,Brooks09JComputChem30p1545} 
+package. \cite{Brooks83JComputChem4p187} 
 For the MM calculations the CHARMM36 force field \cite{Brooks09JComputChem30p1545} 
@@
 method \cite{Repasky02JComputChem23p1601} was used. 
-For the MAT2A system the partitioned QM region consisted of 
-the ATP and MET molecules along with three Mg$^{2+}$ ions 
-whereas for the $pv$ADA system it consisted of 
+For the hybrid QMMM simulations, the MAT2A system was partitioned into a QM
+region consisting of the ATP and MET molecules along with three Mg$^{2+}$ ions 
+and the MM region consists of the rest of the molecules. 
+For the $pv$ADA system the QM region consisted of 
 the adenosine molecule, an hydroxide anion, Zn$^{2+}$ ion and 
@@
 
-Both the systems were minimized within classical molecular mechanics 
-with 50 steps of the
+The energies of both the systems were minimized using 50 steps of the 
 steepest descent method, followed by 2000 steps of the
-adopted basis Newton-Raphson method. 
+adopted basis Newton-Raphson method where only classical molecular mechanics 
+was used for the dynamics. 
 The minimized systems were
 heated slowly to 300 K for 35 ps beginning with harmonic
-constraints on all atoms except on the H atoms and the TIP3 
-waters with gradual reduction of the restraint forces to 1. 
+constraints on all atoms except on the H atoms and the TIP3P 
+water molecules with gradual reduction of the restraint forces to 1. 
 15 ps of equilibration was carried out starting with harmonic 
@@
 discuss the details of the reactions involving the MAT2A and the $pv$ADA enzymes 
-to identify and define the parameters characterizing the reactant and product basins.
+to identify and define the order parameters characterizing the reactant and product basins.
 
-The reaction catalyzed by the human MAT2A enzyme is shown in Fig. \ref{fig:mat2a-reaction}
+The reaction catalyzed by the human MAT2A enzyme is
 between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules resulting in the formation of
-S-adenosyl-L-methionine (SAM). SAM is an essential metabolite which is distributed to almost all 
+S-adenosyl-L-methionine (SAM) which is depicted in Fig. \ref{fig:mat2a-reaction}.
+SAM is an essential metabolite which is distributed to almost all 
 body tissues and fluids, is an universal methyl donor and is of
-fundamental importance to metabolism of compounds such as hormones, neureotransmitters, proteins, 
-and nucleic acids. \cite{Friedel89Drugs38p389} The reaction was experimentally characterized by 
+fundamental importance to the metabolism of compounds such as hormones, neureotransmitters, proteins, 
+and nucleic acids. \cite{Friedel89Drugs38p389} This reaction was experimentally characterized by 
 Firestone et al. \cite{Firestone17JAmChemSoc139p13754} to follow a S$_{\text{N}}2$ transition 
@@
 the range $[-2.0,2.0]$ and this range is divided into 14 overlapping windows 
-($0.09$ {\AA} overlap on either side). For the $pv$ADA system configurations are
-sampled along the order parameter within the range $[-2.0, 2.0]$ where the range is
-divided into 20 overlapping windows ($0.09$ {\AA} overlap on either side of each window). 
-A reactive trajectory from the TPS ensemble is chosen to be the guide trajectory 
+($0.09$ {\AA} overlap on either side). 
+Configurations are sampled for the $pv$ADA system within the interval $[-3.0, 3.0]$ {\AA}
+of the order parameter value. 
+This interval is divided into 20 overlapping windows ($0.045$ {\AA} between 
+neighboring windows). A reactive trajectory from the TPS ensemble is chosen to be the guide trajectory 
 for sampling within each of the windows. To sample in a particular window $w$, a 
-frame from this reactive trajectory is chosen such that the value of the 
-order parameter calculated for this frame falls within $w$. Using the shooting algorithm
+frame from this reactive trajectory is chosen such that the value of its 
+order parameter falls within $w$. Using the shooting algorithm
 the momenta are perturbed and dynamics is run for 10 steps forwards and backwards with a 
 $1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.    
-Within each window 3000 TPS trajectories are collected using the acceptance criteria 
-discussed in sec. \ref{ssec:algorithm}. The equilibrium probability distribution function 
-for the order parameter within window $w$ is then calculated as normalized histogram
-similar to the WHAM procedure. \cite{Kumar92JComputChem13p1011}
+Within each window 3000 TPS trajectories are collected. The 
+probability distribution function for the order parameter within window $w$ 
+is then calculated as normalized histograms which represent the populations 
+of the configurations within the window at equilibrium. 
+Then we calculate the free energy profile in a similar fashion to the WHAM 
+procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
+potential of mean force from umbrella sampling calculations. 
 %------------------------------------------------------------------------------ 

### 2022-04-23 tex/manuscript.tex
added little text and a figure

@@
 
-The reaction catalyzed by $pv$ADA is shown in  
-Fig. \ref{fig:ada-reaction} which is between the adenosine molecule and a hydroxide ion
-which results in the formation of inosine with ammonia as a by product. 
-$pv$ADA is an essential enzyme in the purine salvage pathway for plasmodium vivax and 
+$pv$ADA catalyzes the reaction between adenosine and hydroxide ion 
+resulting in the formation of inosine and ammonia as shown in
+Fig. \ref{fig:ada-reaction}.
+$pv$ADA is an essential enzyme in the purine salvage pathway and 
 inhibiting this enzyme is a promising approach to treat malaria. \cite{Madrid08JBiolChem283p35899}  
@@
 \subsection{Human Methionine Adenosyltransferase 2A}
-%------------------------------------------------------------------------------ 
-Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
-reaction between adenosine triphosphaste and L-methionine which leads to the 
-formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
-The mechanism of this reaction is experimentally determined to follow a 
-SN$_2$ transition state with the sulphur atom of the L-methionine as the 
-nucleophile with the triphosphate leaving group.
+The TPS ensemble was used to analyze the reaction pathway for the S$_{\text{N}}2$
+mechanism and a typical trajectory is provided here
+\begin{figure}
+\includegraphics[scale=0.5]{figures/mat2a-diff60.pdf}
+\caption{A reactive trajectory}
+\label{fig:mat2a-reactive-traj}
+\end{figure}
+The order parameters of the system along a typical reactive trajectory is 
+shown in Fig. \ref{fig:mat2a-reactive-traj}.
 

### 2022-04-25 tex/manuscript.tex
added very few changes today

@@
 \subsection{Human Methionine Adenosyltransferase 2A}
+The binding pocket image is shown for this enzyme here
+% width: 3.330, height: 1.873
 The TPS ensemble was used to analyze the reaction pathway for the S$_{\text{N}}2$
@@
 \begin{figure}
-\includegraphics[scale=0.5]{figures/mat2a-diff60.pdf}
+\includegraphics[scale=1.0]{figures/mat2a-diff60.pdf}
 \caption{A reactive trajectory}
@@
 \begin{figure}
-\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
+\includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
 \caption{The free energy as a function of the order parameter.}

### 2022-04-27 tex/manuscript.tex
added some pretty pictures

@@
 \begin{figure}
-\includegraphics[scale=0.6]{figures/mat2a-reaction.png}
+\includegraphics[width=0.5\textwidth]{figures/mat2a-reaction.png}
 \caption{The reaction catalyzed by human MAT2A enzyme}
@@
 \subsection{Human Methionine Adenosyltransferase 2A}
+\begin{figure}[ht!]
+\includegraphics[scale=0.2]{figures/mat2a-equil.png}
+\label{fig:mat2a-equil}
+\caption{Equilibriated structure of MAT2A enzyme}
+\end{figure}
+
 The binding pocket image is shown for this enzyme here
@@
 \begin{figure}[ht!]
-\includegraphics[scale=0.25]{figures/mat2a-transition-state.png}
+\includegraphics[scale=0.2]{figures/mat2a-transition-state.png}
+%\includegraphics[scale=0.2]{figures/mat2a-active-site2.png}
 \end{figure}
@@
 \begin{figure}[ht!]
-\includegraphics[scale=0.5]{figures/comm-60-mat2a.pdf}
+\includegraphics[scale=1.0]{figures/comm-60-mat2a.pdf}
 \end{figure}
@@
 \begin{figure}[ht!]
-\includegraphics[scale=0.15]{./figures/ada-protein.png}
+\centering
+\includegraphics[scale=0.2]{./figures/ada-equil.png}
+\caption{Equilibriated structure of ADA protein}
 \end{figure}
@@
 \begin{figure}[ht!]
-\includegraphics[scale=0.5]{./figures/ada-fenergy.pdf}
+\includegraphics[scale=1.0]{./figures/ada-fenergy.pdf}
+\label{fig:ada-fenergy}
+\caption{The free energy curve for the ADA reaction}
 \end{figure}

### 2022-04-29 tex/manuscript.tex
added more text and figures

@@
 diffusive reactions whereas it is less accurate otherwise.
+Transition interface sampling (TIS) has been used to obtain 
+reweighted path ensemble which can give access to free energy 
+and committor surfaces. \cite{Rogal10JChemPhys17p174109}
 
@@
 \subsection{Equilibrium distribution of order parameters}
-%Monte carlo methods in trajectory space the probability of a path is 
-%\begin{equation}
-%P(x_A)
-%\end{equation}
 The free energy as a function of the order parameter $\zeta$ is defined
@@
 (protein data bank ID: 7L1A) reported by Niland 
-et al. \cite{Niland21Biochem60p791} and the crystal structure of \textit{pv}ADA 
-in complex with MT-coformycin (protein data bank ID: 3EWC) reported by Ho et 
-al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
+et al. \cite{Niland21Biochem60p791,Ghosh21JAmChemSoc143p18325} and the crystal 
+structure of \textit{pv}ADA in complex with MT-coformycin (protein data bank ID: 3EWC) 
+reported by Ho et al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
 simulations. For calculations involving the MAT2A enzyme the substrates were modified to 
@@
 using the SHAKE procedure. \cite{Ryckaert77JComputPhys23p327} 
-
 %------------------------------------------------------------------------------ 
@@
 enzymes, respectively. 
-
 %------------------------------------------------------------------------------ 
@@
 \subsection{Human Methionine Adenosyltransferase 2A}
+The equilibrated structure of the MAT2A enzyme in complex with 
 \begin{figure}[ht!]
@@
 \label{fig:mat2a-equil}
-\caption{Equilibriated structure of MAT2A enzyme}
+\caption{Equilibriated structure of the MAT2A enzyme in complex with ATP and 
+MET along with 2 Mg$^{2+}$ ions.}
 \end{figure}
+The adenosyl part of the ATP is held together through $\pi-\pi$ interactions 
+by Phe250 and Ile117 residues.
 
-The binding pocket image is shown for this enzyme here
-% width: 3.330, height: 1.873
-The TPS ensemble was used to analyze the reaction pathway for the S$_{\text{N}}2$
-mechanism and a typical trajectory is provided here
 \begin{figure}
-\includegraphics[scale=1.0]{figures/mat2a-diff60.pdf}
+\includegraphics[scale=1.0]{figures/mat2a-diff167.pdf}
 \caption{A reactive trajectory}
@@
 
-\begin{figure}[ht!]
-\includegraphics[scale=1.0]{figures/comm-60-mat2a.pdf}
-\end{figure}
+\begin{figure*}[ht!]
+\centering
+\begin{minipage}[b]{0.45\linewidth}
+\includegraphics[width=\textwidth]{figures/comm-60-mat2a-nocons.pdf}
+%\caption{}
+\label{fig:minipage1}
+\end{minipage}
+\quad
+\begin{minipage}[b]{0.45\linewidth}
+\includegraphics[width=\textwidth]{figures/comm-60-mat2a.pdf}
+%\caption{}
+\label{fig:minipage2}
+\end{minipage}
+\caption{Committor distribution analysis}
+\end{figure*}
+
+The activation energy of this reaction was experimentally measured by 
+Niland et al. \cite{Niland21Biochem60p791} where they measured the rate constant of the reaction 
+at various temperatures while keeping the substrate concentrations near saturation.
+Using the Arrhenius equation they estimated the activation energy to be 
+$17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. Compared to typical enzyme catalyzed 
+reactions calculated activation energy is small.
 
@@
 \end{figure}
-
+From our TPS based calculations we calculated the free energy barrier to be 
+$16\;\text{Kcal mol}^{-1}$.
 %------------------------------------------------------------------------------ 
@@
 %------------------------------------------------------------------------------ 
+The equilibrated structure for the ADA enzyme complexed with adenosine, 
+OH$^{-}$ anion and Zn$^{2+}$ ion is shown in Fig. \ref{fig:ada-equil}.
+The Zn$^{2+}$ cation interacts with His42, His44 and His226 residues
+along with the hydroxide ion in the pocket. Glu229 is part of the 
+QM region and it acts as a proton donor to the adenosine molecule.
+
 \begin{figure}[ht!]
@@
 \caption{Equilibriated structure of ADA protein}
+\label{fig:ada-equil}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=1.0]{figures/ada-diff60.pdf}
+\caption{A reactive trajectory}
+\label{fig:ada-reactive-traj}
 \end{figure}
@@
 \end{figure}
-
-%\begin{figure}[ht]
-%\centering
-%\begin{minipage}[b]{0.32\linewidth}
-%\includegraphics[scale=0.3]{figures/reactant-state.png}
-%\caption{Reactant state}
-%\label{fig:minipage1}
-%\end{minipage}
-%\quad
-%\begin{minipage}[b]{0.32\linewidth}
-%\includegraphics[scale=0.3]{figures/int1.png}
-%\caption{Intermediate state}
-%\label{fig:minipage2}
-%\end{minipage}
-%\quad
-%\begin{minipage}[b]{0.32\linewidth}
-%\includegraphics[scale=0.3]{figures/product-state.png}
-%\caption{Product state}
-%\label{fig:minipage1}
-%\end{minipage}
-%\end{figure}
 %------------------------------------------------------------------------------ 

### 2022-05-01 tex/manuscript.tex
added some more images and text

@@
 milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
-(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
+(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,
+Dellago98JChemPhys108p1964,Bolhuis21AdvTheory4p2000237}
 
@@
 called virtual interface exchange TPS (VIE-TPS) to calculate the free energy 
-landscape. VIE-TPS is shown to produce reasonable free energies for 
-diffusive reactions whereas it is less accurate otherwise.
+landscape. VIE-TPS is shown to produce reasonably accurate free energy profiles for 
+chemical reactions that are diffusive in nature whereas it is less accurate otherwise.
 Transition interface sampling (TIS) has been used to obtain 
@@
 
+Committor analysis and committor distribution analysis are used to deduce 
+the transition states and the reaction coordinates, respectively. 
+Committor analysis is carried out for a reactive trajectory 
+from the TPS ensemble by choosing a time slice randomly from the trajectory, assigning
+momenta picked at random from the Boltzmann distribution for all the atoms, and running 
+dynamics for $250\;fs$. The commitment probability of a time slice is calculated by 
+repeating this procedure several times and finding the fraction of trajectories that go 
+to the reactant, product or neither. The transition state for the reaction is defined as 
+the structure from the time slice that has equal probability to commit to the reactant 
+and the product state. Repeating this process for several trajectories from the TPS
+ensemble results in a collection of transition states called the separatrix. 
+The reaction coordinate might contain not just the atoms from the molecules that 
+are directly involved in the reaction but also residues from side chains that can 
+promote the reaction. Taking this into account the committor distribution analysis
+starts by constraining the residues and molecules that are theorized to be part of the 
+reaction and using this constrained system trajectories are started
+
 \begin{figure}[ht!]
@@
 
+From the committor distribution analysis for a reactive trajectory we see that the 
+constrained region consisting of just the QM part which is the ATP and MET molecules 
+along with two Mg$^{2+}$ ions results in a distribution that is not peaked 
+at 0.5 whereas when the residues ... are constrained we obtain a distribution 
+that is peaked at 0.5. Therefore we conclude that the reaction coordinate consists
+of these residues as well. 
 \begin{figure*}[ht!]
@@
 \includegraphics[width=\textwidth]{figures/comm-60-mat2a-nocons.pdf}
-%\caption{}
 \label{fig:minipage1}
@@
 \includegraphics[width=\textwidth]{figures/comm-60-mat2a.pdf}
-%\caption{}
 \label{fig:minipage2}
@@
 The activation energy of this reaction was experimentally measured by 
-Niland et al. \cite{Niland21Biochem60p791} where they measured the rate constant of the reaction 
-at various temperatures while keeping the substrate concentrations near saturation.
-Using the Arrhenius equation they estimated the activation energy to be 
-$17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. Compared to typical enzyme catalyzed 
+Niland et al. \cite{Niland21Biochem60p791} where they measured the rate 
+constant of the reaction at various temperatures while keeping the 
+substrate concentrations near saturation. Using the Arrhenius equation 
+they estimated the activation energy to be $17.27\;\pm\;1.5\;
+\text{Kcal mol}^{-1}$. Compared to typical enzyme catalyzed 
 reactions calculated activation energy is small.
@@
 
+The formation of Meissenheimer complex is part of the 
+transition state as well as the partial protonation of the 
+N1 atom on the adenosyl moity as observed from experiments. 
+The reaction free energy is supposed to have two peaks with 
+an intermediate state and we observe the same from our calculations. 
 \begin{figure}[ht!]
@@
 
+\begin{figure}
+\centering
+\includegraphics[scale=0.12]{blender-images/ada/ada-trans.png}
+\label{fig:ada-trans}
+\caption{The transition state structure of the reaction catalyzed by the ADA enzyme}
+\end{figure}
+
 \begin{figure}[ht!]
@@
 \end{figure}
+
 %------------------------------------------------------------------------------ 
@@
 free energies of enzymatic reactions within the TPS method using the MAT2A
-and the ADA enzymes as examples.
+and the ADA enzymes as examples. The MAT2A enzymatic reaction consists of 
+the residues that are ... as part of its reaction coordinate. The reaction 
+proceeds through a SN2 mechanism and the transition state consists of 
+$dis(SC)$ being 2.3 and $dis(OC)$ being 2.03 {\AA}. From the calculations 
+we obtain a free energy barrier of 16.0 kcal mol which is in good agreement 
+with the 17.1 kcal mol measured from the experiments using the Arrhenius equation.
+Furthermore we looked at the reaction catalyzed by the ADA enzyme and 
+calculated the free energy profile for this reaction. Here we found two free energy
+barriers separated by an intermediate state which is consistent with 
+experimental findings. We calculate a free energy barrier of 25 kcal mol.
 

### 2022-05-03 tex/manuscript.tex
added a whole bunch of images and text

@@
 The equilibrated structure of the MAT2A enzyme in complex with 
+ATP, MET and two Mg$^{2+}$ ions is shown in Fig. \ref{fig:mat2a-equil}.
 \begin{figure}[ht!]
+\centering
 \includegraphics[scale=0.2]{figures/mat2a-equil.png}
-\label{fig:mat2a-equil}
 \caption{Equilibriated structure of the MAT2A enzyme in complex with ATP and 
 MET along with 2 Mg$^{2+}$ ions.}
+\label{fig:mat2a-equil}
 \end{figure}
-The adenosyl part of the ATP is held together through $\pi-\pi$ interactions 
-by Phe250 and Ile117 residues.
+In the binding pocket the adenosyl moiety and the Phe250 and Ile117 residues 
+are found to have $\pi-\pi$ interactions stabilizing the molecule. The triphosphate 
+moiety forms hydrogen bonds with ... residues.
 
@@
 \includegraphics[scale=1.0]{figures/mat2a-diff167.pdf}
-\caption{A reactive trajectory}
+\caption{Bond breaking ($dis(OC)$), bond forming ($dis(SC)$) distances for a typical 
+reactive trajectory for the reaction catalyzed by the MAT2A enzyme.}
 \label{fig:mat2a-reactive-traj}
@@
 
+%\begin{figure}[ht!]
+%\includegraphics[scale=0.2]{figures/mat2a-transition-state.png}
+%\includegraphics[scale=0.2]{figures/mat2a-active-site2.png}
+%\end{figure}
+
 \begin{figure}[ht!]
-\includegraphics[scale=0.2]{figures/mat2a-transition-state.png}
-%\includegraphics[scale=0.2]{figures/mat2a-active-site2.png}
+\includegraphics[scale=0.12]{blender-images/ada/mat2a-trans-labelled.png}
+\caption{Transition state structure of the reaction catalyzed by the human MAT2A enzyme.
+The S$-$C distance is $2.09$ {\AA} and the O$-$C distance is $2.32$ {\AA} which represents 
+an advanced transition state.}
+\label{fig:mat2a-trans-struct}
 \end{figure}
+The transition state structure follows a S$_{\text{N}}2$ 
+mechanism with a product like structure, the O$-$C bond is cleaved and 
+has a value of $2.32$ {\AA}.
+Experimentally the transition state structures of enzymatic reactions
+can be elucidated using kinetic isotope effects (KIE). \cite{Schramm99MetEnzym308p301}
 
@@
 \end{minipage}
-\caption{Committor distribution analysis}
+\caption{Committor distribution analysis for obtaining the reaction coordinate of the 
+MAT2A catalyzed reaction. The figure on the left has the QM region constrained and the figure 
+on the right has the QM region along with the Gln113, Ser114, Arg249 and Arg264 residues constrained.}
 \end{figure*}
@@
 The equilibrated structure for the ADA enzyme complexed with adenosine, 
-OH$^{-}$ anion and Zn$^{2+}$ ion is shown in Fig. \ref{fig:ada-equil}.
-The Zn$^{2+}$ cation interacts with His42, His44 and His226 residues
-along with the hydroxide ion in the pocket. Glu229 is part of the 
+OH$^{-}$ anion and Zn$^{2+}$ cation is shown in Fig. \ref{fig:ada-equil}.
+The Zn$^{2+}$ cation interacts with the His42, His44 and His226 residues
+and stabilizes the OH$^{-}$ anion in the pocket.
+
+Glu229 is part of the 
 QM region and it acts as a proton donor to the adenosine molecule.
+
+Experimentally KIE studies 
+from Schramm and coworkers \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
+predict the transition state structures for the plasmodium falciparum, human and bovine 
+ADA enzymes. They reported inhibition studies on the plasmodium vivax ADA by using 
+MT coformycin as the inhibitor molecule.
 
@@
 \centering
-\includegraphics[scale=0.12]{blender-images/ada/ada-trans.png}
+\includegraphics[scale=0.12]{blender-images/ada/new-ada-trans.png}
 \label{fig:ada-trans}
-\caption{The transition state structure of the reaction catalyzed by the ADA enzyme}
+\caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
+The formation of the Meissenheimer complex is the rate limiting step and the N-C distance is 
+1.60 {\AA} and the O-C distance is 1.48 {\AA}.}
 \end{figure}
@@
 \label{fig:ada-fenergy}
-\caption{The free energy curve for the ADA reaction}
+\caption{The free energy in $\text{Kcal}\;\text{mol}^{-1}$ as a function of the reaction 
+parameter for the ADA catalyzed reaction. The Meissenheimer complex where the OH$^{-}$ anion
+is bonded to the C atom and the leaving NH$_2$ moeity is also still bonded.}
 \end{figure}

### 2022-05-04 tex/manuscript.tex
made some more corrections

@@
+\pdfoptionpdfminorversion=4
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage{mathptmx}
-%\usepackage[scaled=0.92]{helvet}
+\usepackage[scaled=0.92]{helvet}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Free energy calculations based on the TPS shooting algorithm is 
-appealing but not a lot of examples exist in the literature 
-which compare the computational calculations with experimental
-results for enzymatic reactions.  
+appealing but rare in the literature. Applications of this 
+method to study enzymatic reactions and comparisons of the calculated 
+free energies to experimental measurements can shine light on the 
+accuracy of this approach compared to the standard methods used for 
+free energy calculations.    
 In this manuscript we implement and apply the 
@@
 enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
-These enzymes were chosen not only because of their importance in various 
+Human MAT2A catalyzes the formation of S-adenosyl L-methionine (SAM).
+SAM is an essential metabolite which is 
+distributed to almost all body tissues and fluids, is an universal 
+methyl donor and is of fundamental importance to the metabolism of 
+compounds such as hormones, neureotransmitters, proteins, 
+and nucleic acids. \cite{Friedel89Drugs38p389}
+plasmodium vivax is a parasite that is responsible for the most number 
+of cases of malaria globally. Targeting proteins that are essential to
+the metabolism of this bacteria such as the adenosine deaminase enzyme 
+is a good anti-malarial drug development strategy. \cite{Madrid08JBiolChem283p35899} 
+These enzymes were chosen not only because of their importance in 
 biochemical applications but also because of the availability of experimental 
-results with which comparisons can be drawn. The manuscript is organized as 
+results with which comparisons can be drawn. This manuscript is organized as 
 follows: first we discuss the TPS method and the algorithm that we use 
@@
 S-adenosyl-L-methionine (SAM) which is depicted in Fig. \ref{fig:mat2a-reaction}.
-SAM is an essential metabolite which is distributed to almost all 
-body tissues and fluids, is an universal methyl donor and is of
-fundamental importance to the metabolism of compounds such as hormones, neureotransmitters, proteins, 
-and nucleic acids. \cite{Friedel89Drugs38p389} This reaction was experimentally characterized by 
+ This reaction was experimentally characterized by 
 Firestone et al. \cite{Firestone17JAmChemSoc139p13754} to follow a S$_{\text{N}}2$ transition 
@@
 $pv$ADA is an essential enzyme in the purine salvage pathway and 
-inhibiting this enzyme is a promising approach to treat malaria. \cite{Madrid08JBiolChem283p35899}  
+inhibiting this enzyme is a promising approach to treat malaria.   
 This reaction proceeds with the nucleophilic O atom (O$^{\text{hydroxyl}}$) of the hydroxide anion
@@
 forces, the shooting algorithm was used to generate
-new trajectories within which first a random time slice is chosen from the reactive 
-trajectory and perturbing the momenta of the system. 
-The time step that we used for all our calculations was $1\;fs$ and we propagate the 
-dynamics for a total of 250 steps forward and 250 steps backward which amounts to a 
-total of $0.5\;ps$. 
-A total of 150 and 100 reactive trajectories were generated for the MAT2A and $pv$ADA
+new trajectories. The well known shooting algorithm begins with the choice of a 
+random time slice from the biased reactive trajectory and perturbing the momenta of the system.
+Then the dynamics is propagated for 250 steps forawards and backwards using $1\;fs$ time steps
+resulting in a trajectory with a time period of $0.5\;ps$.
+A total of 150 and 130 reactive trajectories were generated for the MAT2A and $pv$ADA
 enzymes, respectively. 
@@
 The Zn$^{2+}$ cation interacts with the His42, His44 and His226 residues
-and stabilizes the OH$^{-}$ anion in the pocket.
-
-Glu229 is part of the 
-QM region and it acts as a proton donor to the adenosine molecule.
-
+and stabilizes the OH$^{-}$ anion in the pocket. The N1 atom on the adensoyl 
+molecule hydrogen bonds with the carboxylic acid hydrogen atom of the Glu229 
+residue.  
+\begin{figure}[ht!]
+\centering
+\includegraphics[scale=0.2]{./figures/ada-equil.png}
+\caption{Equilibriated structure of ADA protein}
+\label{fig:ada-equil}
+\end{figure}
+The bond forming ($dis(OC)$) and bond breaking ($dis(NC)$) distances 
+for a reactive trajectory from the TPS ensemble is depicted 
+in Fig. \ref{fig:ada-reactive-traj}.
+\begin{figure}[ht!]
+\includegraphics[scale=1.0]{figures/ada-diff60.pdf}
+\caption{A reactive trajectory}
+\label{fig:ada-reactive-traj}
+\end{figure}
+Committor analysis is performed to determine the transition state of this reaction
+and the structure obtained from the calculations is shown in Fig. \ref{fig:ada-trans}.
+At the transition state the N$1-$H bond distance is $1.1\;${\AA}, suggesting that the 
+Glu229 residue has completed the proton donation to the adenosine molecule. This protonation
+then triggers a base rotation which can be observed    
+\begin{figure}
+\centering
+\includegraphics[scale=0.12]{blender-images/ada/new-ada-trans.png}
+\label{fig:ada-trans}
+\caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
+The formation of the Meissenheimer complex is the rate limiting step and the N-C distance is 
+1.60 {\AA} and the O-C distance is 1.48 {\AA}.}
+\end{figure}
 Experimentally KIE studies 
@@
 an intermediate state and we observe the same from our calculations.
-\begin{figure}[ht!]
-\centering
-\includegraphics[scale=0.2]{./figures/ada-equil.png}
-\caption{Equilibriated structure of ADA protein}
-\label{fig:ada-equil}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=1.0]{figures/ada-diff60.pdf}
-\caption{A reactive trajectory}
-\label{fig:ada-reactive-traj}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.12]{blender-images/ada/new-ada-trans.png}
-\label{fig:ada-trans}
-\caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
-The formation of the Meissenheimer complex is the rate limiting step and the N-C distance is 
-1.60 {\AA} and the O-C distance is 1.48 {\AA}.}
-\end{figure}
 

### 2022-05-05 tex/manuscript.tex
almost close to the finish line here

@@
 Configurations are sampled for the $pv$ADA system within the interval $[-3.0, 3.0]$ {\AA}
-of the order parameter value. 
-This interval is divided into 20 overlapping windows ($0.045$ {\AA} between 
-neighboring windows). A reactive trajectory from the TPS ensemble is chosen to be the guide trajectory 
+of the order parameter value which is divided into 20 overlapping 
+windows ($0.045$ {\AA} between neighboring windows). A reactive 
+trajectory from the TPS ensemble is chosen to be the guide trajectory 
 for sampling within each of the windows. To sample in a particular window $w$, a 
@@
 promote the reaction. Taking this into account the committor distribution analysis
-starts by constraining the residues and molecules that are theorized to be part of the 
-reaction and using this constrained system trajectories are started
+starts by constraining the residues and molecules that are hypothesized to be part of the 
+reaction coordinate and using this constrained system as the starting point 
+trajectories are launched assigning momenta at random from the Boltzmann distribution at 
+$300\;\text{K}$.
 
@@
 
+\begin{figure}
+\includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
+\caption{The free energy as a function of the order parameter.}
+\end{figure}
+From our TPS based calculations we calculated the free energy barrier to be 
+$16\;\text{Kcal mol}^{-1}$.
 The activation energy of this reaction was experimentally measured by 
@@
 reactions calculated activation energy is small.  
-
-\begin{figure}
-\includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
-\caption{The free energy as a function of the order parameter.}
-\end{figure}
-From our TPS based calculations we calculated the free energy barrier to be 
-$16\;\text{Kcal mol}^{-1}$.
 %------------------------------------------------------------------------------ 
@@
 \end{figure}
-Committor analysis is performed to determine the transition state of this reaction
-and the structure obtained from the calculations is shown in Fig. \ref{fig:ada-trans}.
-At the transition state the N$1-$H bond distance is $1.1\;${\AA}, suggesting that the 
-Glu229 residue has completed the proton donation to the adenosine molecule. This protonation
-then triggers a base rotation which can be observed    
+The deamination is an aromatic nucleophilic substitution reaction 
+where intermediate Meisenheimer complex is formed at the transition state
+as shown in Fig. \ref{fig:ada-trans}. 
+The approaching O atom of the hydroxide anion is at a distance of $1.48$ {\AA}
+from the C6 atom whereas the N atom of the amine leaving group is at $1.60$ {\AA}.
+The N$1-$H bond distance is $1.1\;${\AA} 
+suggesting that the proton transfer from the Glu229 residue is almost complete.
 \begin{figure}
@@
 \includegraphics[scale=0.12]{blender-images/ada/new-ada-trans.png}
-\label{fig:ada-trans}
 \caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
@@
 1.60 {\AA} and the O-C distance is 1.48 {\AA}.}
+\label{fig:ada-trans}
 \end{figure}
-Experimentally KIE studies 
-from Schramm and coworkers \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
-predict the transition state structures for the plasmodium falciparum, human and bovine 
-ADA enzymes. They reported inhibition studies on the plasmodium vivax ADA by using 
-MT coformycin as the inhibitor molecule.
-
-The formation of Meissenheimer complex is part of the 
-transition state as well as the partial protonation of the 
-N1 atom on the adenosyl moity as observed from experiments. 
-The reaction free energy is supposed to have two peaks with 
-an intermediate state and we observe the same from our calculations.
-
-\begin{figure}[ht!]
+These observations are consistent with the 
+experimental KIE studies from Luo et al. \cite{Luo07JAmChemSoc129p8008}
+who predict the transition state structure for the deamination reaction 
+catalyzed by the plasmodium falciparum ADA enzyme which shares
+$72\%$ identity with the $pv$ADA enzyme that we study here. The frequency of the 
+configurations (order parameter values) that are calculated within each of the 
+windows is used to  
+\begin{figure}[h!]
+\centering
 \includegraphics[scale=1.0]{./figures/ada-fenergy.pdf}
-\label{fig:ada-fenergy}
 \caption{The free energy in $\text{Kcal}\;\text{mol}^{-1}$ as a function of the reaction 
@@
 is bonded to the C atom and the leaving NH$_2$ moeity is also still bonded.}
+\label{fig:ada-fenergy}
 \end{figure}
-
+The free energy profile that we calculated is shown in Fig. \ref{fig:ada-fenergy}.
+The free energy barrier is calculated to be $23\;\text{Kcal}\;\text{mol}^{-1}$.
 %------------------------------------------------------------------------------ 
@@
 %------------------------------------------------------------------------------ 
-We have discussed the application of a scheme to calculate the 
-free energies of enzymatic reactions within the TPS method using the MAT2A
-and the ADA enzymes as examples. The MAT2A enzymatic reaction consists of 
-the residues that are ... as part of its reaction coordinate. The reaction 
-proceeds through a SN2 mechanism and the transition state consists of 
-$dis(SC)$ being 2.3 and $dis(OC)$ being 2.03 {\AA}. From the calculations 
-we obtain a free energy barrier of 16.0 kcal mol which is in good agreement 
-with the 17.1 kcal mol measured from the experiments using the Arrhenius equation.
-Furthermore we looked at the reaction catalyzed by the ADA enzyme and 
-calculated the free energy profile for this reaction. Here we found two free energy
-barriers separated by an intermediate state which is consistent with 
-experimental findings. We calculate a free energy barrier of 25 kcal mol.
+In this study, we presented an application of the TPS method for the calculation 
+of free energy profiles of enzymatic reactions. The order parameter along which the 
+free energy profile is to be obtained is partitioned into windows similar to 
+umbrella sampling. But instead of using bias potentials to sample the configurations
+within the windows the shooting algorithm is used to
+obtain an ensemble of trajectories according to the Monte carlo algorithm. A 
+trajectory is accepted if it visits the window at least in one of the time steps. 
+Using this ensemble of trajectories unbiased equilibrium probability distribution 
+function of the order parameter of the reaction within each window is obtained.
+Furthermore the free energy as a function of the order parameter is computed within 
+each of the windows and the entirte profile is obtained by shifting the free energies
+such that the function is continuous at the boundaries.
 
+The MAT2A enzymatic reaction was analyzed in detail to elucidate the transition state,
+reaction coordinate and the free energy profile. The reaction proceeds through a 
+S$_\text{N}$2 mechanism, the d$_{\text{OC}}$ and d$_{\text{SC}}$ are $2.1$ {\AA}
+and $2.3$ {\AA}, respectively. The reaction coordinate is determined to involve  
+the Gln113, Ser114, Arg249 and Arg264 residues. 
+The free energy barrier for this reaction was calculated to be 
+$16.0\;\text{Kcal}\;\text{mol}^{-1}$ which is in good agreement 
+with the experimentally measured activation energy of $17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. 
+Furthermore the reaction catalyzed by the $pv$ADA enzyme is analyzed in detail to obtain the  
+transition state and calculate the free energy profile.
+The transition state has a Meisenheimer complex like structure with the 
+d$_\text{NC}$ and d$_\text{OC}$ being $1.60$ {\AA} and $1.48$ {\AA}, respectively.
+The free energy barrier for this reaction for found to be $23\;\text{Kcal}\;\text{mol}^{-1}$
+with a double well shape similar to that measured from experiments.  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-05-05 tex/manuscript.tex
added some more text

@@
 The reaction catalyzed by the human MAT2A enzyme is
-between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules resulting in the formation of
-S-adenosyl-L-methionine (SAM) which is depicted in Fig. \ref{fig:mat2a-reaction}.
- This reaction was experimentally characterized by 
-Firestone et al. \cite{Firestone17JAmChemSoc139p13754} to follow a S$_{\text{N}}2$ transition 
-state. The key bond parameters for this reaction consists of the distance between the
-nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as $dis(SC)$), and the distance between 
-oxygen atom of the phosphate group and the $5'-C$ atom of ATP ($dis(OC)$).
-For the TPS calculations we define the reactant state to have $dis(OC)$ to be 
-less than 1.7 {\AA} and the product state to have the $dis(SC)$ less than 1.7 {\AA}.
+between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules 
+resulting in the formation of S-adenosyl-L-methionine (SAM) which is 
+depicted in Fig. \ref{fig:mat2a-reaction}. This reaction follows a S$_{\text{N}}2$ 
+mechanism and was experimentally characterized by 
+Firestone et al. \cite{Firestone17JAmChemSoc139p13754}. The key bond 
+parameters for this reaction consists of the distance between the
+nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as d$_{\mathrm{SC}}$), 
+and the distance between oxygen atom of the phosphate group and the $5'-C$ 
+atom of ATP (d$_{\mathrm{OC}}$). For the TPS calculations we define the 
+reactant state to have d$_{\mathrm{OC}}$ to be less than 1.7 {\AA} and the product 
+state to have the d$_{\mathrm{SC}}$ less than 1.7 {\AA}.
 
@@
 \includegraphics[scale=1.0]{figures/mat2a-diff167.pdf}
-\caption{Bond breaking ($dis(OC)$), bond forming ($dis(SC)$) distances for a typical 
+\caption{Bond breaking ($\mathrm{d}_{\mathrm{OC}}$), bond forming 
+($\mathrm{d}_{\mathrm{SC}}$) distances for a typical 
 reactive trajectory for the reaction catalyzed by the MAT2A enzyme.}
@@
 $300\;\text{K}$.
-
-%\begin{figure}[ht!]
-%\includegraphics[scale=0.2]{figures/mat2a-transition-state.png}
-%\includegraphics[scale=0.2]{figures/mat2a-active-site2.png}
-%\end{figure}
 
@@
 \includegraphics[scale=1.0]{figures/ada-diff60.pdf}
-\caption{A reactive trajectory}
+\caption{Bond breaking ($\mathrm{d}_{\mathrm{NC}}$), bond forming 
+($\mathrm{d}_{\mathrm{OC}}$) distances for a typical 
+reactive trajectory for the reaction catalyzed by the $pv$ADA enzyme.}
 \label{fig:ada-reactive-traj}
@@
 \begin{acknowledgement}
+SGB acknowledges funding from ...
 

### 2022-05-10 tex/manuscript.tex
added some more stuff

@@
 \begin{abstract}
-  Transition path sampling (TPS) has been widely applied for the
-  calculations of reaction rate, transition state and reaction 
-  coordinates of condensed phase systems. Here, we discuss a scheme 
-  for the calculation of free energies with the TPS method in combination 
-  with a window based sampling technique. This 
-  distribution can then be used for the calculations of free energies
-  using standard Boltzmann inversion schemes. We calculate the free
-  energy profiles of the reactions catalyzed by the human methionine 
-  S-adenosyltransferase (MAT2A) enzyme and the plasmodium 
-  vivax adenosine deaminase enzyme to demonstrate the applicability of this 
-  method. 
+  Transition path sampling (TPS) is widely used for the
+  calculations of reaction rates, transition state structures and reaction 
+  coordinates of condensed phase systems. Here we discuss a scheme 
+  for the calculation of free energies using the ensemble of TPS reactive 
+  trajectories in combination with a window based sampling technique 
+  %introduced by 
+  %Radhakrishnan et al. (Radhakrishnan, R.; Schlick, T., J. Chem. Phys. 2004, 121, 2436.) 
+  for enzyme catalyzed reactions. We calculate the free energy profiles of 
+  the reactions catalyzed by the human methionine S-adenosyltransferase (MAT2A) 
+  enzyme and the plasmodium vivax adenosine deaminase enzyme to assess 
+  the accuracy of this method. MAT2A catalyzes the formation of S-adenosine-L-methionine
+  following a S$_{\text{N}}2$ mechanism and using our method we estimate
+  the free energy barrier to be $16\;\text{Kcal}\;\text{mol}^{-1}$ which
+  is in excellent agreement with the experimentally measured activation 
+  energy of $17.27\;\text{Kcal}\;\text{mol}^{-1}$. Furthermore for the 
+  $pv$ADA enzyme catalyzed reaction we estimate a free energy barrier of 
+  $23\;\text{Kcal}\;\text{mol}^{-1}$ and the calcualted free energy profile is
+  similar to that predicted from experimental observations.
+  We show that calculating free energies within TPS is lucrative as it is 
+  bias free, accurate compared to experimental measurements 
+  and computationally reasonable.    
 \end{abstract}
@@
 \includegraphics[width=0.5\textwidth]{figures/mat2a-reaction.png}
-\caption{The reaction catalyzed by human MAT2A enzyme}
+\caption{The reaction catalyzed by the human MAT2A enzyme between ATP and L-methionine resulting in the 
+formation of S-adenosyl L-methionine and triphosphate leaving group.}
 \label{fig:mat2a-reaction}
@@
 
-$pv$ADA catalyzes the reaction between adenosine and hydroxide ion 
+$pv$ADA catalyzes the reaction between adenosine and OH$^{-}$ ion 
 resulting in the formation of inosine and ammonia as shown in
 Fig. \ref{fig:ada-reaction}.
-$pv$ADA is an essential enzyme in the purine salvage pathway and 
-inhibiting this enzyme is a promising approach to treat malaria.   
-This reaction proceeds with the nucleophilic O atom (O$^{\text{hydroxyl}}$) of the hydroxide anion
-attacking the electrophilic C6 atom of the adenosyl ring leading to the formation of a Meisenheimer 
-complex while the N1 atom on the adenosyl ring starts to get 
-protonated by the Glu229 residue of $pv$ADA. At the transition state, experimental findings 
-suggest that the dissociation of the C6$-$N6 bond as well as the protonation of the N1 
-atom are almost complete. \cite{Luo07JAmChemSoc129p8008} The key bond parameters are 
-the distance between the approaching hydroxyl oxygen and the C6 atom (denoted as $dis(OC)$) and the 
-dissociating C6$-$N6 ($dis(NC)$) bond distance. For the TPS simulaions we define the reactant state to 
-have $dis(NC)$ being less than $1.7$ {\AA} and the product state to have the 
-$dis(OC)$ less than $1.7$ {\AA}.
+The nucleophilic O atom of the OH$^{-}$ anion
+attacks the electrophilic C6 atom of the adenosyl ring leading to the 
+formation of a Meisenheimer complex like structure while the N1 atom 
+on the adenosyl moiety gets protonated by the Glu229 residue of $pv$ADA. 
+The key bond parameters are the distance between the approaching hydroxyl 
+oxygen and the C6 atom (denoted as d$_{\text{OC}}$) and the 
+dissociating C6$-$N6 (d$_{\text{NC}}$) bond distance. For the TPS simulaions 
+we define the reactant state to have d$_{\text{NC}}$ being less than $1.7$ {\AA} 
+and the product state to have the d$_{\text{OC}}$ less than $1.7$ {\AA}.
 
@@
 \includegraphics[scale=0.6]{figures/ada-new.png}
-\caption{The reaction catalyzed by adenosine deaminase enzyme}
+\caption{The reaction catalyzed by the $pv$ADA enzyme between adenosine and hydroxide ion
+resulting in the formation of inosine and ammonia.}
 \label{fig:ada-reaction}
@@
 A biased initial reactive trajectory is obtained by using harmonic restraint forces 
-of 1500 to break the $O-C$ bond and form the $S-C$ bond for the MAT2A enzyme and 
-to break the $N-C$ bond and form the $O-C$ bond for the case of the $pv$ADA enzyme. 
+for both the MAT2A and $pv$ADA systems. 
 Using the biased trajectory as the starting point and after removing the restraint 
-forces, the shooting algorithm was used to generate
-new trajectories. The well known shooting algorithm begins with the choice of a 
-random time slice from the biased reactive trajectory and perturbing the momenta of the system.
-Then the dynamics is propagated for 250 steps forawards and backwards using $1\;fs$ time steps
+forces, the shooting algorithm \cite{dellago02AdvChemPhys123} was used to generate
+new trajectories. The shooting algorithm begins with the choice of a 
+random time slice from the biased initial reactive trajectory and perturbs 
+the momenta of all the atoms in the system. Then the dynamics is propagated for 
+250 steps forawards and backwards using $1\;fs$ time steps
 resulting in a trajectory with a time period of $0.5\;ps$.
-A total of 150 and 130 reactive trajectories were generated for the MAT2A and $pv$ADA
-enzymes, respectively. 
+A total of 150 reactive trajectories were generated for both the MAT2A and $pv$ADA
+systems.
 %------------------------------------------------------------------------------ 
@@
 %------------------------------------------------------------------------------ 
-To obtain the free energy profile along the order parameter, we use the method 
-discussed in sec. \ref{ssec:algorithm}. The order paratemer for the MAT2A system 
-is defined to be $dis(OC)-dis(SC)$ and that for the $pv$ADA is defined as 
-$dis(NC)-dis(OC)$. For the MAT2A system, configurations are sampled within 
-the range $[-2.0,2.0]$ and this range is divided into 14 overlapping windows 
-($0.09$ {\AA} overlap on either side). 
-Configurations are sampled for the $pv$ADA system within the interval $[-3.0, 3.0]$ {\AA}
-of the order parameter value which is divided into 20 overlapping 
-windows ($0.045$ {\AA} between neighboring windows). A reactive 
-trajectory from the TPS ensemble is chosen to be the guide trajectory 
-for sampling within each of the windows. To sample in a particular window $w$, a 
-frame from this reactive trajectory is chosen such that the value of its 
+We follow the algorithm discussed previously to calculate the free energy
+profile for the two enzymatic reactions using the TPS reactive trajectories. 
+The order paratemer for the MAT2A system 
+is defined to be d$_{\text{OC}}-$d$_{\text{SC}}$ and that for the $pv$ADA is defined as 
+d$_{\text{NC}}-$d$_{\text{OC}}$. For the MAT2A system, configurations are sampled
+for order parameter values within the range $[-2.0,2.0]$ and this range is divided 
+into 14 overlapping windows ($0.09$ {\AA} between neighboring windows). 
+Configurations are sampled for the $pv$ADA system for order parameter values 
+within the interval $[-3.0, 3.0]$ {\AA} which is divided into 20 overlapping 
+windows ($0.09$ {\AA} between neighboring windows). A reactive 
+trajectory from the TPS ensemble is chosen to guide the sampling 
+within each of the windows in the same spirit as a steered MD trajectory 
+is used to guide the window based umbrella sampling simulations. 
+To sample in a particular window $w$, a frame from this reactive 
+trajectory is chosen such that the value of its 
 order parameter falls within $w$. Using the shooting algorithm
-the momenta are perturbed and dynamics is run for 10 steps forwards and backwards with a 
+the momenta of all the atoms in the system are perturbed 
+and dynamics is run for 10 steps forwards and backwards with a 
 $1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.
+Longer trajectories in our experience end up sampling outside the window 
+more often and hence we make use of very short trajectories in this study.     
 Within each window 3000 TPS trajectories are collected. The 
@@
 of the configurations within the window at equilibrium. 
-Then we calculate the free energy profile in a similar fashion to the WHAM 
+Finally we calculate the free energy profile in a similar fashion to the WHAM 
 procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
@@
 %------------------------------------------------------------------------------ 
-\subsection{Human Methionine Adenosyltransferase 2A}
+\subsection{Human MAT2A}
 The equilibrated structure of the MAT2A enzyme in complex with 
@@
 \end{figure}
-In the binding pocket the adenosyl moiety and the Phe250 and Ile117 residues 
-are found to have $\pi-\pi$ interactions stabilizing the molecule. The triphosphate 
-moiety forms hydrogen bonds with ... residues.
-
+The variation of the key distance parameters of the reaction calculated for 
+a reactive trajectory from the TPS ensemble are shown in Fig. \ref{fig:mat2a-reactive-traj}.
 \begin{figure}
@@
 \end{figure}
-The order parameters of the system along a typical reactive trajectory is 
-shown in Fig. \ref{fig:mat2a-reactive-traj}.
-
 Committor analysis and committor distribution analysis are used to deduce 
@@
 ensemble results in a collection of transition states called the separatrix.
+\begin{figure}[ht!]
+\includegraphics[scale=0.12]{blender-images/ada/mat2a-trans-labelled.png}
+\caption{Transition state structure of the reaction catalyzed by the human MAT2A enzyme.
+The bond parameters d$_{\text{SC}}$ and d$_{\text{OC}}$ are $2.31$ {\AA} and $2.09$ {\AA}, 
+respectively.}
+\label{fig:mat2a-trans-struct}
+\end{figure}
+The transition state resembles a S$_{\text{N}}2$ structure with the 
+d$_{\text{SC}}$ and d$_{\text{OC}}$ distances being $2.31$ {\AA} and $2.09$ {\AA}, respectively.
+Experimentally the transition state structures of enzymatic reactions
+can be elucidated using kinetic isotope effects (KIE). \cite{Schramm99MetEnzym308p301}
+Firestone et al. \cite{Firestone17JAmChemSoc139p13754} used a combination of experimental 
+KIE measurements using the whole protein system and computational density functional 
+theory (DFT) based gas phase simulations of representative molecules to elucidate the 
+transition state structure for this reaction. From their findings they report 
+d$_{\text{SC}}$ and d$_{\text{OC}}$ distances of $2.03$ {\AA} and $2.32$ {\AA}, 
+respectively and our calculated values are in close agreement to this result.
+
 The reaction coordinate might contain not just the atoms from the molecules that 
@@
 trajectories are launched assigning momenta at random from the Boltzmann distribution at 
-$300\;\text{K}$.
-
-\begin{figure}[ht!]
-\includegraphics[scale=0.12]{blender-images/ada/mat2a-trans-labelled.png}
-\caption{Transition state structure of the reaction catalyzed by the human MAT2A enzyme.
-The S$-$C distance is $2.09$ {\AA} and the O$-$C distance is $2.32$ {\AA} which represents 
-an advanced transition state.}
-\label{fig:mat2a-trans-struct}
-\end{figure}
-The transition state structure follows a S$_{\text{N}}2$ 
-mechanism with a product like structure, the O$-$C bond is cleaved and 
-has a value of $2.32$ {\AA}.
-Experimentally the transition state structures of enzymatic reactions
-can be elucidated using kinetic isotope effects (KIE). \cite{Schramm99MetEnzym308p301}
-
-From the committor distribution analysis for a reactive trajectory we see that the 
-constrained region consisting of just the QM part which is the ATP and MET molecules 
-along with two Mg$^{2+}$ ions results in a distribution that is not peaked 
-at 0.5 whereas when the residues ... are constrained we obtain a distribution 
-that is peaked at 0.5. Therefore we conclude that the reaction coordinate consists
-of these residues as well. 
+$300\;\text{K}$ to all the atoms in the system. 
+First we constrained just the QM region in our committor distribution analysis 
+which consists of ATP and MET molecules along with two Mg$^{2+}$ ions and the 
+resulting distribution as shown in Fig. \ref{fig:mat2a-comm-dist} is not peaked 
+at 0.5 which is expected for the reaction coordinate. 
+Next we constrained the residues Gln113, Ser114, Arg249 and Arg264 along with the 
+QM region and obtain a distribution that is peaked at 0.5 and hence we 
+conclude that the reaction coordinate consists of these residues as well. 
 \begin{figure*}[ht!]
@@
 on the right has the QM region along with the Gln113, Ser114, Arg249 and Arg264 residues constrained.}
+\label{fig:mat2a-comm-dist}
 \end{figure*}
-
-\begin{figure}
-\includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
-\caption{The free energy as a function of the order parameter.}
-\end{figure}
-From our TPS based calculations we calculated the free energy barrier to be 
-$16\;\text{Kcal mol}^{-1}$.
 The activation energy of this reaction was experimentally measured by 
@@
 they estimated the activation energy to be $17.27\;\pm\;1.5\;
-\text{Kcal mol}^{-1}$. Compared to typical enzyme catalyzed 
-reactions calculated activation energy is small.  
+\text{Kcal mol}^{-1}$ which is small Compared to typical enzyme catalyzed 
+reactions.
+\begin{figure}
+\includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
+\caption{The free energy as a function of the order parameter.}
+\label{fig:mat2a-fenergy}
+\end{figure}
+The free energy profile from our TPS based calculations is shown in 
+Fig. \ref{fig:mat2a-fenergy} where the barrier is $16\;\text{Kcal mol}^{-1}$.
+This is in excellent agreement with the experimental result.  
 %------------------------------------------------------------------------------ 
-\subsection{Adenosine deaminase reaction coordinates and free energy}
+\subsection{$pv$ADA}
 %------------------------------------------------------------------------------ 
@@
 \end{figure}
-The bond forming ($dis(OC)$) and bond breaking ($dis(NC)$) distances 
+The bond forming (d$_{\text{OC}}$) and bond breaking (d$_{\text{NC}}$) distances 
 for a reactive trajectory from the TPS ensemble is depicted 
@@
 The deamination is an aromatic nucleophilic substitution reaction 
-where intermediate Meisenheimer complex is formed at the transition state
-as shown in Fig. \ref{fig:ada-trans}. 
-The approaching O atom of the hydroxide anion is at a distance of $1.48$ {\AA}
+where an intermediate Meisenheimer complex like structure is formed 
+at the transition state as shown in Fig. \ref{fig:ada-trans}. 
+The approaching O atom of the OH$^{-}$ anion is at a distance of $1.48$ {\AA}
 from the C6 atom whereas the N atom of the amine leaving group is at $1.60$ {\AA}.
@@
 \caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
-The formation of the Meissenheimer complex is the rate limiting step and the N-C distance is 
-1.60 {\AA} and the O-C distance is 1.48 {\AA}.}
+The formation of the Meissenheimer complex like structure with 
+d$_{\text{NC}}=1.60$ {\AA} and d$_{\text{OC}}=1.48$ {\AA} is observed.}
 \label{fig:ada-trans}
@@
 who predict the transition state structure for the deamination reaction 
-catalyzed by the plasmodium falciparum ADA enzyme which shares
-$72\%$ identity with the $pv$ADA enzyme that we study here. The frequency of the 
-configurations (order parameter values) that are calculated within each of the 
-windows is used to  
+catalyzed by the plasmodium falciparum ADA ($pf$ADA) enzyme which shares
+$72\%$ identity with the $pv$ADA enzyme that we analyze here. The next
+step in the reaction is the proton transfer from the OH$^{-}$ ion to the 
+amine leaving group resulting in the formation of inosine and ammonia.   
+%At the transition state, experimental findings 
+%suggest that the dissociation of the C6$-$N6 bond as well as the protonation of the N1 
+%atom are almost complete. \cite{Luo07JAmChemSoc129p8008}
+The free energy profile that we calculated is shown in Fig. \ref{fig:ada-fenergy}.  
 \begin{figure}[h!]
@@
 \end{figure}
-The free energy profile that we calculated is shown in Fig. \ref{fig:ada-fenergy}.
-The free energy barrier is calculated to be $23\;\text{Kcal}\;\text{mol}^{-1}$.
+The free energy barrier is calculated to be $23\;\text{Kcal}\;\text{mol}^{-1}$
+and the overall shape of the function is very similar to that hypothesized 
+by Luo et al. \cite{Luo07JAmChemSoc129p8008}
 %------------------------------------------------------------------------------ 

### 2022-05-10 tex/manuscript.tex
final corrections

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Sree ganesh balasubramani}
+\author{Sree Ganesh Balasubramani}
 \affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
@@
 \title[]
-  {Transition path sampling based calculations of free energies of enzymatic
+  {Transition path sampling based calculations of free energies for enzymatic
   reactions: the case of human methionine adenosyl transferase and plasmodium 
@@
   trajectories in combination with a window based sampling technique 
-  %introduced by 
-  %Radhakrishnan et al. (Radhakrishnan, R.; Schlick, T., J. Chem. Phys. 2004, 121, 2436.) 
   for enzyme catalyzed reactions. We calculate the free energy profiles of 
   the reactions catalyzed by the human methionine S-adenosyltransferase (MAT2A) 
-  enzyme and the plasmodium vivax adenosine deaminase enzyme to assess 
+  enzyme and the plasmodium vivax adenosine deaminase ($pv$ADA) enzyme to assess 
   the accuracy of this method. MAT2A catalyzes the formation of S-adenosine-L-methionine
   following a S$_{\text{N}}2$ mechanism and using our method we estimate
-  the free energy barrier to be $16\;\text{Kcal}\;\text{mol}^{-1}$ which
+  the free energy barrier for this reaction to 
+  be $16\;\text{Kcal}\;\text{mol}^{-1}$ which
   is in excellent agreement with the experimentally measured activation 
-  energy of $17.27\;\text{Kcal}\;\text{mol}^{-1}$. Furthermore for the 
+  energy of $17.27\;\text{Kcal}\;\text{mol}^{-1}$. Furthermore, for the 
   $pv$ADA enzyme catalyzed reaction we estimate a free energy barrier of 
-  $23\;\text{Kcal}\;\text{mol}^{-1}$ and the calcualted free energy profile is
+  $23\;\text{Kcal}\;\text{mol}^{-1}$ and the calculated free energy profile is
   similar to that predicted from experimental observations.
-  We show that calculating free energies within TPS is lucrative as it is 
-  bias free, accurate compared to experimental measurements 
-  and computationally reasonable.    
+  Calculating free energies employing our simple method within TPS 
+  provides significant advantages over such methods as umbrella sampling 
+  since it is free from any applied external bias, accurate compared to 
+  experimental measurements and has a reasonable computational cost.    
 \end{abstract}
@@
 Molecular dynamics (MD) simulations are increasingly being
-used to calculate quantitative estimations of experimentally measurable 
+used to calculate quantitative estimates of experimentally measurable 
 kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
@@
 enzyme activity, fast high throughput screening of thousands of inhibitor 
-molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
+molecules in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
 directed evolution for designing enzymes with enhanced catalytic 
@@
 several enhanced sampling techniques 
-have been developed in the past few decades some of which are 
+have been developed in the past few decades. These include 
 umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
@@
 
-TPS in particular is an attractive method since it samples reactive trajectories
+TPS is an attractive method since it samples reactive trajectories
 in which the rare barrier-crossing events are guaranteed to occur and
@@
 complex for large biological systems.
-TPS is based on the Monte Carlo importance sampling in the 
+TPS is based on Monte Carlo importance sampling in the 
 space of trajectories. It uses the shooting and shifting algorithm 
@@
 \cite{Antoniou01JPhysChemB105p5553,Hay12NatChem4p161,Schwartz09NatChemBiol5p551} 
-protein folding, \cite{Bolhuis03ProcNatlAcadSci100p12129,Juraszek12ChemPhys396p30} etc.
+and protein folding. \cite{Bolhuis03ProcNatlAcadSci100p12129,Juraszek12ChemPhys396p30}
 
 Calculating equilibrium properties such as the free energy of an enzymatic 
-catalysis reaction within TPS is not straighforward. 
+catalysis reaction within TPS is not straightforward. 
 The Monte Carlo algorithm used in TPS has an acceptance criterion that 
@@
 the product basin or vice versa, hence trajectories that are localized in the 
-reactant or product basins are rejected. The localized trajectories contribute
-to the probability density distribution function ($P(\lambda)$) which is the 
-probability of a system at equilibrium to have a particular value of the 
-order parameter $\lambda$. \cite{Dellago09AdvCompSimAppp167}
+reactant or product basins, without visiting both are rejected. 
+If one defines an order parameter $\zeta$, which separates reactant from 
+product basins, then all trajectories, reactive and non-reactive, 
+contribute to the probability density distribution function 
+($P(\zeta)$) which is the probability of a system at equilibrium to have 
+a particular value of the order parameter. \cite{Dellago09AdvCompSimAppp167}
 
-Frenkel has shown that inclusion of rejected states in Markov state Monte 
-Carlo algorith can help speed up the convergence of the calculations of 
-order-parameter distribution of a two-dimensional
+Frenkel has shown that inclusion of rejected states in a Markov state Monte 
+Carlo algorithm can help speed up the convergence of the calculations of 
+the order parameter distribution of a two-dimensional
 Ising model. \cite{Frenkel04ProcNatAcadSci101p17571}
@@
 More recently Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
-have developed a reweighted path ensemble scheme within the standard TPS method 
+have developed a re-weighted path ensemble scheme within the standard TPS method 
 called virtual interface exchange TPS (VIE-TPS) to calculate the free energy 
@@
 Transition interface sampling (TIS) has been used to obtain 
-reweighted path ensemble which can give access to free energy 
+a re-weighted path ensemble which can give access to free energies 
 and committor surfaces. \cite{Rogal10JChemPhys17p174109}
@@
 SAM is an essential metabolite which is 
-distributed to almost all body tissues and fluids, is an universal 
+distributed to almost all body tissues and fluids, is a universal 
 methyl donor and is of fundamental importance to the metabolism of 
@@
 and nucleic acids. \cite{Friedel89Drugs38p389}
-plasmodium vivax is a parasite that is responsible for the most number 
+Plasmodium vivax is a parasite that is responsible for the largest number 
 of cases of malaria globally. Targeting proteins that are essential to
 the metabolism of this bacteria such as the adenosine deaminase enzyme 
-is a good anti-malarial drug development strategy. \cite{Madrid08JBiolChem283p35899} 
+is a possible anti-malarial drug development strategy. \cite{Madrid08JBiolChem283p35899} 
 These enzymes were chosen not only because of their importance in 
@@
 follows: first we discuss the TPS method and the algorithm that we use 
-to calculate the free energies briefly, then we show calculations which apply this 
-method on the two enzymatic reactions. Finally we compare our results to expeiments
+to calculate the free energies, then we show calculations which apply this 
+method on the two enzymatic reactions. Finally we compare our results to experiments
 and provide conclusions. 
@@
 Here we briefly describe the equations describing the TPS method for the purpose of
-introducing notations as well as for the sake of being complete. 
+introducing notations as well as for the sake of completeness. 
 Consider the molecular dynamics simulation of a molecular system consisting 
@@
 Markovian, the probability to go from $z_{i\Delta t}$ to $z_{(i+1)\Delta t}$ depends only 
-on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probablility can 
+on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probability can 
 be expressed as the product of individual probabilities $p(i\rightarrow j)$ as 
@@
 \]
-The set of all reactive trajectories characterized by the distribution function
-given by Eq. \ref{eqn:tpsensem} is the transition path ensemble.
+and $Z_{\mathcal{AB}}(\mathcal{T})$ is the normalization factor for this 
+probability distribution function.
+The set of all reactive trajectories characterized by the probability 
+distribution function given by Eq. \ref{eqn:tpsensem} is the 
+transition path ensemble.
 
 \subsection{Equilibrium distribution of order parameters}
-The free energy as a function of the order parameter $\zeta$ is defined
-as 
+The free energy of a molecular system as a function of the order parameter $\zeta$ 
+is defined as 
 \begin{equation}
@@
 \end{equation}
-where $P(\zeta)$ is the probability distribution of the reaction coordinate
-$\zeta$ and $C$ is an arbitrary constant. At equilibrium the probability distribution of the 
-reaction coordinate 
-can be obtained from the distribution function for the phase space $\rho(\textbf{q})$ as
+where $P(\zeta)$ is the probability distribution function of the reaction coordinate
+$\zeta$, $C$ is an arbitrary constant, $k_{\text{B}}$ is the Boltzmann constant
+and $T$ is the temperature. At equilibrium the probability distribution of the 
+reaction coordinate can be obtained from the distribution function for the phase 
+space $\rho(\textbf{q})$ as
 \begin{equation}
@@
 \end{equation}
-The integration is over the entire phase space. In practice the calculation of this distribution function
+The integration is over the entire phase space. In practice, the calculation of this distribution function
 often proceeds using histogram based methods where the reaction coordinate is 
-divided into bins and the frequency of the occurence of the reaction coordinate within a particular window
-during the course of molecular dynamics simulations is used to calculate the probability distribution. 
-Since some values of the reaction coordinates particulary close to the transition state are 
+divided into bins and the frequency of the occurrence of the reaction coordinate 
+within a particular window during the course of MD simulation is used to calculate 
+the probability distribution function. Since some values of the reaction coordinates 
+particularly close to the transition state are 
 rarely sampled in a conventional MD simulations, enhanced sampling techniques such as the umbrella sampling
 are necessary to sample the low probability regions. TPS is designed to sample these regions in phase space 
-without the application of any external bias, but the trajectories sampled with TPS are only reactive and hence
+without the application of any external bias, but the trajectories 
+sampled with TPS are only reactive and hence
 they are not distributed according to the equilibrium distribution function. 
-The restriction that the pathways start from the reactant state and end in the product state can be relaxed
-to obtain an ensemble of trajectories that will be distributed according to the equilibrium distribution. 
+The restriction that the pathways start from the reactant state and end in the 
+product state can be relaxed to obtain an ensemble of trajectories that will 
+be distributed according to the equilibrium distribution. 
 %------------------------------------------------------------------------------ 
@@
 which consists only of reactive trajectories.
-\item Divide the order parameter into windows $\{\zeta_i\}$ within which the 
-order parameters are given by 
-$\zeta_{i}^{min} < \zeta < \zeta_{i}^{max}$.
+\item The order parameter is divided into windows $\{\zeta_i\}$ and within each window,
+$\zeta_{i}^{min} < \zeta < \zeta_{i}^{max}$ is to be sampled.
 \item Choose a reactive trajectory $\tilde{z}(\mathcal{T})$ from the TPS ensemble as 
@@
 harvest dynamics trajectories starting from the time slice $\tilde{z}_{j\Delta t}$ of the trajectory 
-$\tilde{z}(\mathcal{T})$ and accept the new trajectory if at any time slice the calculated 
+$\tilde{z}(\mathcal{T})$ and accept the new trajectory if for any time slice the calculated 
 order parameter $\hat{\zeta}$ satisfies $\zeta_{i}^{min} < \hat{\zeta} < \zeta_{i}^{max}$.
@@
 < \zeta < \zeta_{i}^{max}$ by constructing histograms obtained from the ensemble of 
-accepted trajectories for the window $i$. 
-\item Combine the probability distributions in successive windows by adjusting the constants $C$
-such that the free energy $A$ from Eq. \ref{eqn:fenergy} is continuous. 
+accepted trajectories for the window referenced by the index $i$. 
+\item Calculate the free energies for each of the windows according to Eq. \ref{eqn:fenergy}
+and combine them such that the function $\text{A}(\zeta)$ is continuous 
+by adjusting the constants $C$. 
 \end{itemize}
@@
 constraints on all atoms except on the H atoms and the TIP3P 
-water molecules with gradual reduction of the restraint forces to 1. 
+water molecules with gradual reduction of the restraint forces. 
 15 ps of equilibration was carried out starting with harmonic 
-constraints of force constant 1 followed by 20 ps of contraint free 
+restraint forces followed by 20 ps of constraint free 
 equilibration to prepare the systems for TPS simulations. 
@@
 reactant state to have d$_{\mathrm{OC}}$ to be less than 1.7 {\AA} and the product 
-state to have the d$_{\mathrm{SC}}$ less than 1.7 {\AA}.
+state to have the d$_{\mathrm{SC}}$ less than 2.0 {\AA}.
 
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{figures/mat2a-reaction.png}
+\includegraphics[width=\columnwidth]{figures/mat2a-reaction.png}
 \caption{The reaction catalyzed by the human MAT2A enzyme between ATP and L-methionine resulting in the 
@@
 oxygen and the C6 atom (denoted as d$_{\text{OC}}$) and the 
-dissociating C6$-$N6 (d$_{\text{NC}}$) bond distance. For the TPS simulaions 
-we define the reactant state to have d$_{\text{NC}}$ being less than $1.7$ {\AA} 
-and the product state to have the d$_{\text{OC}}$ less than $1.7$ {\AA}.
+dissociating C6$-$N6 (d$_{\text{NC}}$) bond distance. For the TPS simulations 
+we define the reactant state to have d$_{\text{NC}}$ being less than $1.8$ {\AA} 
+and the product state to have the d$_{\text{OC}}$ less than $1.8$ {\AA}.
 
 \begin{figure}
-\includegraphics[scale=0.6]{figures/ada-new.png}
+\includegraphics[width=\columnwidth]{figures/ada-new-reaction.png}
 \caption{The reaction catalyzed by the $pv$ADA enzyme between adenosine and hydroxide ion
@@
 the momenta of all the atoms in the system. Then the dynamics is propagated for 
-250 steps forawards and backwards using $1\;fs$ time steps
+250 steps forwards and backwards using $1\;fs$ time steps
 resulting in a trajectory with a time period of $0.5\;ps$.
@@
 profile for the two enzymatic reactions using the TPS reactive trajectories. 
-The order paratemer for the MAT2A system 
+The order parameter for the MAT2A system 
 is defined to be d$_{\text{OC}}-$d$_{\text{SC}}$ and that for the $pv$ADA is defined as 
@@
 within each of the windows in the same spirit as a steered MD trajectory 
-is used to guide the window based umbrella sampling simulations. 
-To sample in a particular window $w$, a frame from this reactive 
+is used to guide the window based umbrella sampling simulations (though in this case,
+of course, this is an unbiased dynamically exact reactive trajectory). 
+To sample in a particular window $\zeta_i$, a frame from this reactive 
 trajectory is chosen such that the value of its 
-order parameter falls within $w$. Using the shooting algorithm
+order parameter ($\zeta$) falls within 
+$\zeta_{i}^{min} < \zeta < \zeta_{i}^{max}$. Using the shooting algorithm
 the momenta of all the atoms in the system are perturbed 
@@
 \includegraphics[scale=0.2]{figures/mat2a-equil.png}
-\caption{Equilibriated structure of the MAT2A enzyme in complex with ATP and 
+\caption{Equilibrated structure of the MAT2A enzyme in complex with ATP and 
 MET along with 2 Mg$^{2+}$ ions.}
@@
 dynamics for $250\;fs$. The commitment probability of a time slice is calculated by 
-repeating this procedure several times and finding the fraction of trajectories that go 
+repeating this procedure many times and finding the fraction of trajectories that go 
 to the reactant, product or neither. The transition state for the reaction is defined as 
@@
 are directly involved in the reaction but also residues from side chains that can 
-promote the reaction. Taking this into account the committor distribution analysis
-starts by constraining the residues and molecules that are hypothesized to be part of the 
+promote the reaction. \cite{Schramm18Biochem57p3299} Taking this into account the 
+committor distribution analysis
+begins by constraining the residues and molecules that are hypothesized to be part of the 
 reaction coordinate and using this constrained system as the starting point 
@@
 they estimated the activation energy to be $17.27\;\pm\;1.5\;
-\text{Kcal mol}^{-1}$ which is small Compared to typical enzyme catalyzed 
-reactions.
+\text{Kcal mol}^{-1}$.
 \begin{figure}
 \includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
-\caption{The free energy as a function of the order parameter.}
+\caption{The free energy as a function of the order 
+parameter for the reaction catalyzed
+by the human MAT2A enzyme. The activation barrier is 
+calculated to be $16\;\text{Kcal mol}^{-1}$. The solid colored disks represent the free energies 
+calculated within the individual windows which is adjusted to make the free energy as a function 
+of the order parameter a continuous function.}
 \label{fig:mat2a-fenergy}
@@
 \includegraphics[scale=0.2]{./figures/ada-equil.png}
-\caption{Equilibriated structure of ADA protein}
+\caption{Equilibrated structure of $pv$ADA enzyme complexed with adenosine, 
+OH$^{-}$ anion and Zn$^{2+}$.}
 \label{fig:ada-equil}
@@
 at the transition state as shown in Fig. \ref{fig:ada-trans}. 
-The approaching O atom of the OH$^{-}$ anion is at a distance of $1.48$ {\AA}
-from the C6 atom whereas the N atom of the amine leaving group is at $1.60$ {\AA}.
+The approaching O atom of the OH$^{-}$ anion is at a distance of $1.49$ {\AA}
+from the C6 atom whereas the N atom of the amine leaving group is at $1.62$ {\AA}
+from the C6 atom.
 The N$1-$H bond distance is $1.1\;${\AA} 
@@
 \caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
-The formation of the Meissenheimer complex like structure with 
-d$_{\text{NC}}=1.60$ {\AA} and d$_{\text{OC}}=1.48$ {\AA} is observed.}
+The formation of the Meisenheimer complex like structure with 
+d$_{\text{NC}}=1.62$ {\AA} and d$_{\text{OC}}=1.49$ {\AA} is observed.}
 \label{fig:ada-trans}
@@
 \includegraphics[scale=1.0]{./figures/ada-fenergy.pdf}
-\caption{The free energy in $\text{Kcal}\;\text{mol}^{-1}$ as a function of the reaction 
-parameter for the ADA catalyzed reaction. The Meissenheimer complex where the OH$^{-}$ anion
-is bonded to the C atom and the leaving NH$_2$ moeity is also still bonded.}
+\caption{The free energy as a function of the reaction 
+parameter for the ADA catalyzed reaction. The activation barrier is calculated as 
+$21\;\text{Kcal}\;\text{mol}^{-1}$. The solid colored disks represent the free energies 
+calculated within the individual windows which is adjusted to make the free energy as a function 
+of the order parameter a continuous function.}
 \label{fig:ada-fenergy}
 \end{figure}
-The free energy barrier is calculated to be $23\;\text{Kcal}\;\text{mol}^{-1}$
+The free energy barrier is calculated to be $21\;\text{Kcal}\;\text{mol}^{-1}$
 and the overall shape of the function is very similar to that hypothesized 
@@
 of free energy profiles of enzymatic reactions. The order parameter along which the 
-free energy profile is to be obtained is partitioned into windows similar to 
-umbrella sampling. But instead of using bias potentials to sample the configurations
+free energy profile is to be obtained is partitioned into windows as is done in 
+umbrella sampling, but instead of using bias potentials to sample the configurations
 within the windows the shooting algorithm is used to
-obtain an ensemble of trajectories according to the Monte carlo algorithm. A 
+obtain an ensemble of trajectories according to the Monte Carlo algorithm. A 
 trajectory is accepted if it visits the window at least in one of the time steps. 
-Using this ensemble of trajectories unbiased equilibrium probability distribution 
+Using this ensemble of trajectories the unbiased equilibrium probability distribution 
 function of the order parameter of the reaction within each window is obtained.
 Furthermore the free energy as a function of the order parameter is computed within 
-each of the windows and the entirte profile is obtained by shifting the free energies
+each of the windows and the entire profile is obtained by shifting the free energies
 such that the function is continuous at the boundaries.
@@
 reaction coordinate and the free energy profile. The reaction proceeds through a 
-S$_\text{N}$2 mechanism, the d$_{\text{OC}}$ and d$_{\text{SC}}$ are $2.1$ {\AA}
-and $2.3$ {\AA}, respectively. The reaction coordinate is determined to involve  
+S$_\text{N}$2 mechanism, the d$_{\text{OC}}$ and d$_{\text{SC}}$ are $2.09$ {\AA}
+and $2.31$ {\AA}, respectively. The reaction coordinate is determined to involve  
 the Gln113, Ser114, Arg249 and Arg264 residues. 
 The free energy barrier for this reaction was calculated to be 
-$16.0\;\text{Kcal}\;\text{mol}^{-1}$ which is in good agreement 
+$16\;\text{Kcal}\;\text{mol}^{-1}$ which is in good agreement 
 with the experimentally measured activation energy of $17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. 
@@
 The transition state has a Meisenheimer complex like structure with the 
-d$_\text{NC}$ and d$_\text{OC}$ being $1.60$ {\AA} and $1.48$ {\AA}, respectively.
-The free energy barrier for this reaction for found to be $23\;\text{Kcal}\;\text{mol}^{-1}$
+d$_\text{NC}$ and d$_\text{OC}$ being $1.62$ {\AA} and $1.49$ {\AA}, respectively.
+The free energy barrier for this reaction for found to be $21\;\text{Kcal}\;\text{mol}^{-1}$
 with a double well shape similar to that measured from experiments. 
+These results show that free energy simulations based on unbiased transition 
+path sampling computations can result in accurate and computationally
+efficient predictions of both free energy profiles and mechanisms.  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{acknowledgement}
-SGB acknowledges funding from ...
-
+SGB acknowledges funding from the NIH through grant number GM127594.
 \end{acknowledgement}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{suppinfo}
+%\begin{suppinfo}
 
-\end{suppinfo}
+%\end{suppinfo}
 

### 2022-06-05 tex/manuscript.tex
changes made to the manuscript

@@
 
-$pv$ADA catalyzes the reaction between adenosine and OH$^{-}$ ion 
+$pv$ADA catalyzes the reaction between adenosine and OH$^{-}$ anion
 resulting in the formation of inosine and ammonia as shown in
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figures/ada-new-reaction.png}
+\includegraphics[width=\columnwidth]{figures/ada-reaction.png}
 \caption{The reaction catalyzed by the $pv$ADA enzyme between adenosine and hydroxide ion
@@
 \begin{figure}[ht!]
-\includegraphics[scale=0.12]{blender-images/ada/mat2a-trans-labelled.png}
+\includegraphics[scale=0.12]{figures/mat2a-trans-labelled.png}
 \caption{Transition state structure of the reaction catalyzed by the human MAT2A enzyme.
@@
 \centering
-\includegraphics[scale=0.12]{blender-images/ada/new-ada-trans.png}
+\includegraphics[scale=0.12]{figures/new-ada-trans.png}
 \caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
@@
 \centering
-\includegraphics[scale=1.0]{./figures/ada-fenergy.pdf}
+\includegraphics[scale=1.0]{figures/ada-fenergy.pdf}
 \caption{The free energy as a function of the reaction 

### 2022-06-07 tex/resubmission1/response/response.tex
added the new files

\pdfoptionpdfminorversion=4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcbfk,manuscript=article,layout=traditional]{achemso}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage[scaled=0.92]{helvet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\usepackage{microtype,xparse,tcolorbox}
\usepackage{helvet}
\newcommand{\newtext}[1]{{\color{blue} #1}}
\newcommand{\vamsee}[1]{{\color{applegreen} #1}}
\newcommand{\filipp}[1]{{\color{red} #1}}

\newenvironment{reviewer-comment }{}{}
\tcbuselibrary{skins}
\tcolorboxenvironment{reviewer-comment }{empty,
  left = 1em, top = 1ex, bottom = 1ex,
  borderline west = {2pt} {0pt} {black!20},
}
\ExplSyntaxOn
\NewDocumentEnvironment {response} { +m O{black!20} } {
  \IfValueT {#1} {
    \begin{reviewer-comment~}
      \setlength\parindent{2em}
      \noindent
      \ttfamily #1
    \end{reviewer-comment~}
  }
  \par\noindent\ignorespaces
} { \bigskip\par }

\NewDocumentCommand \Reviewer { m } {
  \section*{Comments~by~Reviewer~#1}
}
\ExplSyntaxOff
%\AtBeginDocument{\maketitle\thispagestyle{empty}\noindent}

% You can get probably get rid of these definitions:
%\newcommand\meta[1]{$\langle\hbox{#1}\rangle$}
%\newcommand\PaperTitle[1]{``\textit{#1}''}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Sree Ganesh Balasubramani}
\affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
\author{Steven D. Schwartz}
\affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
%\author{I. Ken Groupleader}
%\altaffiliation{A shared footnote}
\email{sschwartz@email.arizona.edu}
%\phone{+123 (0)123 4445556}
%\fax{+123 (0)123 4445557}
%\affiliation[Unknown University]
%{Department of Chemistry, Unknown University, Unknown Town}
%\alsoaffiliation[Second University]
%{Department of Chemistry, Second University, Nearby Town}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]
  {Transition path sampling based calculations of free energies for enzymatic
  reactions: the case of human methionine adenosyl transferase and plasmodium 
  vivax adenosine deaminase}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{TPS,WHAM}
\keywords{American Chemical Society, \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%\date{\today}

\maketitle

%Additions to the manuscript are indicated in  \textit{\newtext{blue text}}.

\Reviewer{1}

We would like to thank the Reviewer for his/her time and for evaluating
the manuscript.

%\begin{response}
%{The current work shows the application of the BOLAS method to two enzyme-catalyzed reactions. 
%The BOLAS method is a transition path sampling algorithm that can be extended to calculate free 
%energies along a pre-defined reaction coordinate. There are already several methods to calculate 
%free energy profiles and many have been applied to enzymatic reactions, but none is a black-box, 
%and therefore the implementation and application of this method to an enzymatic reaction is an 
%important advance. The text is well written and clear.}
%reply
%TPS based free energy calculations have been attempted previously for the calculations
%involving model problems involving analytical potential energy surfaces or for 
%configurational or structural changes of an enzyme. A chemical reaction catalyzed by 
%an enzyme can involve a complicated reaction coordinate and TPS based method using the 
%shooting algorithm provides a bias free technique for calculating the free energy
%of the reaction.

%\end{response}

\begin{response}
{First one needs to know how the specific parameters of the method affect 
the results. Or at least, how or why they were chosen. If not for all parameters, 
at least for the most relevant ones (but this needs also to be specified). 
For example:
* why use 150 reactive trajectories?}
%reply

%The application of TPS for stuyding the kinetics and transition states of 
%enzyme catalyzed reactions has been repeorted several times by our group as well
%as by others we decided to not list out every detail. But we have now included 
%some details about the choice the number of trajectories and so on.
The number of reactive trajectories in a TPS ensemble is decided based 
on how quickly the shooting algorithm results in decorrelation of the reactive 
trajectories from the initial biased trajectory. If the decorrelation is fast 
then fewer trajectoreis are sufficient for further committor analysis. 
To be sure that we do not use trajecotries that are correlated to the initial 
biased trajectory we ignore the first 100 trajectories for committor analysis. 
This choice for the number of trajectories is also based on computational cost, 
while obtaining the additional trajectories may not provide any more information 
we use up computational resources in the process.

\end{response}

\begin{response}
  {why should each reaction last 0.5 ps? How is it known if that time is too short 
  or too long?}
We used longer trajectories (1 ps) and checked to make sure that the reaction occurs within
the time that we used (0.5 ps) before collecting the TPS ensemble.

\end{response}

{\begin{response}
{* why 20 windows?}
%reply
This choice was made based on the order parameter window that we are scanning.
20 windows results in a window size of 0.39 angstroms and we use a overlap of
0.09 angstroms on either side of the window to make sure that we have configurations
that are common to neighboring windows which is necessary to obtain a continuous free energy
profile. Smaller window sizes result in inefficient sampling because the TPS based method
does not include bias forces to keep the trajectories within the windows and we end up
sampling outside the windows especially in regions where the free energy has a steeper slope.

% We are somewhat at loss how to respond to this criticism,
% which seems to contradict Reviewer B's statements,
% ``I have to stress that the paper discusses complex material and still
% it is in general clearly written. It can be accessible for readers who
% are less familiar with RPA or many-body theory.''

\end{response}

\begin{response}
{* When calculating the commitor, you cannot say "repeating this procedure many times". 
How many are needed to obtain a given precision or to obtain the given figure?
\newline
* As before, when calculating the separatrix, you can not say "repeating this 
process for several trajectories".}
%reply
Since this is the discussion of the committor analysis and committor distribution 
analysis in general and not for the specific case for the enzymes we had not
provided the number of trajectories that were used in these calculations.

\end{response}

\begin{response}{Some of these questions arise because nowhere in 
the manuscript are confidence intervals given. Neither in the energy 
profile nor in the transition state structure distances (Fig. 5). 
Therefore the effect of these parameters on the precision of the 
results is not known.
}
The transition state structure depicted in Fig. 5 is a representative
from the ensemble of transition states that were obtained from committor
analysis. We could provide the mean and standard deviation of the bond distances
but this does not provide the any more

\end{response}

\begin{response}{Some of these questions also arise because the 
code to implement the method is not made public. Doing so would 
not only clarify some questions but also prevent others from 
reinventing the wheel, that is, coding the implementation again 
to check the results of this paper or to apply the method to their specific enzymes.}

% REPLY
We agree that the availability of code is necessary for reproducibility of the results.
TPS based free energies calculations are akin to vanilla TPS calculations with 
only a few modifications to the acceptance criteria as discussed in the manuscript as well
as in the original BOLAS paper. We followed the algorithm stated in the manuscript
and implemented it as python scripts but it was hard coded to model the MAT2A and 
$\textit{pv}$ADA catalyzed reactions. We would release it in a public repository once it is 
tailored for general purpose usage.  
\end{response}

\begin{response}{The second piece of information needed is how the method 
performs compared to others. Both in terms of computational cost and in 
terms of accuracy. Comparison with experiment is interesting but not very 
enlightening as the computational results depend on the method to calculate 
the free energy profile (tested here) but also on the chosen reaction coordinate, 
the QM region and the QM method, among others. A comparison with Umbrella Sampling, 
which also uses a reaction coordinate and also uses windows would be ideal. Again, 
to make fair comparisons we would need error bars to estimate the convergence of 
the results and reports of computational costs.}

% REPLY

\end{response}

\begin{response}{Beyond these general issues raised so far, there are some other points 
that deserve some attention.
\newline
1. How is the initial reactive trajectory generated? Although obtaining an initial reactive 
trajectory is not the crux of the algorithm it is a non-trivial step that should be more 
detailed. In particular, for the pvADA enzyme, which involves a two-step reaction, how 
do the way these 2 steps take place in the initial trajectory condition of the TPS ensemble?}

% REPLY
For the \textit{pv}ADA enzyme the two steps in the reaction are the nucleophilic attack 
of the OH$^{-}$ on the C6 atom and the transfer of the proton from the OH$^{-}$ 
to the N6 atom with the formation of NH$_3$. We are including the time series of the 
N6-H1 distances in the supporting information which would show the sequence of events 
along a typical TPS trajectory for this reaction.

\end{response}

\begin{response}{2. Related to the previous question, considering that the barriers 
surrounding the intermediate are higher than kT, shouldn't the trajectories remain 
trapped in the intermediate? In other words, isn't the duration of the whole trajectory 
determining the mechanism by enforcing a ballistic trajectory? Is the d(NC)-d(OC) still 
a good reaction coordinate to describe the second step? If not, shouldn't that underestimate 
the barrier, something that should be compensated by a (not calculated) transmission coefficient?}

% REPLY

\end{response}

\begin{response}
{3. I find the free energy profiles in fig 7 and 11 worrying. Take the 4th window in 
Fig. 7: it is clearly showing a smaller slope than the interpolation and this is because 
there are jumps between the windows. The first point of this window and the next one are 
"disconnected" from the other ones. Why is this so? Repeating the calculations with smaller 
windows would show the effect of window size on the result (that would address one of my 
questions regarding the choice of parameters).}
% REPLY
\end{response}

\begin{response}
{4. The author should check they cite the appropriate works. I am afraid 
ref 20 is not an application of TPS to enzyme catalysis. To my knowledge, The first 
applications of TPS to enzyme catalysis are  $J. Am. Chem. Soc.2005, 127, 13822-13831$ 
and $J. Phys. Chem. B 2007, 111, 20, 5708-5718$.}
% REPLY

We thank the Reviewer for pointing out this mistake. We have made the appropriate changes to that
citation in the manuscript. 
\end{response}

\begin{response}{Overall this is a relevant work, but the authors need to go 
deeper into the scientific details to be useful to other members of the 
community, because what the manuscript should be describing is not that the 
method works --we already knew that-- but how to make it work in enzymatic reactions.}
% REPLY
\end{response}

\section*{Other Changes}

\end{document}

### 2022-06-07 tex/resubmission1/supp-info/si.tex
added the new files

--- 2022-05-10 tex/manuscript.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jpcbfk,manuscript=article,layout=traditional]{achemso}
+\documentclass[journal=jpcbfk,manuscript=suppinfo,layout=traditional]{achemso}
 
@@
 %\begin{tocentry}
-
-%Some journals require a graphical entry for the Table of Contents.
-%This should be laid out ``print ready'' so that the sizing of the
-%text is correct.
-
-%Inside the \texttt{tocentry} environment, the font used is Helvetica
-%8\,pt, as required by \emph{Journal of the American Chemical
-%Society}.
-
-%The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-%permitted for  \emph{Journal of the American Chemical Society}
-%graphical table of content entries. The box will not resize if the
-%content is too big: instead it will overflow the edge of the box.
-
-%This box and the associated title will always be printed on a
-%separate page at the end of the document.
-
-%\end{tocentry}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The abstract environment will automatically gobble the contents
-%% if an abstract is not used by the target journal.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-  Transition path sampling (TPS) is widely used for the
-  calculations of reaction rates, transition state structures and reaction 
-  coordinates of condensed phase systems. Here we discuss a scheme 
-  for the calculation of free energies using the ensemble of TPS reactive 
-  trajectories in combination with a window based sampling technique 
-  %introduced by 
-  %Radhakrishnan et al. (Radhakrishnan, R.; Schlick, T., J. Chem. Phys. 2004, 121, 2436.) 
-  for enzyme catalyzed reactions. We calculate the free energy profiles of 
-  the reactions catalyzed by the human methionine S-adenosyltransferase (MAT2A) 
-  enzyme and the plasmodium vivax adenosine deaminase enzyme to assess 
-  the accuracy of this method. MAT2A catalyzes the formation of S-adenosine-L-methionine
-  following a S$_{\text{N}}2$ mechanism and using our method we estimate
-  the free energy barrier to be $16\;\text{Kcal}\;\text{mol}^{-1}$ which
-  is in excellent agreement with the experimentally measured activation 
-  energy of $17.27\;\text{Kcal}\;\text{mol}^{-1}$. Furthermore for the 
-  $pv$ADA enzyme catalyzed reaction we estimate a free energy barrier of 
-  $23\;\text{Kcal}\;\text{mol}^{-1}$ and the calcualted free energy profile is
-  similar to that predicted from experimental observations.
-  We show that calculating free energies within TPS is lucrative as it is 
-  bias free, accurate compared to experimental measurements 
-  and computationally reasonable.    
-\end{abstract}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Start the main part of the manuscript here.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-Molecular dynamics (MD) simulations are increasingly being
-used to calculate quantitative estimations of experimentally measurable 
-kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
-\cite{Karplus02NatStructMolBiol9p646,
-Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
-Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
-This has helped in proposing mechanisms and explanations for 
-enzyme activity, fast high throughput screening of thousands of inhibitor 
-molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
-directed evolution for designing enzymes with enhanced catalytic 
-activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
-Straightforward MD simulations adequately sample 
-the long-lived stable reactant and product states 
-but the high energy barrier-crossing events that occur rarely and at long 
-time scales are not easily accessible using this method alone.
-To access such states at a higher frequency using MD simulations 
-several enhanced sampling techniques 
-have been developed in the past few decades some of which are 
-umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
-metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
-steered MD, \cite{Park04JChemPhys120p5946}
-milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
-(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,
-Dellago98JChemPhys108p1964,Bolhuis21AdvTheory4p2000237}
-
-TPS in particular is an attractive method since it samples reactive trajectories
-in which the rare barrier-crossing events are guaranteed to occur and
-it does not require a priori knowledge of the reaction coordinate which can be 
-complex for large biological systems.
-TPS is based on the Monte Carlo importance sampling in the 
-space of trajectories. It uses the shooting and shifting algorithm 
-to collect an ensemble of reactive pathways that are real dynamical 
-trajectories generated without the application of any bias forces. \cite{dellago02AdvChemPhys123} 
-Therefore the TPS ensemble can be used to obtain 
-kinetic information such as the rate of the reaction as well as to elucidate 
-the transition state, reaction coordinate and the reaction mechanism. 
-TPS has been successfully applied to study problems such as enzyme catalysis, 
-\cite{Antoniou01JPhysChemB105p5553,Hay12NatChem4p161,Schwartz09NatChemBiol5p551} 
-protein folding, \cite{Bolhuis03ProcNatlAcadSci100p12129,Juraszek12ChemPhys396p30} etc.
-
-Calculating equilibrium properties such as the free energy of an enzymatic 
-catalysis reaction within TPS is not straighforward. 
-The Monte Carlo algorithm used in TPS has an acceptance criterion that 
-is only satisfied if the trajectory begins in the reactant basin and ends in 
-the product basin or vice versa, hence trajectories that are localized in the 
-reactant or product basins are rejected. The localized trajectories contribute
-to the probability density distribution function ($P(\lambda)$) which is the 
-probability of a system at equilibrium to have a particular value of the 
-order parameter $\lambda$. \cite{Dellago09AdvCompSimAppp167}
-
-Frenkel has shown that inclusion of rejected states in Markov state Monte 
-Carlo algorith can help speed up the convergence of the calculations of 
-order-parameter distribution of a two-dimensional
-Ising model. \cite{Frenkel04ProcNatAcadSci101p17571}
-Inclusion of rejected trajectories within TPS was first discussed by Radhakrishnan 
-et al. \cite{Radhakrishnan04JChemPhys121p2436} who developed `BOLAS' which combines 
-the shooting and shifting algorithms of TPS with a window based sampling technique 
-in the spirit of umbrella sampling to calculate equilibrium free energy 
-differences. Peters et al. used a variant of the BOLAS scheme called the 
-equilibrium path sampling (EPS) where they use aimless shooting 
-combined with window based sampling for free energy calculations. 
-\cite{Peters08JAmChemSoc130p17342,Beckham10epsbook} 
-More recently Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
-have developed a reweighted path ensemble scheme within the standard TPS method 
-called virtual interface exchange TPS (VIE-TPS) to calculate the free energy 
-landscape. VIE-TPS is shown to produce reasonably accurate free energy profiles for 
-chemical reactions that are diffusive in nature whereas it is less accurate otherwise.
-Transition interface sampling (TIS) has been used to obtain 
-reweighted path ensemble which can give access to free energy 
-and committor surfaces. \cite{Rogal10JChemPhys17p174109}
-
-Free energy calculations based on the TPS shooting algorithm is 
-appealing but rare in the literature. Applications of this 
-method to study enzymatic reactions and comparisons of the calculated 
-free energies to experimental measurements can shine light on the 
-accuracy of this approach compared to the standard methods used for 
-free energy calculations.    
-In this manuscript we implement and apply the 
-algorithm developed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
-to calculate the free energies of reactions catalyzed by two systems:
-the human MAT2A enzyme and the plasmodium vivax adenosine deaminase 
-enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
-Human MAT2A catalyzes the formation of S-adenosyl L-methionine (SAM).
-SAM is an essential metabolite which is 
-distributed to almost all body tissues and fluids, is an universal 
-methyl donor and is of fundamental importance to the metabolism of 
-compounds such as hormones, neureotransmitters, proteins, 
-and nucleic acids. \cite{Friedel89Drugs38p389}
-plasmodium vivax is a parasite that is responsible for the most number 
-of cases of malaria globally. Targeting proteins that are essential to
-the metabolism of this bacteria such as the adenosine deaminase enzyme 
-is a good anti-malarial drug development strategy. \cite{Madrid08JBiolChem283p35899} 
-These enzymes were chosen not only because of their importance in 
-biochemical applications but also because of the availability of experimental 
-results with which comparisons can be drawn. This manuscript is organized as 
-follows: first we discuss the TPS method and the algorithm that we use 
-to calculate the free energies briefly, then we show calculations which apply this 
-method on the two enzymatic reactions. Finally we compare our results to expeiments
-and provide conclusions. 
-%------------------------------------------------------------------------------ 
-\section{Methods}
-%------------------------------------------------------------------------------ 
-\subsection{Transition path sampling}
-%------------------------------------------------------------------------------ 
-Here we briefly describe the equations describing the TPS method for the purpose of
-introducing notations as well as for the sake of being complete. 
-Consider the molecular dynamics simulation of a molecular system consisting 
-of $N$ atoms. Starting from the initial positions and momenta of each atom at time $t_0$,
-the time evolution is carried out by finding the position ($\textbf{q}$) and 
-momenta ($\textbf{p}$) of each atom at regular intervals of time $t_i = t_0 + i\Delta t$  
-($i = 0,1,2,\ldots$). At each of the time slices, the positions and 
-momenta can be collectively represented as
-the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is 
-run for a total 
-time of $\mathcal{T}$ the number of time slices is given by 
-$L = \mathcal{T}/\Delta t +1$ and for this sequence of 
-times the state of the system can be represented as 
-\begin{equation}
-z(\mathcal{T}) = \{z_0, z_{\Delta t}, z_{2\Delta t},\ldots,z_{\mathcal{T}}\}
-\end{equation}
-The probability to obtain a particular sequence of states is determined by the initial 
-conditions and the type of dynamics used for the time evolution. If the time evolution is 
-Markovian, the probability to go from $z_{i\Delta t}$ to $z_{(i+1)\Delta t}$ depends only 
-on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probablility can 
-be expressed as the product of individual probabilities $p(i\rightarrow j)$ as 
-\begin{equation}
-\mathcal{P}[z(\mathcal{T})] = \rho(z_0)\Pi_{i=0}^{\mathcal{T}/\Delta t-1} p(z_{i\Delta t}\rightarrow z_{(i+1)\Delta t}),
-\end{equation}
-where $\rho(z_0)$ is the equilibrium distribution of the initial conditions and for a canonical ensemble this can be 
-expressed as 
-\begin{equation}
-\rho(z_0) = \exp(-\beta H(z_0))/Z
-\end{equation}
-where 
-\begin{equation}
-Z = \int dz \exp(-\beta H(z)) 
-\end{equation}
-is the canonical partition function.
-
-Within the transition path sampling, the trajectories of interest start in the reactant region of the 
-phase space ($\mathcal{A}$) and end in the product region ($\mathcal{B}$). Such reactive trajectories have 
-a restricted probability distribution function given by
-\begin{equation}
-\mathcal{P}_{\mathcal{AB}}[z_{\mathcal{T}}] = h_{\mathcal{A}}(z_0)\mathcal{P}[z(\mathcal{T})]
-h_{\mathcal{B}}(z_{\mathcal{T}})/Z_{\mathcal{AB}}(\mathcal{T})\label{eqn:tpsensem}
-\end{equation}
-where 
-\[
-    h_{\mathcal{A}/\mathcal{B}}(z)= 
-\begin{cases}
-    1, & \text{if } z\in \mathcal{A}/\mathcal{B}\\
-    0,              & \text{otherwise}
-\end{cases}
-\]
-The set of all reactive trajectories characterized by the distribution function
-given by Eq. \ref{eqn:tpsensem} is the transition path ensemble.
-
-\subsection{Equilibrium distribution of order parameters}
-The free energy as a function of the order parameter $\zeta$ is defined
-as 
-\begin{equation}
-\text{A}(\zeta) = -k_{\text{B}}T\text{ln}(P(\zeta)) + C, \label{eqn:fenergy}
-\end{equation}
-where $P(\zeta)$ is the probability distribution of the reaction coordinate
-$\zeta$ and $C$ is an arbitrary constant. At equilibrium the probability distribution of the 
-reaction coordinate 
-can be obtained from the distribution function for the phase space $\rho(\textbf{q})$ as
-\begin{equation}
-P(\zeta) = \int d\textbf{V} \rho(\textbf{q})\delta\left[\zeta-\tilde{\zeta}(\textbf{q})\right].
-\end{equation}
-The integration is over the entire phase space. In practice the calculation of this distribution function
-often proceeds using histogram based methods where the reaction coordinate is 
-divided into bins and the frequency of the occurence of the reaction coordinate within a particular window
-during the course of molecular dynamics simulations is used to calculate the probability distribution. 
-Since some values of the reaction coordinates particulary close to the transition state are 
-rarely sampled in a conventional MD simulations, enhanced sampling techniques such as the umbrella sampling
-are necessary to sample the low probability regions. TPS is designed to sample these regions in phase space 
-without the application of any external bias, but the trajectories sampled with TPS are only reactive and hence
-they are not distributed according to the equilibrium distribution function. 
-The restriction that the pathways start from the reactant state and end in the product state can be relaxed
-to obtain an ensemble of trajectories that will be distributed according to the equilibrium distribution. 
-%------------------------------------------------------------------------------ 
-\subsection{Algorithm for free energy calculations}\label{ssec:algorithm}
-%------------------------------------------------------------------------------ 
-We follow the steps proposed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
-to sample TPS trajectories within windows of the order parameter. 
-\begin{itemize}
-\item An appropriate order parameter is chosen for the reaction of interest 
-that can distinguish the reactant and product states sufficiently well and the TPS ensemble is harvested 
-which consists only of reactive trajectories.
-\item Divide the order parameter into windows $\{\zeta_i\}$ within which the 
-order parameters are given by 
-$\zeta_{i}^{min} < \zeta < \zeta_{i}^{max}$.
-\item Choose a reactive trajectory $\tilde{z}(\mathcal{T})$ from the TPS ensemble as 
-a guiding trajectory for the window sampling. Pick a time slice from this trajectory 
-$\tilde{z}_{j\Delta t}$ such that the order parameter calculated for this time slice
-$\tilde{\zeta}$ satisfies $\zeta_{i}^{min} < \tilde{\zeta} < \zeta_{i}^{max}$.
-\item Using the shooting and shifting algorithm familiar from TPS simulations, \cite{dellago02AdvChemPhys123}
-harvest dynamics trajectories starting from the time slice $\tilde{z}_{j\Delta t}$ of the trajectory 
-$\tilde{z}(\mathcal{T})$ and accept the new trajectory if at any time slice the calculated 
-order parameter $\hat{\zeta}$ satisfies $\zeta_{i}^{min} < \hat{\zeta} < \zeta_{i}^{max}$.
-\item Calculate the probability distribution ($P_i(\zeta)$) within the window $\zeta_{i}^{min} 
-< \zeta < \zeta_{i}^{max}$ by constructing histograms obtained from the ensemble of 
-accepted trajectories for the window $i$. 
-\item Combine the probability distributions in successive windows by adjusting the constants $C$
-such that the free energy $A$ from Eq. \ref{eqn:fenergy} is continuous. 
-\end{itemize}
-%------------------------------------------------------------------------------ 
-\section{Computational details}
-%------------------------------------------------------------------------------ 
-\subsection{System preparation}
-%------------------------------------------------------------------------------
-We used the crystal structure of the human MAT2A enzyme 
-bound to methylthioadenosine and di-imido triphosphate (PNPNP)
-(protein data bank ID: 7L1A) reported by Niland 
-et al. \cite{Niland21Biochem60p791,Ghosh21JAmChemSoc143p18325} and the crystal 
-structure of \textit{pv}ADA in complex with MT-coformycin (protein data bank ID: 3EWC) 
-reported by Ho et al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
-simulations. For calculations involving the MAT2A enzyme the substrates were modified to 
-L-methionine and adenosyl triphosphate (ATP) whereas for calculations with the \textit{pv}ADA
-enzyme the substrates were modified to $2^{'}-$deoxyadenosine and hydroxide ion. 
-We solvated both the systems using TIP3P water molecules \cite{Jorgensen83JChemPhys79p926}
-in a nanodroplet sphere whose volume was fixed at $15$ {\AA}
-away from the protein's surface and the total charge was neutralized 
-using potassium ions.
-
-MD simulations within classical and hybrid quantum mechanics/molecular mechanics 
-(QMMM) approximations were carried out using the CHARMM program 
-package. \cite{Brooks83JComputChem4p187} 
-For the MM calculations the CHARMM36 force field \cite{Brooks09JComputChem30p1545} 
-was used and for the QM calculations the approximate PM3 semiempirical 
-method \cite{Repasky02JComputChem23p1601} was used. 
-For the hybrid QMMM simulations, the MAT2A system was partitioned into a QM
-region consisting of the ATP and MET molecules along with three Mg$^{2+}$ ions 
-and the MM region consists of the rest of the molecules. 
-For the $pv$ADA system the QM region consisted of 
-the adenosine molecule, an hydroxide anion, Zn$^{2+}$ ion and 
-a portion of the Glu229 residue which was partitioned within the generalized 
-hybrid orbital (GHO) scheme. \cite{Gao98JPhysChemA102p4714}
-
-The energies of both the systems were minimized using 50 steps of the 
-steepest descent method, followed by 2000 steps of the
-adopted basis Newton-Raphson method where only classical molecular mechanics 
-was used for the dynamics. 
-The minimized systems were
-heated slowly to 300 K for 35 ps beginning with harmonic
-constraints on all atoms except on the H atoms and the TIP3P 
-water molecules with gradual reduction of the restraint forces to 1. 
-15 ps of equilibration was carried out starting with harmonic 
-constraints of force constant 1 followed by 20 ps of contraint free 
-equilibration to prepare the systems for TPS simulations. 
-During the heating and equilibration steps, bonds that contain 
-hydrogen atoms in the MM region were restrained to their equilibrium values 
-using the SHAKE procedure. \cite{Ryckaert77JComputPhys23p327} 
-%------------------------------------------------------------------------------ 
-\subsection{Transition path sampling}
-%------------------------------------------------------------------------------
-TPS simulations require the definition of unique reactant and product states
-to collect reactive trajectories based on the Monte Carlo algorithm. Here we 
-discuss the details of the reactions involving the MAT2A and the $pv$ADA enzymes 
-to identify and define the order parameters characterizing the reactant and product basins.
-
-The reaction catalyzed by the human MAT2A enzyme is
-between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules 
-resulting in the formation of S-adenosyl-L-methionine (SAM) which is 
-depicted in Fig. \ref{fig:mat2a-reaction}. This reaction follows a S$_{\text{N}}2$ 
-mechanism and was experimentally characterized by 
-Firestone et al. \cite{Firestone17JAmChemSoc139p13754}. The key bond 
-parameters for this reaction consists of the distance between the
-nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as d$_{\mathrm{SC}}$), 
-and the distance between oxygen atom of the phosphate group and the $5'-C$ 
-atom of ATP (d$_{\mathrm{OC}}$). For the TPS calculations we define the 
-reactant state to have d$_{\mathrm{OC}}$ to be less than 1.7 {\AA} and the product 
-state to have the d$_{\mathrm{SC}}$ less than 1.7 {\AA}.
-
+\section{Initial reactive trajectory}
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{figures/mat2a-reaction.png}
-\caption{The reaction catalyzed by the human MAT2A enzyme between ATP and L-methionine resulting in the 
-formation of S-adenosyl L-methionine and triphosphate leaving group.}
-\label{fig:mat2a-reaction}
+\includegraphics[scale=1]{figures/ada-oh-nh60.pdf}
+\caption{The distance between the N6-H1 atoms and O1-H1 atoms as a function of time. The proton transfer from the hydroxide
+anion to the NH$_2$ group occurs after the nucleophilic attack of the O1 atom on the C6 atom.}
 \end{figure}
 
-$pv$ADA catalyzes the reaction between adenosine and OH$^{-}$ ion 
-resulting in the formation of inosine and ammonia as shown in
-Fig. \ref{fig:ada-reaction}.
-The nucleophilic O atom of the OH$^{-}$ anion
-attacks the electrophilic C6 atom of the adenosyl ring leading to the 
-formation of a Meisenheimer complex like structure while the N1 atom 
-on the adenosyl moiety gets protonated by the Glu229 residue of $pv$ADA. 
-The key bond parameters are the distance between the approaching hydroxyl 
-oxygen and the C6 atom (denoted as d$_{\text{OC}}$) and the 
-dissociating C6$-$N6 (d$_{\text{NC}}$) bond distance. For the TPS simulaions 
-we define the reactant state to have d$_{\text{NC}}$ being less than $1.7$ {\AA} 
-and the product state to have the d$_{\text{OC}}$ less than $1.7$ {\AA}.
-
-\begin{figure}
-\includegraphics[scale=0.6]{figures/ada-new.png}
-\caption{The reaction catalyzed by the $pv$ADA enzyme between adenosine and hydroxide ion
-resulting in the formation of inosine and ammonia.}
-\label{fig:ada-reaction}
-\end{figure}
-
-A biased initial reactive trajectory is obtained by using harmonic restraint forces 
-for both the MAT2A and $pv$ADA systems. 
-Using the biased trajectory as the starting point and after removing the restraint 
-forces, the shooting algorithm \cite{dellago02AdvChemPhys123} was used to generate
-new trajectories. The shooting algorithm begins with the choice of a 
-random time slice from the biased initial reactive trajectory and perturbs 
-the momenta of all the atoms in the system. Then the dynamics is propagated for 
-250 steps forawards and backwards using $1\;fs$ time steps
-resulting in a trajectory with a time period of $0.5\;ps$.
-A total of 150 reactive trajectories were generated for both the MAT2A and $pv$ADA
-systems.
-%------------------------------------------------------------------------------ 
-\subsection{Free energy calculations}
-%------------------------------------------------------------------------------ 
-We follow the algorithm discussed previously to calculate the free energy
-profile for the two enzymatic reactions using the TPS reactive trajectories. 
-The order paratemer for the MAT2A system 
-is defined to be d$_{\text{OC}}-$d$_{\text{SC}}$ and that for the $pv$ADA is defined as 
-d$_{\text{NC}}-$d$_{\text{OC}}$. For the MAT2A system, configurations are sampled
-for order parameter values within the range $[-2.0,2.0]$ and this range is divided 
-into 14 overlapping windows ($0.09$ {\AA} between neighboring windows). 
-Configurations are sampled for the $pv$ADA system for order parameter values 
-within the interval $[-3.0, 3.0]$ {\AA} which is divided into 20 overlapping 
-windows ($0.09$ {\AA} between neighboring windows). A reactive 
-trajectory from the TPS ensemble is chosen to guide the sampling 
-within each of the windows in the same spirit as a steered MD trajectory 
-is used to guide the window based umbrella sampling simulations. 
-To sample in a particular window $w$, a frame from this reactive 
-trajectory is chosen such that the value of its 
-order parameter falls within $w$. Using the shooting algorithm
-the momenta of all the atoms in the system are perturbed 
-and dynamics is run for 10 steps forwards and backwards with a 
-$1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.
-Longer trajectories in our experience end up sampling outside the window 
-more often and hence we make use of very short trajectories in this study.     
-Within each window 3000 TPS trajectories are collected. The 
-probability distribution function for the order parameter within window $w$ 
-is then calculated as normalized histograms which represent the populations 
-of the configurations within the window at equilibrium. 
-Finally we calculate the free energy profile in a similar fashion to the WHAM 
-procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
-potential of mean force from umbrella sampling calculations. 
-%------------------------------------------------------------------------------ 
-\section{Results and discussion}
-%------------------------------------------------------------------------------ 
-\subsection{Human MAT2A}
-The equilibrated structure of the MAT2A enzyme in complex with 
-ATP, MET and two Mg$^{2+}$ ions is shown in Fig. \ref{fig:mat2a-equil}.
-\begin{figure}[ht!]
-\centering
-\includegraphics[scale=0.2]{figures/mat2a-equil.png}
-\caption{Equilibriated structure of the MAT2A enzyme in complex with ATP and 
-MET along with 2 Mg$^{2+}$ ions.}
-\label{fig:mat2a-equil}
-\end{figure}
-The variation of the key distance parameters of the reaction calculated for 
-a reactive trajectory from the TPS ensemble are shown in Fig. \ref{fig:mat2a-reactive-traj}.
-\begin{figure}
-\includegraphics[scale=1.0]{figures/mat2a-diff167.pdf}
-\caption{Bond breaking ($\mathrm{d}_{\mathrm{OC}}$), bond forming 
-($\mathrm{d}_{\mathrm{SC}}$) distances for a typical 
-reactive trajectory for the reaction catalyzed by the MAT2A enzyme.}
-\label{fig:mat2a-reactive-traj}
-\end{figure}
-Committor analysis and committor distribution analysis are used to deduce 
-the transition states and the reaction coordinates, respectively. 
-Committor analysis is carried out for a reactive trajectory 
-from the TPS ensemble by choosing a time slice randomly from the trajectory, assigning
-momenta picked at random from the Boltzmann distribution for all the atoms, and running 
-dynamics for $250\;fs$. The commitment probability of a time slice is calculated by 
-repeating this procedure several times and finding the fraction of trajectories that go 
-to the reactant, product or neither. The transition state for the reaction is defined as 
-the structure from the time slice that has equal probability to commit to the reactant 
-and the product state. Repeating this process for several trajectories from the TPS
-ensemble results in a collection of transition states called the separatrix.
-\begin{figure}[ht!]
-\includegraphics[scale=0.12]{blender-images/ada/mat2a-trans-labelled.png}
-\caption{Transition state structure of the reaction catalyzed by the human MAT2A enzyme.
-The bond parameters d$_{\text{SC}}$ and d$_{\text{OC}}$ are $2.31$ {\AA} and $2.09$ {\AA}, 
-respectively.}
-\label{fig:mat2a-trans-struct}
-\end{figure}
-The transition state resembles a S$_{\text{N}}2$ structure with the 
-d$_{\text{SC}}$ and d$_{\text{OC}}$ distances being $2.31$ {\AA} and $2.09$ {\AA}, respectively.
-Experimentally the transition state structures of enzymatic reactions
-can be elucidated using kinetic isotope effects (KIE). \cite{Schramm99MetEnzym308p301}
-Firestone et al. \cite{Firestone17JAmChemSoc139p13754} used a combination of experimental 
-KIE measurements using the whole protein system and computational density functional 
-theory (DFT) based gas phase simulations of representative molecules to elucidate the 
-transition state structure for this reaction. From their findings they report 
-d$_{\text{SC}}$ and d$_{\text{OC}}$ distances of $2.03$ {\AA} and $2.32$ {\AA}, 
-respectively and our calculated values are in close agreement to this result.
-
-The reaction coordinate might contain not just the atoms from the molecules that 
-are directly involved in the reaction but also residues from side chains that can 
-promote the reaction. Taking this into account the committor distribution analysis
-starts by constraining the residues and molecules that are hypothesized to be part of the 
-reaction coordinate and using this constrained system as the starting point 
-trajectories are launched assigning momenta at random from the Boltzmann distribution at 
-$300\;\text{K}$ to all the atoms in the system. 
-First we constrained just the QM region in our committor distribution analysis 
-which consists of ATP and MET molecules along with two Mg$^{2+}$ ions and the 
-resulting distribution as shown in Fig. \ref{fig:mat2a-comm-dist} is not peaked 
-at 0.5 which is expected for the reaction coordinate. 
-Next we constrained the residues Gln113, Ser114, Arg249 and Arg264 along with the 
-QM region and obtain a distribution that is peaked at 0.5 and hence we 
-conclude that the reaction coordinate consists of these residues as well. 
-\begin{figure*}[ht!]
-\centering
-\begin{minipage}[b]{0.45\linewidth}
-\includegraphics[width=\textwidth]{figures/comm-60-mat2a-nocons.pdf}
-\label{fig:minipage1}
-\end{minipage}
-\quad
-\begin{minipage}[b]{0.45\linewidth}
-\includegraphics[width=\textwidth]{figures/comm-60-mat2a.pdf}
-\label{fig:minipage2}
-\end{minipage}
-\caption{Committor distribution analysis for obtaining the reaction coordinate of the 
-MAT2A catalyzed reaction. The figure on the left has the QM region constrained and the figure 
-on the right has the QM region along with the Gln113, Ser114, Arg249 and Arg264 residues constrained.}
-\label{fig:mat2a-comm-dist}
-\end{figure*}
-The activation energy of this reaction was experimentally measured by 
-Niland et al. \cite{Niland21Biochem60p791} where they measured the rate 
-constant of the reaction at various temperatures while keeping the 
-substrate concentrations near saturation. Using the Arrhenius equation 
-they estimated the activation energy to be $17.27\;\pm\;1.5\;
-\text{Kcal mol}^{-1}$ which is small Compared to typical enzyme catalyzed 
-reactions.
-\begin{figure}
-\includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
-\caption{The free energy as a function of the order parameter.}
-\label{fig:mat2a-fenergy}
-\end{figure}
-The free energy profile from our TPS based calculations is shown in 
-Fig. \ref{fig:mat2a-fenergy} where the barrier is $16\;\text{Kcal mol}^{-1}$.
-This is in excellent agreement with the experimental result.  
-%------------------------------------------------------------------------------ 
-\subsection{$pv$ADA}
-%------------------------------------------------------------------------------ 
-The equilibrated structure for the ADA enzyme complexed with adenosine, 
-OH$^{-}$ anion and Zn$^{2+}$ cation is shown in Fig. \ref{fig:ada-equil}.
-The Zn$^{2+}$ cation interacts with the His42, His44 and His226 residues
-and stabilizes the OH$^{-}$ anion in the pocket. The N1 atom on the adensoyl 
-molecule hydrogen bonds with the carboxylic acid hydrogen atom of the Glu229 
-residue.  
-\begin{figure}[ht!]
-\centering
-\includegraphics[scale=0.2]{./figures/ada-equil.png}
-\caption{Equilibriated structure of ADA protein}
-\label{fig:ada-equil}
-\end{figure}
-The bond forming (d$_{\text{OC}}$) and bond breaking (d$_{\text{NC}}$) distances 
-for a reactive trajectory from the TPS ensemble is depicted 
-in Fig. \ref{fig:ada-reactive-traj}.
-\begin{figure}[ht!]
-\includegraphics[scale=1.0]{figures/ada-diff60.pdf}
-\caption{Bond breaking ($\mathrm{d}_{\mathrm{NC}}$), bond forming 
-($\mathrm{d}_{\mathrm{OC}}$) distances for a typical 
-reactive trajectory for the reaction catalyzed by the $pv$ADA enzyme.}
-\label{fig:ada-reactive-traj}
-\end{figure}
-The deamination is an aromatic nucleophilic substitution reaction 
-where an intermediate Meisenheimer complex like structure is formed 
-at the transition state as shown in Fig. \ref{fig:ada-trans}. 
-The approaching O atom of the OH$^{-}$ anion is at a distance of $1.48$ {\AA}
-from the C6 atom whereas the N atom of the amine leaving group is at $1.60$ {\AA}.
-The N$1-$H bond distance is $1.1\;${\AA} 
-suggesting that the proton transfer from the Glu229 residue is almost complete.
-\begin{figure}
-\centering
-\includegraphics[scale=0.12]{blender-images/ada/new-ada-trans.png}
-\caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
-The formation of the Meissenheimer complex like structure with 
-d$_{\text{NC}}=1.60$ {\AA} and d$_{\text{OC}}=1.48$ {\AA} is observed.}
-\label{fig:ada-trans}
-\end{figure}
-These observations are consistent with the 
-experimental KIE studies from Luo et al. \cite{Luo07JAmChemSoc129p8008}
-who predict the transition state structure for the deamination reaction 
-catalyzed by the plasmodium falciparum ADA ($pf$ADA) enzyme which shares
-$72\%$ identity with the $pv$ADA enzyme that we analyze here. The next
-step in the reaction is the proton transfer from the OH$^{-}$ ion to the 
-amine leaving group resulting in the formation of inosine and ammonia.   
-%At the transition state, experimental findings 
-%suggest that the dissociation of the C6$-$N6 bond as well as the protonation of the N1 
-%atom are almost complete. \cite{Luo07JAmChemSoc129p8008}
-The free energy profile that we calculated is shown in Fig. \ref{fig:ada-fenergy}.  
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{./figures/ada-fenergy.pdf}
-\caption{The free energy in $\text{Kcal}\;\text{mol}^{-1}$ as a function of the reaction 
-parameter for the ADA catalyzed reaction. The Meissenheimer complex where the OH$^{-}$ anion
-is bonded to the C atom and the leaving NH$_2$ moeity is also still bonded.}
-\label{fig:ada-fenergy}
-\end{figure}
-The free energy barrier is calculated to be $23\;\text{Kcal}\;\text{mol}^{-1}$
-and the overall shape of the function is very similar to that hypothesized 
-by Luo et al. \cite{Luo07JAmChemSoc129p8008}
-%------------------------------------------------------------------------------ 
-\section{Conclusions}
-%------------------------------------------------------------------------------ 
-In this study, we presented an application of the TPS method for the calculation 
-of free energy profiles of enzymatic reactions. The order parameter along which the 
-free energy profile is to be obtained is partitioned into windows similar to 
-umbrella sampling. But instead of using bias potentials to sample the configurations
-within the windows the shooting algorithm is used to
-obtain an ensemble of trajectories according to the Monte carlo algorithm. A 
-trajectory is accepted if it visits the window at least in one of the time steps. 
-Using this ensemble of trajectories unbiased equilibrium probability distribution 
-function of the order parameter of the reaction within each window is obtained.
-Furthermore the free energy as a function of the order parameter is computed within 
-each of the windows and the entirte profile is obtained by shifting the free energies
-such that the function is continuous at the boundaries.
-
-The MAT2A enzymatic reaction was analyzed in detail to elucidate the transition state,
-reaction coordinate and the free energy profile. The reaction proceeds through a 
-S$_\text{N}$2 mechanism, the d$_{\text{OC}}$ and d$_{\text{SC}}$ are $2.1$ {\AA}
-and $2.3$ {\AA}, respectively. The reaction coordinate is determined to involve  
-the Gln113, Ser114, Arg249 and Arg264 residues. 
-The free energy barrier for this reaction was calculated to be 
-$16.0\;\text{Kcal}\;\text{mol}^{-1}$ which is in good agreement 
-with the experimentally measured activation energy of $17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. 
-Furthermore the reaction catalyzed by the $pv$ADA enzyme is analyzed in detail to obtain the  
-transition state and calculate the free energy profile.
-The transition state has a Meisenheimer complex like structure with the 
-d$_\text{NC}$ and d$_\text{OC}$ being $1.60$ {\AA} and $1.48$ {\AA}, respectively.
-The free energy barrier for this reaction for found to be $23\;\text{Kcal}\;\text{mol}^{-1}$
-with a double well shape similar to that measured from experiments.  
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "Acknowledgement" section can be given in all manuscript
-%% classes.  This should be given within the "acknowledgement"
-%% environment, which will make the correct section or running title.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgement}
-SGB acknowledges funding from ...
-
-\end{acknowledgement}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The same is true for Supporting Information, which should use the
-%% suppinfo environment.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{suppinfo}
-
-\end{suppinfo}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The appropriate \bibliography command should be placed here.
-%% Notice that the class file automatically sets \bibliographystyle
-%% and also names the section correctly.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{manuscript}
-
 \end{document}

### 2022-06-08 tex/resubmission1/manuscript.tex
added some references

--- 2022-06-05 tex/manuscript.tex
@@
 directed evolution for designing enzymes with enhanced catalytic 
-activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
+activities, \cite{Nunez06JPhysChemA110p463,Thyme09Nature461p1300,
+Bloom09PNAS106p9995,Zoi16JAmChemSoc138p3403,Schafer19JAmChemSoc141p10431} etc.    
 Straightforward MD simulations adequately sample 
@@
 TPS has been successfully applied to study problems such as enzyme catalysis, 
-\cite{Antoniou01JPhysChemB105p5553,Hay12NatChem4p161,Schwartz09NatChemBiol5p551} 
+\cite{Basner05JAmChemSoc127p13822,Ramon07JPhysChemB111p5708,Hay12NatChem4p161,Schwartz09NatChemBiol5p551} 
 and protein folding. \cite{Bolhuis03ProcNatlAcadSci100p12129,Juraszek12ChemPhys396p30}

### 2022-06-09 tex/resubmission1/response/response.tex
added a supporting information and writing some sections there

@@
 \author{Sree Ganesh Balasubramani}
-\affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
+\affiliation{Department of Chemistry and Biochemistry, University of Arizona, 
+Tucson, Arizona 85721, United States}
 \author{Steven D. Schwartz}
-\affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
+\affiliation{Department of Chemistry and Biochemistry, University of Arizona, 
+Tucson, Arizona 85721, United States}
 %\author{I. Ken Groupleader}
@@
 biased trajectory we ignore the first 100 trajectories for committor analysis. 
-This choice for the number of trajectories is also based on computational cost, 
-while obtaining the additional trajectories may not provide any more information 
-we use up computational resources in the process.
-
+Additionally we have added a discussion on the acceptance percentage of reactive 
+trajectories in the manuscript. 
 \end{response}
@@
 sampling outside the windows especially in regions where the free energy has a steeper slope.  
-
-% We are somewhat at loss how to respond to this criticism,
-% which seems to contradict Reviewer B's statements,
-% ``I have to stress that the paper discusses complex material and still
-% it is in general clearly written. It can be accessible for readers who
-% are less familiar with RPA or many-body theory.''
-
 \end{response}
@@
 provided the number of trajectories that were used in these calculations.
-
 \end{response}
@@
 }
+We have added a supporting information file which contains analysis on the transition
+state ensemble providing more information about the structures that we had calculated. 
 The transition state structure depicted in Fig. 5 is a representative
-from the ensemble of transition states that were obtained from committor
-analysis. We could provide the mean and standard deviation of the bond distances
-but this does not provide the any more
-
+from the ensemble of transition states.
 \end{response}
@@
 % REPLY
-For the \textit{pv}ADA enzyme the two steps in the reaction are the nucleophilic attack 
-of the OH$^{-}$ on the C6 atom and the transfer of the proton from the OH$^{-}$ 
-to the N6 atom with the formation of NH$_3$. We are including the time series of the 
-N6-H1 distances in the supporting information which would show the sequence of events 
-along a typical TPS trajectory for this reaction.
-
+The reaction catalyzed by the \textit{pv}ADA enzyme consists of 
+the nucleophilic attack of the OH$^{-}$ on the C6 atom, proton transfer from the 
+GLU 223 residue to the N1 atom in the adenosyl moeity and finally the 
+the transfer of the proton from the OH$^{-}$ to the N6 atom resulting in the formation 
+of NH$_3$ and inosine. We have included the time series of the key distances in 
+the supporting information which shows that we collect trajectories which contain all these
+steps taking the system from the reactant state to the product state with the order of events
+as predicted from experiments. 
 \end{response}

### 2022-06-09 tex/resubmission1/manuscript.tex
added a supporting information and writing some sections there

@@
 A total of 150 reactive trajectories were generated for both the MAT2A and $pv$ADA
-systems.
+systems. The percentage of TPS trajectories that were accepted were 24$\%$ and 
+19$\%$ for the MAT2A and $pv$ADA systems, respectively.
+For optimal sampling with a good balance between the number of 
+accepted reactive trajectories and decorrelation from the initial constrained
+trajectory as well from one another, the acceptance percentage 
+is typically $\sim$ 40. \cite{dellago02AdvChemPhys123}  
+For complex systems such as enzymes we find that the lower the acceptance 
+percentage the quicker is the decorrelation. 
+%While a 40$\%$ acceptance has been found to provide the best compromise between the
+%number of accepted trajectories and decorrelation of the
+%trajectories from one another, 
+%we find that a lower acceptance
+%ratio provides quicker decorrelation from the initial constrained
+%trajectory, especially in more complex systems such as enzymes.
 %------------------------------------------------------------------------------ 

### 2022-06-09 tex/resubmission1/supp-info/si.tex
added a supporting information and writing some sections there

@@
 \usepackage{mathptmx}
+\usepackage{multirow}
+\usepackage{tabularx,booktabs}
+\usepackage{color, colortbl}
+\usepackage{subcaption}
 \usepackage[scaled=0.92]{helvet}
@@
 \section{Initial reactive trajectory}
+The reaction catalyzed by \textit{pv}ADA is between adenosine and OH$^{-}$ anion resulting
+in the formation of inosine. Experimental studies suggest that the GLU 223 residue 
+protonates the N1 atom of the adenosyl moeity and this transfer of proton is almost complete 
+at the transition state. \cite{Luo07JAmChemSoc129p8008} The protonation of the N1 atom 
+as observed in a representative reactive trajectory from the TPS ensemble is shown in 
+Fig. \ref{sfig:he2time}.
+
+\begin{figure}
+\includegraphics[scale=1]{figures/ada-ohe-nhe60.pdf}
+\caption{The distance between the N1-HE2 atoms (d$_{\text{NH}}$) and the OE2-HE2 atoms 
+(d$_{\text{OH}}$)in {\AA} as a function of time. The proton is 
+transferred from the GLU 229 residue onto the N1 atom of the adenosyl moeity.}
+\label{sfig:he2time}
+\end{figure}
+
 \begin{figure}
@@
 
+\section{Transition state ensemble for MAT2A enzyme}
+\begin{table*}[ht!]
+\caption{Order parameters d$_{\text{SC}}$ and d$_{\text{OC}}$ in {\AA} for the 11 transition states obtained from 
+the TPS ensemble along with the mean and standard deviation (Stddev).}
+\centering
+\begin{tabular}{l c c}
+\hline\hline
+Transition state & d$_{\text{SC}}$ ({\AA}) & d$_{\text{OC}}$ ({\AA})\\
+\hline
+1 & 2.30 & 2.10 \\
+2 & 2.31 & 2.08 \\
+3 & 2.31 & 2.09 \\
+4 & 2.34 & 2.14 \\
+5 & 2.35 & 2.17 \\
+6 & 2.41 & 2.16 \\
+7 & 2.41 & 2.11 \\
+8 & 2.40 & 2.15 \\
+9 & 2.35 & 2.13 \\
+10& 2.38 & 2.12 \\
+11& 2.29 & 2.06 \\
+\hline
+Mean & 2.35 & 2.12 \\
+Stddev & 0.04 & 0.03 \\
+\hline\hline
+\end{tabular}
+%
+\end{table*}
+
+\section{Transition state ensemble for \textit{pv}ADA enzyme}
+\begin{table*}[ht!]
+\caption{Order parameters d$_{\text{OC}}$ and d$_{\text{NC}}$ in {\AA} for the 11 transition states obtained from 
+the TPS ensemble along with the mean and standard deviation (Stddev).}
+\centering
+\begin{tabular}{l c c}
+\hline\hline
+Transition state & d$_{\text{OC}}$ ({\AA})& d$_{\text{NC}}$ ({\AA})\\
+\hline
+1& 1.48  &1.60 \\
+2& 1.45  &1.51 \\
+3& 1.48  &1.58 \\
+4& 1.5   &1.57 \\
+5& 1.55  &1.55 \\
+6& 1.55  &1.64 \\
+7& 1.52  &1.63 \\
+8& 1.52  &1.68 \\
+9& 1.49  &1.52 \\
+10&1.50  &1.64 \\
+11&1.49  &1.62 \\
+\hline
+Mean & 1.50 & 1.59 \\
+Stddev & 0.03 & 0.05 \\
+\hline\hline
+\end{tabular}
+%
+\end{table*}
+
+\bibliography{si}
+
 \end{document}

### 2022-06-10 tex/resubmission1/response/response.tex
added new figure and some more text

@@
 Additionally we have added a discussion on the acceptance percentage of reactive 
-trajectories in the manuscript. 
+trajectories in the manuscript. It has been observed that when the acceptance 
+percentage is low the trajectories decorrelate faster.  
 \end{response}
@@
   or too long?}
-We used longer trajectories (1 ps) and checked to make sure that the reaction occurs within
-the time that we used (0.5 ps) before collecting the TPS ensemble.
-
+We used longer trajectories (1 ps) during the initial stages of our calculations
+and checked to make sure that the reaction occurs within
+the 0.5 ps time length before collecting the TPS ensemble. 
 \end{response}
@@
 Since this is the discussion of the committor analysis and committor distribution 
-analysis in general and not for the specific case for the enzymes we had not
+analysis in general and not for the specific case for the enzymes, we had not
 provided the number of trajectories that were used in these calculations. 
@@
 % REPLY
-
+The calculated geometry of the intermediate resembles that of the Meissenheimer complex which is
+similar to that predicted from experiments. In our TPS reactive trajectories we observe that 
+once the complex is formed there is a rapid proton transfer from the nucleophilic OH$^{-}$ 
+to the NH$_2$ group resulting in the release of NH$_3$ and the free energy change due to the 
+proton transfer is 1.7 kCal mol$^{-1}$. The order parameter that is chosen is based on the fact that
+it allows for the best definition of the reactant and product states.    
 \end{response}
@@
 % REPLY
+Since there are no bias forces constraining the trajectories to stay within a defined window, 
+in regions where the free energy curve has a large slope the trajectories tend to only sample 
+configurations almost completely outside the window with acceptance percentages $\leq 10$.
+This disconnection is simply an artifact of the histogram 
 \end{response}

### 2022-06-13 tex/resubmission1/response/response.tex
added new changes

@@
 at least for the most relevant ones (but this needs also to be specified). 
+\newline
 For example:
@@
 Additionally we have added a discussion on the acceptance percentage of reactive 
-trajectories in the manuscript. It has been observed that when the acceptance 
-percentage is low the trajectories decorrelate faster.  
+trajectories in the manuscript. 
+%It has been observed that when the acceptance 
+%percentage is low the trajectories decorrelate faster.  
 \end{response}
@@
 %reply
-Since this is the discussion of the committor analysis and committor distribution 
-analysis in general and not for the specific case for the enzymes, we had not
-provided the number of trajectories that were used in these calculations. 
+We have included a few lines in the committor analysis 
+discussion where the details of the number of trajectories used is mentioned.  
 \end{response}
@@
 }
-We have added a supporting information file which contains analysis on the transition
-state ensemble providing more information about the structures that we had calculated. 
-The transition state structure depicted in Fig. 5 is a representative
-from the ensemble of transition states.
+We have added a supporting information file which contains geometrical 
+parameters of the transition state structures that we had calculated 
+along with the mean and standard deviations. 
+The transition state structures depicted in Figs. 5 and 10 are 
+representatives from the ensemble. We have also updated Figs. 7 
+and 11 to contain standard deviations of the free energy that were
+calculated using the bootstrapping method within the windows.  
 \end{response}
@@
 
+We have added discussions in the manuscript and a supporting information 
+document on the efficiency of sampling, 
+statistical analysis of the transition state structures and free energy profiles. 
 % REPLY
-
 \end{response}
@@
 to the NH$_2$ group resulting in the release of NH$_3$ and the free energy change due to the 
-proton transfer is 1.7 kCal mol$^{-1}$. The order parameter that is chosen is based on the fact that
-it allows for the best definition of the reactant and product states.    
+proton transfer is 1.7 kCal mol$^{-1}$. 
+%The order parameter that is chosen is based on the fact that
+%it allows for the best definition of the reactant and product states.    
 \end{response}
@@
 in regions where the free energy curve has a large slope the trajectories tend to only sample 
-configurations almost completely outside the window with acceptance percentages $\leq 10$.
-This disconnection is simply an artifact of the histogram 
+configurations that are almost completely outside the window with acceptance percentages $\leq 10$.
+This results in populations that are skewed in favor of regions with lower free energies. We sampled 
+3000 trajectories within each window and made sure that the change in the free energy profile
+is negligible as a function of number of sampled trajectories to make sure we have achieved 
+convergence. Moreover the largest standard deviations that we obtained from bootstrapping 
+analysis was 0.9 kCal mol$^{-1}$ which also provides a measure for convergence.
+
 \end{response}
@@
 \end{response}
-
+We thank the reviewer for the constructive comments to make the manuscript better. We have 
+addressed the questions by adding more statistical analysis and discussions of the details 
+of the calculations. 
 \section*{Other Changes}

### 2022-06-13 tex/resubmission1/manuscript.tex
added new changes

@@
 dynamics for $250\;fs$. The commitment probability of a time slice is calculated by 
-repeating this procedure many times and finding the fraction of trajectories that go 
+repeating this procedure at least 50 times for each time slice and finding the fraction of trajectories that go 
 to the reactant, product or neither. The transition state for the reaction is defined as 
@@
 and the product state. Repeating this process for several trajectories from the TPS
-ensemble results in a collection of transition states called the separatrix.
+ensemble results in a collection of transition states called the separatrix. We calculated 
+11 transition state structures for the MAT2A and the \textit{pv}ADA enzymes and details 
+regarding them are reported in the supporting information. 
 \begin{figure}[ht!]
@@
 The free energy profile from our TPS based calculations is shown in 
-Fig. \ref{fig:mat2a-fenergy} where the barrier is $16\;\text{Kcal mol}^{-1}$.
-This is in excellent agreement with the experimental result.  
+Fig. \ref{fig:mat2a-fenergy}. It includes standard deviations calculated 
+using the bootstrapping method for the populations of configurations 
+sampled within each window. \cite{Hub10JChemTheoryComput6p3713}
+The calculated free energy barrier is $16\;\text{Kcal mol}^{-1}$.
+which is in excellent agreement with the experimental result.  
 %------------------------------------------------------------------------------ 
@@
 \end{figure}
+The free energy profile along with standard deviations calculated 
+using the bootstrapping method for the populations of configurations 
+sampled within each window \cite{Hub10JChemTheoryComput6p3713} is 
+depicted in Fig. \ref{fig:ada-fenergy}.
 The free energy barrier is calculated to be $21\;\text{Kcal}\;\text{mol}^{-1}$
@@
 efficient predictions of both free energy profiles and mechanisms.
+Free energy calculations within the TPS shooting algorithm do not employ any 
+bias forces constraining the trajectories within a defined window. 
+We observe that in regions where the free energy curve has a large 
+slope the trajectories tend to majorly sample configurations that are 
+outside the window with acceptance percentages $\leq 10$. While this improves
+the decorrelation time of the sampled trajectories, techniques within 
+TPS to enhance the sampling in these regions will be necessary for its
+widespread applications to complex biomolecular systems.  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{suppinfo}
+\begin{suppinfo}
+The supporting information is available free of charge at
 
-%\end{suppinfo}
+\end{suppinfo}
 

### 2022-06-13 tex/resubmission1/supp-info/si.tex
added new figure and some more text
added new changes

@@
 %\begin{tocentry}
-\section{Initial reactive trajectory}
+\section{\textit{pv}ADA catalyzed reaction}
 The reaction catalyzed by \textit{pv}ADA is between adenosine and OH$^{-}$ anion resulting
@@
 as observed in a representative reactive trajectory from the TPS ensemble is shown in 
-Fig. \ref{sfig:he2time}.
+Fig. \ref{sfig:he2time}. The synamics of the proton transfer from the OH group to the 
+NH$_2$ group which are both bonded to the adenosine moeity forming a Meissenheimer complex is 
+depicted in Fig. \ref{sfig:n6proton}.
 
@@
 anion to the NH$_2$ group occurs after the nucleophilic attack of the O1 atom on the C6 atom.}
+\label{sfig:n6proton}
 \end{figure}
@@
 
+%\section{Bootstrapping analysis of the free energy curves}
+
 \bibliography{si}

### 2022-06-13 tex/resubmission1/response/response.tex
added all of Steve's suggestions

@@
 %and therefore the implementation and application of this method to an enzymatic reaction is an 
-%important advance. The text is well written and clear.}
-%reply
+%important advance. The text is well written and clear.\newline
+%\begin{response}
+%Having said that, one has to keep in mind that this work does not introduce a new method, but 
+%applies a known method to some new systems. For these types of contributions to be really useful 
+%to the scientific community, one has to address the specific requirements or adaptations needed 
+%for these new systems and what are the advantages. Some of these aspects are missing in the 
+%current version of the manuscript. In other words, the paper has to explore and describe the 
+%necessary details to help others apply the method and decide whether it is worth applying.}
+%REPLY
+
 %TPS based free energy calculations have been attempted previously for the calculations
@@
 %of the reaction. 
-
 %\end{response}
@@
 
+The number of reactive trajectories in a TPS ensemble is determined 
+by how quickly the shooting algorithm results in decorrelation of 
+the reactive trajectories from the initial biased trajectory. If the 
+decorrelation is rapid, then fewer trajectories are sufficient for 
+further committor analysis. To be sure that we do not use trajectories 
+that are correlated to the initial biased trajectory we ignore the 
+first 100 trajectories for committor analysis. We have added a discussion 
+on the acceptance percentage of reactive trajectories in the manuscript.
+
 %The application of TPS for stuyding the kinetics and transition states of 
@@
 %some details about the choice the number of trajectories and so on.
-The number of reactive trajectories in a TPS ensemble is decided based 
-on how quickly the shooting algorithm results in decorrelation of the reactive 
-trajectories from the initial biased trajectory. If the decorrelation is fast 
-then fewer trajectoreis are sufficient for further committor analysis. 
-To be sure that we do not use trajecotries that are correlated to the initial 
-biased trajectory we ignore the first 100 trajectories for committor analysis. 
-Additionally we have added a discussion on the acceptance percentage of reactive 
-trajectories in the manuscript. 
+%The number of reactive trajectories in a TPS ensemble is decided based 
+%on how quickly the shooting algorithm results in decorrelation of the reactive 
+%trajectories from the initial biased trajectory. If the decorrelation is fast 
+%then fewer trajectoreis are sufficient for further committor analysis. 
+%To be sure that we do not use trajecotries that are correlated to the initial 
+%biased trajectory we ignore the first 100 trajectories for committor analysis. 
+%Additionally we have added a discussion on the acceptance percentage of reactive 
+%trajectories in the manuscript. 
 %It has been observed that when the acceptance 
@@
 20 windows results in a window size of 0.39 angstroms and we use a overlap of 
-0.09 angstroms on either side of the window to make sure that we have configurations
+0.09 {\AA} on either side of the window to make sure that we have configurations 
 that are common to neighboring windows which is necessary to obtain a continuous free energy 
@@
 sampling outside the windows especially in regions where the free energy has a steeper slope. 
+To be honest, such choices lie in the art of doing such computations. 
 \end{response}
@@
 %reply
-We have included a few lines in the committor analysis 
-discussion where the details of the number of trajectories used is mentioned.  
+Completely reasonable criticism. We have included a few lines in the committor 
+analysis discussion where the details of the number of trajectories used is mentioned. 
 \end{response}
@@
 to check the results of this paper or to apply the method to their specific enzymes.}
-
 % REPLY
@@
 only a few modifications to the acceptance criteria as discussed in the manuscript as well 
-as in the original BOLAS paper. We followed the algorithm stated in the manuscript
-and implemented it as python scripts but it was hard coded to model the MAT2A and 
-$\textit{pv}$ADA catalyzed reactions. We would release it in a public repository once it is 
-tailored for general purpose usage.  
+as in the original BOLAS paper. To be completely honest, we thought we had invented the 
+method until a careful search found the original manuscript. We followed the algorithm 
+stated in the manuscript and implemented it as python scripts but it was hard coded 
+to model the MAT2A and pvADA catalyzed reactions. We would release it in a public 
+repository once it is tailored for general purpose usage.
 \end{response}
@@
 the results and reports of computational costs.}
-
-We have added discussions in the manuscript and a supporting information 
-document on the efficiency of sampling, 
-statistical analysis of the transition state structures and free energy profiles. 
-% REPLY
+%REPLY
+We have added discussions in the manuscript and a supporting information document 
+on the efficiency of sampling, statistical analysis of the transition state structures 
+and free energy profiles. We also must respectfully disagree – comparison with experiment 
+is exactly what we are interested in. The problems we have addressed are not “toy problems.” 
+There are full enzymatically catalyzed chemical reactions. The entire point to the 
+algorithm is the choice of coordinate along which windows are chosen is NOT a random guess – 
+it is what the actual dynamics of the system demands. Our goal is to provide evidence that 
+the method is capable of making mechanistic predictions PRIOR to experiment, and the way 
+for us to most efficiently do that is to verify our result by comparison to experiment – 
+not to other methods that have inherent limitations. 
 \end{response}
@@
 % REPLY
-The reaction catalyzed by the \textit{pv}ADA enzyme consists of 
-the nucleophilic attack of the OH$^{-}$ on the C6 atom, proton transfer from the 
-GLU 223 residue to the N1 atom in the adenosyl moeity and finally the 
-the transfer of the proton from the OH$^{-}$ to the N6 atom resulting in the formation 
-of NH$_3$ and inosine. We have included the time series of the key distances in 
-the supporting information which shows that we collect trajectories which contain all these
-steps taking the system from the reactant state to the product state with the order of events
-as predicted from experiments. 
+We thank the reviewer for this question. We have added a discussion in the manuscript
+which contains the details about how the initial biased reactive trajectories were
+generated using harmonic distance constraints to form and break bonds for both the 
+MAT2A and $pv$ADA systems. We have also added a discussion about the 
+time series of certain distance parameters for the $pv$ADA reaction which provide 
+information about the sequence of the reactions steps as observed in a typical 
+TPS trajectory.
+
+%The reaction catalyzed by the \textit{pv}ADA enzyme consists of 
+%the nucleophilic attack of the OH$^{-}$ on the C6 atom, proton transfer from the 
+%GLU 223 residue to the N1 atom in the adenosyl moeity and finally the 
+%the transfer of the proton from the OH$^{-}$ to the N6 atom resulting in the formation 
+%of NH$_3$ and inosine. We have included the time series of the key distances in 
+%the supporting information which shows that we collect trajectories which contain all these
+%steps taking the system from the reactant state to the product state with the order of events
+%as predicted from experiments. 
 \end{response}
@@
 % REPLY
-The calculated geometry of the intermediate resembles that of the Meissenheimer complex which is
-similar to that predicted from experiments. In our TPS reactive trajectories we observe that 
-once the complex is formed there is a rapid proton transfer from the nucleophilic OH$^{-}$ 
-to the NH$_2$ group resulting in the release of NH$_3$ and the free energy change due to the 
-proton transfer is 1.7 kCal mol$^{-1}$. 
-%The order parameter that is chosen is based on the fact that
-%it allows for the best definition of the reactant and product states.    
+The calculated geometry of the intermediate resembles that of the Meissenheimer complex 
+which is similar to that predicted from experiments. In our TPS reactive trajectories we 
+observe that once the complex is formed there is a rapid proton transfer from the 
+nucleophilic OH$^{-}$ to the NH$_2$ group resulting in the release of NH$_3$ and the free 
+energy change due to the proton transfer is 1.7 kcal mol$^{-1}$.
+
 \end{response}
@@
 convergence. Moreover the largest standard deviations that we obtained from bootstrapping 
-analysis was 0.9 kCal mol$^{-1}$ which also provides a measure for convergence.
+analysis was 0.9 kcal mol$^{-1}$ which also provides a measure for convergence.
 
@@
 ref 20 is not an application of TPS to enzyme catalysis. To my knowledge, The first 
-applications of TPS to enzyme catalysis are  $J. Am. Chem. Soc.2005, 127, 13822-13831$ 
-and $J. Phys. Chem. B 2007, 111, 20, 5708-5718$.}
+applications of TPS to enzyme catalysis are  \textit{J. Am. Chem. Soc.} 2005, 127, 13822-13831 
+and \textit{J. Phys. Chem. B} 2007, 111, 20, 5708-5718.}
 % REPLY
@@
 % REPLY
+We thank the reviewer for the constructive comments to make the manuscript better. We 
+have addressed the questions by adding more statistical analysis and discussions of 
+the details of the calculations in the manuscript and supporting information. 
 \end{response}
-We thank the reviewer for the constructive comments to make the manuscript better. We have 
-addressed the questions by adding more statistical analysis and discussions of the details 
-of the calculations. 
+
 \section*{Other Changes}
-
+We have provided a diff.pdf file which contains a color coded presentation of 
+all the changes that were made to the manuscript after we received the reviewer's 
+comments. 
 \end{document}

### 2022-06-13 tex/resubmission1/response/response.tex
Some final tweaks to the response

@@
 and checked to make sure that the reaction occurs within 
-the 0.5 ps time length before collecting the TPS ensemble. 
+the 0.5 ps time length before collecting the TPS ensemble. A statement regarding this 
+is added to the manuscript. 
 \end{response}
@@
 time series of certain distance parameters for the $pv$ADA reaction which provide 
-information about the sequence of the reactions steps as observed in a typical 
-TPS trajectory.
+information about the sequence of the steps that are part of the reaction as observed 
+in a typical TPS trajectory.
 
@@
 nucleophilic OH$^{-}$ to the NH$_2$ group resulting in the release of NH$_3$ and the free 
-energy change due to the proton transfer is 1.7 kcal mol$^{-1}$.
+energy change due to the proton transfer is 1.7 kcal mol$^{-1}$. We believe that d(NC)-d(OC) 
+is still a good reaction coordinate to describe the proton transfer step which is accompanied by 
+the release of NH$_3$.
 
@@
 Since there are no bias forces constraining the trajectories to stay within a defined window, 
-in regions where the free energy curve has a large slope the trajectories tend to only sample 
-configurations that are almost completely outside the window with acceptance percentages $\leq 10$. 
+in regions where the free energy curve has a large slope the trajectories more often tend to sample 
+configurations that are outside the window with acceptance percentages $\leq 10$. 
 This results in populations that are skewed in favor of regions with lower free energies. We sampled 
 3000 trajectories within each window and made sure that the change in the free energy profile 
-is negligible as a function of number of sampled trajectories to make sure we have achieved 
+is negligible as a function of the number of sampled trajectories to make sure that we have achieved 
 convergence. Moreover the largest standard deviations that we obtained from bootstrapping 
@@
 \section*{Other Changes}
-We have provided a diff.pdf file which contains a color coded presentation of 
-all the changes that were made to the manuscript after we received the reviewer's 
-comments. 
+We have provided a diff.pdf file which highlights all the changes that were made to 
+the manuscript after we received the reviewer's comments. 
 \end{document}

### 2022-06-13 tex/resubmission1/manuscript.tex
More small changes

@@
 for both the MAT2A and $pv$ADA systems.
-Using the biased trajectory as the starting point and after removing the restraint 
+For the MAT2A system harmonic constraint with a force constant of 35 kcal mol$^{-1}$ {\AA}$^{-2}$
+was used for the bond breaking C5-O5 distance and the bond forming SD-C5 distance,
+and dynamics was run for $500\;fs$ starting with the reactant state and ending up in the product state.   
+For the $pv$ADA system harmonic constraints with a force constant of 40 kcal mol$^{-1}${\AA}$^{-2}$ 
+was used for the bond breaking (C6-N6) and bond forming (O1-C6) distances, and for the proton 
+transfer distances between the OE2 atom on the GLU229 residue to the N1 atom on 
+the adenosyl moeity and the dynamics was propagated for $500\;fs$ going from the reactant to 
+the product state. Using the biased trajectory as the starting point and after removing the restraint 
 forces, the shooting algorithm \cite{dellago02AdvChemPhys123} was used to generate
@@
 resulting in a trajectory with a time period of $0.5\;ps$.
+The trajectories were propagated for a longer $1\;ps$ time period 
+to make sure that the chemical reaction occurs within $0.5\;ps$. 
 A total of 150 reactive trajectories were generated for both the MAT2A and $pv$ADA

### 2022-06-13 tex/resubmission1/supp-info/si.tex
More small changes

@@
 The reaction catalyzed by \textit{pv}ADA is between adenosine and OH$^{-}$ anion resulting
-in the formation of inosine. Experimental studies suggest that the GLU 223 residue 
+in the formation of inosine. Experimental studies suggest that the GLU 229 residue 
 protonates the N1 atom of the adenosyl moeity and this transfer of proton is almost complete 
@@
 as observed in a representative reactive trajectory from the TPS ensemble is shown in 
-Fig. \ref{sfig:he2time}. The synamics of the proton transfer from the OH group to the 
+Fig. \ref{sfig:he2time}. The dynamics of the proton transfer from the OH group to the 
 NH$_2$ group which are both bonded to the adenosine moeity forming a Meissenheimer complex is 
